### 2023-05-12 FoodWebsCheck.Rmd
First commit

---
title: "Check consistency"
author: "L.A.S."
output: html_document
editor_options: 
  chunk_output_type: console
---

## Setup

```{r setup, eval=T,echo=FALSE }
#load(".RData")
require(igraph)
require(stringr)
require(devtools)
if( require(meweasmo) == FALSE)
  install_github("lsaravia/meweasmo")
if( require(multiweb) == FALSE)
  install_github("lsaravia/multiweb")
require(tidyverse)

```

# Check Weddell Sea

```{r WeddellSea, eval=F,echo=F,message=T,warning=T}

#
dtot <- read_csv("Data/WeddellSea_links.csv")

# Check valid names 
#
#
require(taxize)
dfn <- unique(dtot$consumer)
con_names <- gnr_resolve(dfn, best_match_only = TRUE, canonical=TRUE)
# Species not found
anti_join(data.frame(user_supplied_name=dfn),con_names)
# 1 Cadulus dalli antarcticum

dfn <- unique(dtot$resource)
res_names <- gnr_resolve(dfn, best_match_only = TRUE, canonical=TRUE)
anti_join(data.frame(user_supplied_name=dfn),res_names)
# 1             Phytodetritus
# 2                  Sediment
# 3         Silicioflagellata
# 4 Cadulus dalli antarcticum
gM <- graph_from_edgelist(as.matrix(dtot), directed  = T)

# Basal species
#
basal_sp <- (V(gM)[ (degree(gM,mode="in")==0) ])$name

# CLasificación completa
clasall <- classification(basal_sp, db = 'itis') # 24 not found

# Solamente clase
class <- tax_name(basal_sp, get = "class", db = "itis")
class %>% count(is.na(class))                                # 27 NA

## class = Bacillariophyceae

```