### 2023-09-29 manuscript.tex
Add manuscript Latex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[vmargin=1in,hmargin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}

\usepackage{pdflscape,booktabs}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage[running]{lineno}
\linenumbers
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\hypertarget{the-response-of-trophic-interaction-networks-to-multiple-stressors-in-a-marine-latitudinal-gradient-of-the-southern-hemisphere}{%
\section{The response of trophic interaction networks to multiple
stressors in a marine latitudinal gradient of the Southern
Hemisphere}\label{the-response-of-trophic-interaction-networks-to-multiple-stressors-in-a-marine-latitudinal-gradient-of-the-southern-hemisphere}}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Ecological networks are increasingly used as conceptual frameworks for
  understanding ecological community structure, identifying key species,
  and for ecosystem management and biodiversity conservation. Nowadays,
  there is a growing interest in understanding how ecological networks
  respond to environmental and anthropogenic stressors along
  geographical gradients.
\item
  We review for the first time the state-of-the-art knowledge on
  stressor response of marine food webs along the southwest Atlantic to
  Antarctic gradient (45 - 78ºS), considering the following study areas:
  San Jorge Gulf, Beagle Channel, Burdwood Bank, Scotia Sea, Potter Cove
  (Antarctica) and Weddell Sea (Antarctica). The aim of this review is
  threefold: 1) describe the complexity and structure of marine food
  webs along the southwest Atlantic to Antarctic axis from a network
  perspective; 2) identify the ongoing environmental and anthropogenic
  stressors for each marine ecosystem containing the food webs; and 3)
  resume proven food web changes and elaborate hypotheses on how the
  identified stressors might affect food web features, combining
  information on node- and network-level properties.
\item
  Environmental and anthropogenic stressor effects in the study gradient
  have been mostly assessed individually and at the organism and/or
  population, i.e., at the node level, with one exception: the effect of
  fisheries in the San Jorge Gulf food web. To address the most
  plausible stressors effects on the selected food webs, we built
  hypotheses for each study area considering the following: a)
  stressor(s), b) parameter(s) affected, c) node-level properties of the
  affected species, and d) network-level properties of the food web.
\item
  The most common stressor reported along the gradient is global
  warming, except for Beagle Channel and Burdwood Bank, which are more
  influenced by the introduction of an alien species and fisheries,
  respectively. We provide particular hypotheses on how warming might
  impact food webs.
\item
  We demonstrate the advantages of applying a network approach to better
  understand and predict the effect of stressors in marine ecosystems of
  the Southern Hemisphere.
\end{enumerate}

Keywords: anthropogenic stressors, environmental stressors, food webs,
latitudinal gradient, Southern Hemisphere

\hypertarget{introduction}{%
\subsection{1. Introduction}\label{introduction}}

The application of a network perspective has emerged as a powerful tool
to tackle the complexity of species interactions, facilitating a better
understanding of the structure and functioning of ecosystems (Belgrano,
Scharler, Scharler, Dunne, \& Ulanowicz, 2005; Thompson et al., 2012;
D'Alelio, Libralato, Wyatt, \& Ribera d'Alcalà, 2016). Trophic networks
(or food webs) allow identifying properties and key species that may be
crucial for ecosystem stability, and hence important for ecosystem
management and biodiversity conservation (Thompson et al., 2012). There
is a growing interest in understanding how ecological networks respond
to environmental and anthropogenic stressors along geographical
gradients (Baiser, Gotelli, Buckley, Miller, \& Ellison, 2012; Cirtwill,
Stouffer, \& Romanuk, 2015; Strydom et al., 2021; Bauer et al., 2022).
Yet, only a few studies have described variation in food web structure
along latitudinal gradients in marine ecosystems. The few that have come
from the Global North (Wood, Russell, Hanson, Williams, \& Dunne, 2015;
Kortsch et al., 2019; Frelat et al., 2022; Pecuchet et al., 2022),
whereas no studies, nor meta-analyses, on geographical variation in
marine food webs exist for the Global South (Southern Hemisphere).

Here we review for the first time the state-of-the-art knowledge on
stressor response of marine food webs along the southwest Atlantic to
Antarctic gradient (45 - 78ºS, Figure 1). We focus on proven and
expected changes in food webs driven by stressors in selected areas
along this large-scale latitudinal gradient. The aim of this review is
threefold: 1) describe the complexity and structure of marine food webs
along the southwest Atlantic to Antarctic axis from a network
perspective; 2) identify the ongoing environmental and anthropogenic
stressors for each marine ecosystem containing the food webs; and 3)
resume proven food web changes and elaborate hypotheses on how the
identified stressors might affect food web features (e.g.~energy flow,
stability), combining information on node- and network-level properties.
Finally, we suggest which additional data and analyses are needed to
gain insights into the stressors' effects on food web properties in the
southwest Atlantic to Antarctic region.

\hypertarget{the-structure-of-marine-food-webs-in-the-southwest-atlantic---antarctic-region}{%
\subsection{2. The structure of marine food webs in the Southwest
Atlantic - Antarctic
region}\label{the-structure-of-marine-food-webs-in-the-southwest-atlantic---antarctic-region}}

Together, the southwest and the Atlantic sector of the Southern Ocean
comprise one of the most biologically productive regions of the world's
oceans (Acha, Mianzan, Guerrero, Favero, \& Bava, 2004; Whitehouse et
al., 2008; Ballerini et al., 2014; Latorre et al., 2023). The referred
region extends from San Jorge Gulf (45ºS) in the Patagonian shelf to the
Weddell Sea (78ºS) in the Southern Ocean, and covers a well-connected
oceanic latitudinal gradient (R. P. Matano, Palma, \& Piola, 2010;
Steele, Thorpe, \& Turekian, 2010; Guihou, Piola, Palma, \& Chidichimo,
2020).

Throughout this latitudinal gradient, many investigations have been
carried out addressing the trophic ecology of specific species and
prey-predator relationships (Vinuesa \& Varisco, 2007; Sailley et al.,
2013; Pasotti et al., 2015; Riccialdelli, Newsome, Fogel, \& Fernández,
2017; Saunders, Hill, Tarling, \& Murphy, 2019; Riccialdelli et al.,
2020), but few studies have considered the complexity of the ecosystem
in terms of a high resolution of species and their prey-predator
interactions (but see Jacob et al. (2011), Marina et al. (2018), Funes,
Saravia, Cordone, Iribarne, \& Galván (2022), López-López, Genner,
Tarling, Saunders, \& O'Gorman (2022), Rodriguez, Marina, Schloss, \&
Saravia (2022), Marina et al. (in rev.)). Neglecting this complexity
might lead to a misunderstanding of the structure and functioning of the
ecosystems, and ultimately reduce the ability to predict ecosystem
responses to perturbations (Montoya, Woodward, Emmerson, \& Solé, 2009;
Woodward et al., 2010).

In this review, we consider marine areas in the Southern hemisphere for
which highly-resolved empirical food webs, in terms of species diversity
and trophic interactions, have been previously described. These areas
include: (1) San Jorge Gulf (45 - 47ºS, 65 - 68ºW), (2) Beagle Channel
(\textasciitilde54ºS, 68ºW), (3) Burdwood Bank (\textasciitilde54ºS,
59ºW), (4) Scotia Sea (48 - 58ºS, 50ºW), (5) Potter Cove (62ºS, 58ºW,
Antarctica), and (6) Weddell Sea (74 - 78ºS, 30ºW, Antarctica) (Figure
1). The geographic locations of these marine biomes span from temperate
to Antarctic, and are exposed to both environmental (sea warming,
glacial retreat) and anthropogenic (fishery, pollution) stressors.

\begin{figure}
\centering
\includegraphics[width=6.27in,height=3.528in]{Figures/Map_final.jpg}
\caption{Map of the study areas along the southwest Atlantic - Antarctic
latitudinal gradient. The areas are marked with numbers from one to six.
Smaller areas (1. San Jorge Gulf, 2. Beagle Channel, 3. Burdwood Bank,
and 5. Potter Cove) are shown on the panels to the left, whereas the
larger areas (4. North and South Scotia Sea, and 6. Weddell Sea) are
marked with a red rectangle on the map. The map was drawn using the
`marmap' R package (Pante, Simon-Bouhet, \& Irisson, 2023). Continental
contour shapefiles were obtained from www.ign.gob.ar.}
\end{figure}

San Jorge Gulf is the northernmost study area considered in this review.
It is a partially enclosed basin spanning approximately 34,000 km2 and
\textasciitilde100 m of maximum depths, located from Cabo dos Bahias
(44°55'S) to Cabo tres Puntas (47°06'S) (Figure 1). The shallower
northern and southern ends of the Gulf (with depths \textasciitilde40 m)
present two prominent frontal systems. These systems are the areas of
highest productivity (Bogazzi et al., 2005; Glembocki, Williams,
Góngora, Gagliardini, \& Orensanz, 2015). The Gulf's productivity
supports large invertebrate and vertebrate fisheries (María Eva Góngora,
González-Zevallos, Pettovello, \& Mendía, 2012), as well as marine
mammal and seabird populations (Yorio, 2009). The San Jorge Gulf food
web comprises 165 nodes and 1015 trophic interactions with a connectance
of 0.04. The percentage of top predators is 16\%, 78\% of intermediate
nodes and 6\% of basal nodes; 60\% of predators are omnivorous (Table
1). The most connected nodes are: the Argentine red shrimp Pleoticus
muelleri, the squat lobster Gimothea gregaria, squids (Illex argentinus
as dominant species) and Amphipoda. Notably, these nodes present
mid-trophic positions in the food web (3.0, 2.5, 3.6 and 2,
respectively) (Funes et al., 2022).

At a southern latitude, the Beagle Channel (54°52'S, 68°8'W) is an
interoceanic passage located at the southernmost tip of South America,
spanning 240 km in length and 5 km in width (\textasciitilde1200 km2),
and a range depth of 0 - 140 m (Figure 1). It features complex
coastlines, varying bathymetry, a prevailing west-to-east circulation
pattern, and a significant longitudinal gradient of glacial freshwater
discharge (Schloss et al., 2023). Bruno, Riccialdelli, Acha, \&
Fernández (2023) suggests that locally produced suspended particulate
organic matter (mostly composed by phytoplankton) and organic matter
accumulated in the sediments (macroalgae-originated detritus) are the
primary food sources for the marine ecosystem, as opposed to
allochthonous materials. Beagle Channel food web includes 145 nodes and
1115 trophic interactions with a connectance of 0.05 (Table 1). The food
web is suggested to have a wasp-waist structure, where the following
species play a crucial role in the dynamics of the ecosystem: Fueguian
sprat Sprattus fuegensis, longtail southern cod Patagonotothen ramsayi,
black southern cod P. tessellata, frogmouth Cottoperca trigloides, and
squat lobster Grimothea gregaria (Riccialdelli et al., 2020). Moreover,
the squat lobster has been identified as a species responsible for
linking modules and connecting the entire food web (Rodriguez et al.,
2022).

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II, meaning the shallow submarine
plateau named Burdwood Bank with a 200 m isobath boundary, and a deep
slope that reaches 4000 m in depth, respectively (Tombesi, Rabufetti, \&
Lovrich, 2020; Administración de Parques Nacionales, 2022) (Figure 1).
Physical features in the plateau are fairly stable, with salinity
averaging 34 all year round and temperature ranging between 4 and 8ºC
(Acha et al., 2004; Piola \& Falabella, 2009). The plateau is surrounded
by steep flanks of up to 4000 m depth, protected by the Marine Protected
Area Namuncurá - Burdwood Bank II (32,000 km2; \textasciitilde55º -
56ºS, \textasciitilde58º - 62ºW). Intense upwelling and mixing occur in
relation to the slope, entraining deep nutrient-rich waters into the
photic layer (Piola \& Falabella, 2009; Ricardo P. Matano, Palma, \&
Combes, 2019), and resulting in a fairly homogeneous water column both
spatially and temporally (Glorioso \& Flather, 1995; Ricardo P. Matano
et al., 2019). The Burdwood Bank food web comprises 379 nodes and 1788
interactions, with a connectance of 0.01, and an asymmetric degree
distribution (i.e.~most of the species have a relatively low number of
interactions and few species concentrate most of them). Almost half of
the consumers are omnivores (0.48), and the network displays a
small-world pattern (Marina et al., in rev.) (Table 1).

The Scotia Sea is a deep-sea basin, delimited by the Drake Passage to
the West and by the island complex of the Scotia Arc to the North, East,
and South, with an approximate extension of 1.5 x 106 km2 and a depth
range of 0 - 3000 m (E. j. Murphy et al., 2006) (Figure 1). Its
oceanography is dominated by the Antarctic Circumpolar Current, which is
spatially structured by frontal systems (Whitworth, 1980). The South
Antarctic Circumpolar Current Front subdivides the Scotia Sea into two
biogeographic regions: the Northern Scotia Sea is characterized by
higher and more variable temperatures, and the Southern Scotia Sea by
lower and more stable temperatures and influenced by seasonal sea ice
(Raymond, 2011). The analysis of the Northern and Southern Scotia Sea
food webs shows that the former is relatively more complex than the
latter: with higher species richness (218 vs 192) and interactions
(10008 vs 7241) and a higher connection overall (0.21 vs 0.20). As
expected from a more complex network, the path length is shorter in the
Northern Scotia Sea food web. In the same sense, the Southern Scotia Sea
network displays a greater proportion of omnivores and a lower mean
trophic level (López-López et al., 2022) (Table 1).

In the Antarctic realm, Potter Cove is a \textasciitilde9 km2 fjord with
a depth range of 0 - 200 m located at 25 de Mayo/King George Island
(62°14'S, 58°38'W, South Shetland Islands) on the West Antarctic
Peninsula (Figure 1). The cove, bordered by the Fourcade Glacier, is
divided into three areas: a) the internal cove, a high
glacier-influenced, soft sediment zone with a 50 m maximum depth; b) the
central cove, a mixed substrate area with low meltwater influence and an
80 m maximum depth; and c) the external cove, ice-free for 60 years with
a 185 m maximum depth and rocky substrate (Jerosch et al., 2018; Neder
et al., 2022; Neder, 2023). Potter Cove's high-latitude location results
in variable environmental conditions due to photoperiod length
seasonality. Sea ice often covers this area in winter (Schloss et al.,
2012). With low phytoplankton biomass, macroalgae, and microphytobenthos
are likely the primary food sources for secondary benthic production (M.
L. Quartino \& Boraso de Zaixso, 2008). The Potter Cove food web
includes 110 nodes and 649 interactions, with an asymmetric degree
distribution, and a connectance value of 0.05 (Table 1). It presents a
modular structure (groups of species interact more strongly with each
other than with species belonging to other groups), that's positively
associated with stability, since perturbation can be retained within
modules (Rodriguez et al., 2022).

Located between 74 and 78ºS, the high Antarctic Weddell Sea shelf spans
approximately 450 km from East to West (Jacob et al., 2011) (Figure 1).
The water depth ranges from 200 to 500 m, with shallower regions being
covered by continental ice that forms the coastline along the eastern
and southern parts of the Weddell Sea. Within this shelf area, exists a
complex three-dimensional benthic habitat characterized by substantial
benthic biomasses and an intermediate to high diversity when compared to
benthic boreal communities (Dayton, 1990; Teixidó, Garrabou, \& Arntz,
2002). The Weddell Sea food web exhibits a high level of network
complexity, featuring the greatest number of nodes (490) and trophic
interactions (16041) among the analyzed food webs in this review (Table
1). Its connectance (0.07) and other structural properties are
intermediate compared to the other food webs. Recently, the interaction
strengths of this food web were estimated, revealing the presence of
numerous weak and few strong interactions, which is consistent with
findings in other complex food webs (McCann, Hastings, \& Huxel, 1998;
Nilsson \& McCann, 2016). Notably, this asymmetric distribution of
interaction strength is likely to promote community persistence.

\footnotesize

Table 1. Complexity and structure properties of the marine food webs
considered in the present review. Refer to Table 3 for definition of
properties. mean TL: mean trophic level. Food webs are ordered by
increasing latitude.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1538}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0839}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0839}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1259}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1259}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0979}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1049}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.2238}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Food web}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Nodes}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Links}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Connectance}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Path length}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{mean TL}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Omnivory}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Reference}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
San Jorge Gulf & 165 & 1015 & 0.04 & 2.17 & 3.02 & 0.63 & Funes et al.
(2022) \\
Beagle Channel & 145 & 1115 & 0.05 & 2.12 & 2.37 & 0.55 & Rodriguez et
al. (2022) \\
Burdwood Bank & 379 & 1788 & 0.01 & 2.99 & 2.52 & 0.49 & Marina et al.
(in rev.) \\
Potter Cove & 110 & 649 & 0.05 & 2.33 & 2.22 & 0.46 & Marina et al.
(2018); Rodriguez et al. (2022) \\
N Scotia Sea & 218 & 10008 & 0.21 & 1.87 & 3.29 & 0.73 & López-López et
al. (2022) \\
S Scotia Sea & 192 & 7241 & 0.20 & 1.90 & 3.21 & 0.71 & López-López et
al. (2022) \\
Weddell Sea & 490 & 16041 & 0.07 & 2.19 & 2.62 & 0.51 & Jacob et al.
(2011) \\
\end{longtable}

\normalsize

\hypertarget{environmental-and-anthropogenic-stressors-in-the-southwest-atlantic---antarctic-region}{%
\subsection{3. Environmental and anthropogenic stressors in the
Southwest Atlantic - Antarctic
region}\label{environmental-and-anthropogenic-stressors-in-the-southwest-atlantic---antarctic-region}}

A stressor is any environmental or anthropogenic variable that causes a
quantifiable change, irrespective of its direction (increase or
decrease), in a biological response (Côté, Darling, \& Brown, 2016; Orr
et al., 2020). In the light of this, the southwest Atlantic - Antarctic
marine biota has been and is currently subjected to a variety of
stressors (e.g.~sea warming, glacial retreat, ice changes,
acidification, species invasion, fisheries, and contamination). Although
it is well-known that multiple stressors act in concert at any given
time (e.g.~warming and fishery; acidification and contaminants) (Julian
Gutt et al., 2015, 2021), to date stressor assessments have been
performed individually. Moreover, the potential for interactive effects
of two or more stressors (synergy or antagonism) (Crain, Kroeker, \&
Halpern, 2008; Côté et al., 2016) is almost unknown for the region
(Rowlands et al., 2021).

In the following subsections, we describe the main environmental and
anthropogenic stressors, and the species (or trophic species) and
parameters affected reported for the marine ecosystems that contain the
food webs considered in this review. In Table 2 we provide an exhaustive
list of all stressors affecting species inhabiting each area,
considering: the type of stressor, species and parameter affected, and
locality.

\newpage
\begin{landscape}
\scriptsize

Table 2. Environmental and anthropogenic stressors reported for the
study areas: San Jorge Gulf, Beagle Channel, Burdwood Bank, Scotia Sea
(North and South), Potter Cove, and Weddell Sea. Stressor categories:
sea warming; glacial retreat; sediment in the water column; iceberg
scouring; sea ice extent; ocean acidification; ocean acidification +
plastics; microplastics; mercury; urban \& industrial pollution;
fishery; alien species. The species affected were considered at the node
level, whereas effects were considered at the organism and population
levels. Categories of affected parameters and variables: metabolism;
biomass; distribution; diet (see text for explanation). ``Locality''
indicates whether a stressor for a given species was reported for the
study area (`In situ') or in another area (`Elsewhere').

\begin{longtable}{ L{2.1cm} L{2.5cm} L{4cm} L{2.5cm} L{2cm} L{3cm} }
\hline
\textbf{Study area} & \textbf{Stressor} & \textbf{Species affected} & \textbf{Parameter affected} & \textbf{Locality} & \textbf{Reference} \\
\hline
\endhead
\endfoot
\hline
\endlastfoot
\textbf{San Jorge Gulf} & & & & & \\
& Sea warming & Fish assemblage \& its prey & Distribution & In situ &
Galván et al. (2022) \\
& Fishery & Demersal fish community & Biomass & In situ & Galván et al.
(2022) \\
& Fishery & Macroinvertebrates, fishes and seabirds & Diet & In situ &
González-Zevallos \& Yorio (2006); González-Zevallos \& Yorio (2011);
Yorio, González-Zevallos, Gatto, Biagioni, \& Castillo (2017) \\
& Alien species & Fish assemblage \& its prey & Distribution & In situ &
Galván et al. (2022); J. Ciancio, Beauchamp, \& Pascual (2010) \\
& Urban \& industrial pollution & Seabirds \& benthic assemblage &
Biomass & Elsewhere & Moore \& Dwyer (1974); Buskey, White, \& Esbaugh
(2016) \\
\textbf{Beagle Channel} & & & & & \\
& Urban \& industrial pollution & Macroalgae; Mytilus edulis chilensis,
Patagonotothen tessellata & Metabolism & In situ & Giarratano \& Amin
(2010); Ferreira, Lo Nostro, Fernández, \& Genovese (2021); Kaminsky et
al. (in prep.) \\
& Mercury & Phytoplankton, Zooplankton, \emph{Grimothea gregaria} &
Metabolism & In situ & Fioramonti, Ribeiro Guevara, Becker, \&
Riccialdelli (2022) \\
& Microplastics & Mytilus edulis chilensis, Nacella magellanica &
Metabolism & In situ & Analía F. Pérez et al. (2020); Ojeda et al.
(2021) \\
& Alien species: Chinook salmon Oncorhynchus tshawytscha &
Patagonotothen tessellata, Sprattus fuegensis & Diet/Biomass & In situ
(the presence), Elsewhere (changes in prey biomass) & Fernández,
Ciancio, Ceballos, Riva-Rossi, \& Pascual (2010); J. E. Ciancio,
Pascual, Botto, Frere, \& Iribarne (2008) \\
\textbf{Burdwood Bank} & & & & & \\
& Mercury & Dissostichus eleginoides, Sprattus fuegensis, Patagonotothen
ramsayi, Cottoperca trigloides (fishes) \& squids & Metabolism & In situ
& Fioramonti et al. (2022) \\
& Microplastics & Henricia obesa \& Odontaster penicillatus (sea stars);
Patagonotothen guntheri \& P. ramsayi (fishes) & Metabolism & In situ &
Cossi et al. (2021); A. F. Pérez et al. (2021) \\
& Fishery & Target: Dissostichus eleginoides. Bycatch: Macrourus sp.,
Coelorinchus sp. (fishes), Daption capense, Thalassarche melanophris,
Macronectes giganteus, T. chrysostoma, Diomedea epomophora (seabirds),
30+ spp macrobenthos & Biomass & Elsewhere & Gaitán \& Marí (2016);
Patricia A. Martínez, Wöhler, Troccoli, Di Marco, \& Maydana (2022);
Administración de Parques Nacionales (2022); Tamini et al. (2023) \\
& Fishery & Thalassarche melanophris, Macronectes giganteus, Daption
capense, Diomedea epomophora, M. halli, Procellaria aequinoctialis
(seabirds) & Diet & Elsewhere & Tamini et al. (2023) \\
\textbf{Scotia Sea, North \& South} & & & & & \\
& Mercury & Krefftichthys anderssoni, Protomyctophum bolini, Electrona
antarctica, Gymnoscopelus nicholsi, Gymnoscopelus braueri, Dissostichus
eleginoides & Metabolism & In situ & Seco et al. (2021) \\
& Ocean acidification + plastics & Euphausia superba, Limacina
retroversa (Pteropoda) & Metabolism & In situ & Rowlands et al. (2021);
Manno, Peck, Corsi, \& Bergami (2022) \\
& Sea warming & Euphausia superba & Metabolism & In situ & E. J. Murphy
et al. (2007); Perry et al. (2020) \\
& Sea warming & Euphausia superba & Distribution & In situ & Atkinson et
al. (2019) \\
& Fishery & Euphausia superba & Biomass & In situ & P. N. Trathan et al.
(2021) \\
\textbf{Potter Cove (Antarctica)} & & & & & \\
& Sediment in water column & Microphytobenthos, macroalgae, benthic
filter feeders (ascidians), pelagic filter feeders (krill, salps) &
Metabolism & In situ & Sahade et al. (2015); D. Deregibus et al. (2016);
Fuentes et al. (2016); Hoffmann et al. (2019) \\
& Sea ice extent & Krill & Biomass & Elsewhere & Flores et al. (2012) \\
& Sea ice extent & Benthic community (macroalgae, invertebrates) &
Metabolism & Elsewhere & Clark et al. (2013); Campana et al. (2018) \\
& Glacial retreat & Benthic community (macroalgae, invertebrates) &
Biomass & In situ & María Liliana Quartino, Deregibus, Campana, Latorre,
\& Momo (2013); Lagger, Servetto, Torre, \& Sahade (2017); Lagger et al.
(2018) \\
& Iceberg scouring & Benthic community & Biomass & In situ & D.
Deregibus, Quartino, Zacher, Campana, \& Barnes (2017); Dolores
Deregibus et al. (2023) \\
& Sea warming & Phytoplankton & Metabolism, biomass & In situ & Antoni
et al. (2020); Latorre et al. (2023) \\
& Sea warming & Zooplankton & Metabolism, biomass, diet & In situ &
Garcia et al. (2016); Garcia et al. (2019) \\
& Sea warming & Fish & Metabolism & In situ, elsewhere & Strobel, Leo,
Pörtner, \& Mark (2013); Souza et al. (2018); Saravia et al. (2021) \\
\textbf{Weddell Sea (Antarctica)} & & & & & \\
& Iceberg scouring & Macrobentos & Biomass & In situ & Isla (2023);
Julian Gutt et al. (2015); Dan A. Smale, Brown, Barnes, Fraser, \&
Clarke (2008); J. Gutt, Starmans, \& Dieckmann (1996) \\
& Iceberg scouring & Hexactinellida sponges & Biomass & In situ & Julian
Gutt \& Starmans (2001); Pineda-Metz, Gerdes, \& Richter (2020); Julian
Gutt (2001); Julian Gutt \& Piepenburg (2003) \\
& Ocean acidification & Primary producers / krill / foraminifera /
flagellates & Metabolism & Elsewhere & Deppeler et al. (2020); Isla
(2023); Julian Gutt et al. (2015); Moy, Howard, Bray, \& Trull (2009) \\
& Ocean acidification & Euphausia superba, Pleuragramma antarcticum &
Metabolism & In situ & So Kawaguchi et al. (2010); S. Kawaguchi et al.
(2013); Piñones \& Fedorov (2016); Mintenbeck et al. (2012) \\
& Sea warming & Large (diatoms) \& small (cryptophytes) phytoplankton,
zooplankton (salps) & Metabolism & Elsewhere & Isla (2023); Julian Gutt
et al. (2015); Trebilco, Melbourne-Thomas, \& Constable (2020) \\
& Sea warming & Euphausia superba, Nototheniid fishes, Pleuragramma
antarcticum & Metabolism & In situ & Meyer et al. (2017); Hill,
Phillips, \& Atkinson (2013); Mintenbeck et al. (2012); Constable et al.
(2014); Mintenbeck et al. (2012) \\
& Sea ice extent & Phytoplankton, Lobodon carcinophaga, Hydrurga
leptonyx, Leptonychotes weddellii, Ommataphoca rossii, Mirounga leonina,
Arctocephalus gazella & Metabolism & In situ & Pineda-Metz et al.
(2020); Wege, Salas, \& LaRue (2021); Siniff, Garrott, Rotella, Fraser,
\& Ainley (2008) \\
& Sea ice extent & Euphausia superba, Pleuragramma antarcticum,
Pagodroma nivea, Thalassoica antarctica, Pygoscelis adeliae & Metabolism
& In situ & Orgeira, Alvarez, \& Salvó (2021); Braithwaite, Meeuwig,
Letessier, Jenner, \& Brierley (2015); Hill et al. (2013); Mintenbeck et
al. (2012) \\
& Sea ice extent & Megaptera novaeangliae & Metabolism & Elsewhere &
Pallin et al. (2023) \\
& Sea ice extent & Aptenodytes forsteri & Metabolism & Elsewhere &
Orgeira et al. (2021) \\
\end{longtable}

\end{landscape}
\normalsize

\hypertarget{san-jorge-gulf}{%
\subsubsection{3.1 San Jorge Gulf}\label{san-jorge-gulf}}

San Jorge Gulf experienced several environmental and anthropogenic
stressors (Table 2). Trawl fisheries discard several species and add new
trophic interactions to the food web which resulted in a decrease in
trait variability and the stability of the system (Rincón-Díaz, Bovcon,
Cochia, Góngora, \& Galván, 2021; Funes et al., 2022). Moreover, it
changed the availability of prey to several predators. For example,
Merluccius hubbsi one of the main bycatch species, became prey item to
non-diving seabirds, like the kelp gull Larus dominicanus
(González-Zevallos \& Yorio, 2006) and reef fishes (Funes, Marinao, \&
Galván, 2019). Although juveniles of M. hubbsi are largely the main
bycatch item, 29 other cartilaginous and 69 bony local fish species were
also registered as incidental catch between 2005 and 2014 (María E.
Góngora, Bovcon, \& Cochia, 2009; Bovcon, Góngora, Marinao, \&
González-Zevallos, 2013; Ruibal Nuñez, 2020). This level of impact
triggered a shift in the functional diversity of the assemblages
homogenizing the trophic function of fishes (Rincón-Díaz et al., 2021).
Other functional changes were a decrease in the maximum sizes of
individual fish, together with a drop in elasmobranchs biomass and an
increase in crustaceans biomass (Funes, 2020). The significant increase
in crustacean biomass was mostly due to the increase in Pleoticus
muelleri and Grimothea gregaria populations (Funes, 2020). These species
rapidly became the most important prey for the most abundant fishes in
the area: M. hubbsi and Mustelus schmitti (Belleggia, Giberto, \&
Bremec, 2017; Pasti et al., 2021). However, the above-mentioned effects
of trawl fisheries on the structure and function of the San Jorge Gulf
community may have changed again, since the fisheries ceased activity in
2015 (Annex I, Resolution CFP No 7/2018), remaining a minimum of
artisanal activity.

Sea warming is another important environmental stressor in the San Jorge
Gulf, because of the southward shifts of northern fish populations to
the Gulf (Galván et al., 2022). The San Jorge Gulf is especially prone
to be affected by climate-driven shifts in species ranges, because it is
located in the ecotone between two biogeographic provinces, the
Argentine (30°S - 44°S) and the Magellanic (43°S - 55°S) (Balech \&
Ehrlich, 2008). In addition to this tropicalization from northern fish
populations, alien species are also documented to affect the demersal
assemblage (fish and macroinvertebrates) in the Gulf (Vinuesa, 2005 ;
Galván et al., 2022). Finally, San Jorge Gulf is exposed to urban and
industrial pollution due to an oil monobuoy from which oil manipulation
and general oil transport along the Patagonian coast registered several
oil spills and chronic oil discharges (García-Borboroglu, Boersma,
Reyes, \& Skewgar, 2008). Other marine systems impacted by oil spill
showed an important decrease in marine bird populations (Irons, Kendall,
Erickson, McDonald, \& Lance, 2000), seals and macroalgae (e.g. Garrott,
Eberhardt, \& Burn (1993); Paine et al. (1996)), with examples of
cascading effects throughout the food web (Paine et al., 1996; Peterson,
2001).

\hypertarget{beagle-channel}{%
\subsubsection{3.2 Beagle Channel}\label{beagle-channel}}

A main anthropogenic stressor in the Beagle Channel is the introduction
of invasive species (Table 2). Salmonidae were introduced to Tierra del
Fuego in the 1930s. Especially Chinook salmon Oncorhynchus tshawytscha
causes concern. The detection of Chinook salmon in Tierra del Fuego
dates back to April 2006, and its population has been expanding since
then (Fernández et al., 2010; Nardi, Fernández, Vanella, \& Chalde,
2019). Being a top predator Chinook salmon can compete with several
native species in the Beagle Channel (Correa \& Gross, 2008), and prey
over native species such as Notothenioids (Eleginops maclovinus,
Patagonotothen tessellata, P. cornucola, P. sima, Paranotothenia
magellanica, Harpagifer bispinis), the Atherinidae Odontesthes smitti
and O. nigricans, the Fuegian sprat Sprattus fuegensis, and larvae of
king crabs (Lithodes antarcticus and Paralomis granulosa) (Fernández et
al., 2010). J. E. Ciancio et al. (2008) observed that Chinook salmon in
the Southern Patagonian Shelf Ecosystem Area primarily feed on sprats
and display trophic levels comparable to those of intermediate-sized
fish and cephalopod predators species, showing significant dietary
overlap with Magellanic penguins (Spheniscus magellanicus). Another
potential competitor for Chinook salmon in the Beagle Channel is the
Commerson's dolphin (Cephalorhynchus commersonii), which shares a
similar diet (Riccialdelli et al., 2013).

In the Beagle Channel, other anthropogenic stressors include
contaminants like metals, perfluorinated compounds, hydrocarbons, and
microplastics found in animal tissue and sediments. Some studies
reported varying heavy metal levels in sea mussel tissues and sediments,
indicating bioavailability differences (Duarte, Giarratano, Amin, \&
Comoglio, 2011). Other research found similar carbon and carbohydrate
levels in Ushuaia Bay's surface sediments to those in hypertrophic
ecosystems, along with hydrocarbons and heavy metals linked to port and
industrial activities (Gil, Torres, Amin, \& Esteves, 2011).
Commendatore, Nievas, Amin, \& Esteves (2012) identified three distinct
hydrocarbon accumulation zones in the Bay. Llorca et al. (2012) found
high levels of perfluorinated compounds in algae, fish, and guano.
Ferreira et al. (2021) showed that black southern cod in Ushuaia Bay may
be exposed to endocrine-disrupting compounds from urban and industrial
pollution. Analía F. Pérez et al. (2020) and Ojeda et al. (2021) found
microplastics in M. edulis chilensis and Nacella magellanica,
respectively. These studies link pollutants to Ushuaia (54°48′S 68°18′W)
and its port and industrial activities. We expect the Beagle Channel
area closest to the city to be most affected. Contaminants can spread
through the food web, magnifying the risk to higher-level organisms.
Fioramonti et al. (2022) studied mercury transfer in three southwest
Atlantic Ocean food webs and found biodilution in the Beagle Channel and
Atlantic coast of Tierra del Fuego webs. However, mercury concentrations
rose with benthivory in these webs, and higher mercury was found in
phytoplankton and Grimothea gregaria. Squat lobsters connect pelagic and
benthic habitats, so any disruption to them could disrupt the food web
(Rodriguez et al., 2022). Dodino et al. (2022) found the highest mercury
levels in Magellanic penguin feathers from offshore colonies in Tierra
del Fuego. Recently, Ushuaia's kelp forests have seen a decrease in
biodiversity and changes in macroalgae composition due to urban
pollution (Kaminsky et al., in prep.).

\hypertarget{burdwood-bank}{%
\subsubsection{3.3 Burdwood Bank}\label{burdwood-bank}}

The stressors reported for the ecosystem of the Marine Protected Areas
Namuncurá within Burdwood Bank I and II are mostly anthropogenic; there
is a lack of studies focusing on the environmental stressors that are
known to be occurring at a regional scale (e.g.~sea surface warming,
Franco, Combes, \& González Carman (2020)) (Table 2). Several fisheries
targeting demersal fishes operate in the vicinity and within the
ecosystem of the Marine Protected Areas (i.e.~Marine National Reserve
management category). The Patagonian fisheries on the toothfish
Dissostichus eleginoides has gained prominence in recent years (P. A.
Martínez, Wöhler, \& Troccoli, 2015; L. Allega et al., 2020; Gorini,
Lukaszewicz, \& Giussi, 2021; Administración de Parques Nacionales,
2022). Although these are regulated by the Argentinean government,
incidental catches do occur, where demersal fishes of the genera
Coelorinchus and Macrourus, seabirds and benthic invertebrates (30+
taxa) are the most common bycatches (Gaitán \& Marí, 2016; Patricia A.
Martínez et al., 2022). Noteworthy, among the invertebrates caught, 8
species are indicator taxa of vulnerable marine ecosystems (Gaitán \&
Marí, 2016; Schejter \& Albano, 2021). Independent assessments of these
bycatches suggest no significant impact on the communities (Gaitán \&
Marí, 2016; Patricia A. Martínez et al., 2022). Besides bycatch, species
of seabirds are being affected by the fishery due to discards, altering
its diet; the most frequently encountered species are Black-browed
albatross Thalassarche melanophris, Southern giant petrel Macronectes
giganteus, Cape petrel Daption capense, Southern royal albatross
Diomedea epomophora, Northern giant petrel M. halli, and White-chinned
petrel Procellaria aequinoctialis (Tamini et al., 2023). Nevertheless,
there is a lack of knowledge considering the species' role in the
ecosystem and the potential joint effect of both target fishes and
bycatch in a broader food web framework.

Anthropogenic contaminants such as microplastics and mercury are present
in the water column of the Burdwood Bank ecosystem (Cossi et al., 2021;
Di Mauro et al., 2022; Fioramonti et al., 2022). Microplastics are
distributed all along the water column, from surface to deep waters
(3-2450 m) (Di Mauro et al., 2022). More importantly, microplastics were
found in soft tissues of benthic macroinvertebrates (sea stars Henricia
obesa and Odontaster penicillatus) and benthopelagic fishes
(Patagonotothen guntheri and P. ramsayi), which not only incorporated
the contaminant from the environment through their filter-feeding system
but could also get it indirectly from prey organisms already containing
plastics in their tissues (Cossi et al., 2021). Notably, one of the
contaminated species, the long tail southern cod Patagonotothen ramsayi,
is part of the core group of species that drive the ecosystem through
the suggested wasp-waist control (Riccialdelli et al., 2020). Mercury
transfer and biomagnification are ongoing processes in the ecosystem,
which are occurring at a greater pace than near coastal areas (Beagle
Channel) (Fioramonti et al., 2022). It's important to note that the
Fuegian sprat Sprattus fuegensis, a pelagic fish with a mid-trophic
level in the food web, presented the highest levels of mercury
(Fioramonti et al., 2022). Considering the wasp-waist control of the
Fuegian sprat in the food web (Riccialdelli et al., 2020), a rapid and
widespread contamination to the top predators is expected (Fioramonti et
al., 2022).

In recent years, licenses for seismic studies and exploration of
hydrocarbon resources have been granted all along the northwestern limit
of the Marine Protected Areas (Secretaría de Gobierno de Energía Res. Nº
65/2018). Although effects from these types of surveys on marine mammals
and seabirds are well-known for other regions of the world (Nowacek et
al., 2015), there is no particular knowledge for this ecosystem.
However, several documents warn of the potential negative effects this
may have on the species inhabiting the Burdwood Bank area (Lucrecia
Allega et al., 2019; de Haro et al., 2022).

Despite evidence of warming at surface, mid-water and bottom layers (100
m) in Burdwood Bank (Franco et al., 2020), specific studies on the
oceanographic aspects of the system are lacking. In this sense, there is
a big question mark about the environmental stressors impact on the
species and trophic interactions in Burdwood Bank.

\hypertarget{scotia-sea}{%
\subsubsection{3.4 Scotia Sea}\label{scotia-sea}}

The Scotia Sea is a vast and heterogenous oceanic region, where
especially the areas around South Georgia island represents an area of
interest, here referred to as `Northern Scotia Sea'. The majority of
studies analyzing the stressors' effects come from this area.

The Scotia Sea has experienced one of the largest levels of warming
within the polar regions (Whitehouse et al., 2008). Together with the
Southern Annular Mode anomalies this has caused a long-term decrease in
krill abundance; more pronounced in the northern than in the southern
Scotia Sea (E. J. Murphy et al., 2007; Atkinson et al., 2019). Over the
past 90 years, the krill also showed an increase in mean body length
(Atkinson et al., 2019), which may also alter predator-prey interactions
and allow reaching cooler feeding grounds near the seabed, with the
potential to link krill to unexpected predators (Schmidt et al., 2011).
Another consequence of the mentioned environmental stressors, is a
change in krill distribution due to a southward contraction (Atkinson et
al., 2019). It is noteworthy here, that the distributional shift has not
affected all species inhabiting the Northern Scotia Sea, and most
abundant calanoid copepods have maintained their distribution (Tarling,
Ward, \& Thorpe, 2018). Considering the above, food web models of the
Northern Scotia Sea have suggested shifts in the structure and
functioning of the food web, from krill-based to non-krill-based, where
myctophid fishes and squid are playing important roles (Hill, Keeble,
Atkinson, \& Murphy, 2012; Saunders et al., 2019).

The principal anthropogenic stressor of the Scotia Sea is commercial
fisheries. Krill fishery not only stresses the targeted species, but
also the many predators that depend on it as a food source (Hilborn et
al., 2017). Yet, data currently available from monitoring of krill and
its predators remain insufficient, hence identifying the potential
fishery impacts on the ecosystem is difficult (P. N. Trathan et al.,
2021). Apart from krill fishery, two other commercial fisheries operate
in the Scotia Sea, targeting Patagonian toothfish species Dissostichus
eleginoides and D. mawsoni. The D. eleginoides stock is linked to the
stock at South Georgia (`Northern Scotia Sea') (Collins, Brickle, Brown,
\& Belchier, 2010), while the D. mawsoni stock is linked to the
Antarctic continental shelf (`Southern Scotia Sea') (Soeffker et al.,
2022). Despite the fact that the mentioned fisheries are certified by
the Marine Stewardship Council standards, Philip N. Trathan (2023)
identified several concerns regarding aspects of fisheries itself, other
species' population status (i.e.~recovery of baleen whales), and
environmental stressors (i.e.~sea warming), which could have synergistic
effects on the food web structure and functioning that are still
unknown.

Mercury transfer and biomagnification are current processes occurring in
the Scotia Sea, where the total concentration of contaminants increase
with trophic level and are highest in notothenioid and myctophid fishes
(e.g.~Dissostichus eleginoides, Gymnoscopelus nicholsi), and seabirds
(Seco et al., 2021). During years of low Antarctic krill abundance,
predators must deal with both the stress of reduced prey availability
and the concurrent rise in mercury exposure (Seco et al., 2021).

\hypertarget{potter-cove-antarctica}{%
\subsubsection{3.5 Potter Cove
(Antarctica)}\label{potter-cove-antarctica}}

Regional warming in the last half century has been one of the main
factors driving changes in Potter Cove (Western Antarctic Peninsula)
(Hendry, Meredith, \& Ducklow, 2018; Chown et al., 2022). Sudden
environmental changes have occurred, such as sea surface temperature
increase, salinity decrease, suspended particulate matter loading and
chlorophyll-a increase, all linked to climatic cycles (Southern Annular
Mode and El Niño Southern Oscillation) (Schloss et al., 2012; Bers,
Momo, Schloss, \& Abele, 2013). Particularly, sea warming has produced
drastic environmental and biological transformations (e.g.~shifts in
dominance of benthic community) in the Potter Cove ecosystem (Rückamp,
Braun, Suckro, \& Blindow, 2011; Schloss et al., 2012; María Liliana
Quartino et al., 2013; Sahade et al., 2015; Hernández, Lopez, Piquet,
Mac Cormack, \& Buma, 2019), a system highly dependent on sea-ice
dynamics (Table 2). In Potter Cove, total sea ice cover has decreased
since 1991 (Schloss et al., 2012). Changes in the annual timing of
landfast ice formation and breakup of the sea ice cover has multiple
effects on species in the food web (Michel et al., 2019; Rossi et al.,
2019). Warmer winters and springs result in earlier sea-ice melt,
causing an abrupt increase in the light available benthic primary
producers (Dolores Deregibus et al., 2020). Sea ice also mediates
physical disturbances to the benthos by influencing sedimentation and
iceberg scouring. These factors affect the production of macroalgae,
albeit in opposite ways (D. Deregibus et al., 2016, 2017), and
microphytobenthos (Hoffmann et al., 2019; Braeckman et al., 2021). On
the other hand, sea ice is an important habitat for diatoms and its
associated consumers, including copepods and krill (Flores et al.,
2012), and thus important for bentho-pelagic nutrient and carbon cycling
during winter. Additionally, a decrease in winter sea ice cover produces
an increase in physical perturbation on benthic shallow communities in
coastal shallows due to ice scouring (D. Deregibus et al., 2017). The
glacier surrounding Potter Cove has been receding at an increasing rate
since 1950 (Rückamp et al., 2011), which has caused a massive discharge
of sediment-laden meltwater (Meredith et al., 2018). Large quantities of
suspended particles affect growth, survival and reproduction of benthic
species. This had led to a major shift in the benthic community
structure, from a filter feeders--ascidian domination to a mixed
assemblage with scavengers and opportunistic species (Pasotti et al.,
2015; Sahade et al., 2015), and the metabolic balance in benthos went
from net autotrophy to heterotrophy (Hoffmann et al., 2019; Braeckman et
al., 2021). Additionally, massive stranding events of the tunicate Salpa
thompsoni and the euphausiid Euphausia superba (krill) linked to the
presence of glacial meltwater have been reported (Fuentes et al., 2016).
Rising temperatures leading to ice and glacial melting has also
substantial impacts on pelagic primary productivity, since it reduces
water salinity, affecting water column stratification, light penetration
and nutrient availability for photosynthesis (Schloss et al., 2012). In
Potter Cove, changes in biomass of most phytoplankton species have been
observed under heat wave conditions, resulting in a shift from a
microplankton to a nanoplankton dominated community (Antoni et al.,
2020; Latorre et al., 2023). This means that in areas strongly affected
by glacier melt, the planktonic food web is dominated by the microbial
loop (ciliates and heterotrophic dinoflagellates preys upon
nanophytoplankton, which are sequentially available prey for small
omnivorous copepods), instead of being predominantly herbivorous (Garcia
et al., 2016, 2019). In addition, phytoplankton species under these
warming conditions showed a decrease in metabolic rates and in the
quality of the fatty acids composition (Latorre et al., 2023).

\hypertarget{weddell-sea-antarctica}{%
\subsubsection{3.6 Weddell Sea
(Antarctica)}\label{weddell-sea-antarctica}}

In the Weddell Sea several stressors have already been triggered by
global warming effects (Table 2), from which spatial and temporal
reduction in sea ice is suggested to be driving changes in pelagic and
benthic communities (Constable et al., 2014; Julian Gutt et al., 2021).
Sea ice extent has reached new record lows since the satellite era began
in 1978. Sea warming has been substantial in recent years: in 2017 the
mean temperature for February reached 1.45°C, the highest monthly mean
ever recorded (0.56°C above the climatological mean) (Turner et al.,
2020). In this context, it has been suggested that Antarctic krill
Euphausia superba has already declined as a result of productivity
changes caused by sea ice declines (Atkinson, Siegel, Pakhomov, \&
Rothery, 2004). Declining sea ice cover allows increased access to krill
by predators (So Kawaguchi, Nicol, \& Press, 2009), which further
contributes to decreased krill abundances. This in turn can result in
reduced carbon export due to decreased fecal pellets from krill (Pauli
et al., 2021).

On the other hand, sea ice loss increases light availability, triggering
primary production by phytoplankton in the short-term (Pineda-Metz et
al., 2020; Dinniman, St-Laurent, Arrigo, Hofmann, \& van Dijken, 2023).
Furthermore, a warmer and more stable water column with a shallow upper
mixed layer, is expected to enhance the dominance of cryptophytes over
diatoms which may favor salp populations over krill populations; thereby
reducing the magnitude of energy transfer to higher trophic levels and
the seabed (Isla, 2023).

Higher trophic level species are expected to respond differently to
global warming in the Weddell Sea. In this regard, Notothenioid fishes
(e.g.~Pleuragramma antarcticum) face multiple stressors (sea warming,
sea ice decline, ocean acidification) that threaten their survival.
While some species exhibit physiological plasticity to compensate for
increased oxygen demand, most notothenioid fishes are stenothermal and
unable to adjust their metabolic functioning (Mintenbeck et al., 2012).
It is anticipated that the thermal habitat preferred by the Antarctic
toothfish Dissostichus mawsoni could undergo contraction in the coming
three decades (Constable et al., 2014), further emphasizing the
potential impact of global warming effects on these species. Marine
mammals, Weddell and crabeater seals (Leptonychotes weddellii and
Lobodon carcinophagus), exhibit varying levels of sensitivity. The
latter exhibits a preference for breeding in close proximity to krill
swarms, and thus is particularly susceptible to reductions in sea ice
concentrations and sea surface temperatures. Rescued sea ice can
significantly affect their post- and future breeding foraging success
(Siniff et al., 2008; Wege et al., 2021). In addition, the Antarctic
petrel Thalassoica antarctica, which is one of the most abundant
seabirds in the area, is a sea ice dependent forager and breeder
(Orgeira et al., 2021); therefore the loss of this habitat will degrade
its foraging success and reproductive output over time. The snow petrel
Pagodroma nivea also relies heavily on sea ice for foraging and its
colonies in the Weddell Sea may be affected. The emperor penguin
Aptenodytes forsteri uses sea ice for breeding; any reductions could
constrain its populations, as evidenced by recent observations in the
Bellingshausen Sea region (Fretwell, Boutet, \& Ratcliffe, 2023). The
Arctic tern Sterna paradisaea spends the summer in the Weddell Sea
exploiting krill swarms under receding ice edges, then declines in ice
cover may reduce prey availability and quality of molting habitats.
Continued warming is thus expected to gradually erode the abundance and
distribution of these `primary species' that dominate the avifauna of
the Weddell Sea through loss of critical sea ice habitat (Orgeira et
al., 2021). Variations in sea ice extent have implications for great
whales, such as the Humpback whale Megaptera novaeangliae, since their
primary prey resource is the Antarctic krill (Braithwaite et al., 2015).
Decreased winter ice coverage results in reduced suitable habitat and
lower krill abundance (Braithwaite et al., 2015), with flow-on effects
for whale body condition observed historically in harvest data. With
climate projections indicating ongoing sea ice losses, further threats
to critical krill populations pose risks to the energy intake and
reproductive success and long-term viability of humpback whale
populations dependent on consistent Antarctic feeding (Pallin et al.,
2023).

Iceberg scouring is a major factor in the high biodiversity of benthic
communities in the Weddell Sea (Julian Gutt \& Starmans, 2001). Even at
600 m depths, iceberg scouring has a strong effect on the benthic
environment, disrupting the upper layers of the seabed and removing
macrofauna. This patchy disturbance and distribution pattern occurs
roughly every 200 square meters on the Antarctic continental shelf.
Global warming is predicted to raise iceberg scouring frequency (J. Gutt
et al., 1996; Julian Gutt, 2001; Dan A. Smale et al., 2008), disrupting
the environment (J. Gutt et al., 1996; Dan A. Smale \& Barnes, 2008).
Julian Gutt et al. (2015) expects tipping points are significant due to
these effects and lack of knowledge.

\hypertarget{from-species-stressors-to-food-web-effects}{%
\subsection{4. From species' stressors to food web
effects}\label{from-species-stressors-to-food-web-effects}}

A major challenge in contemporary ecology lies in predicting the effects
of stressors on complex multispecies systems, such as food webs. Network
analysis has proved to be a powerful tool to tackle this issue, since it
can capture the effects of individual and multiple stressors on
communities and ecosystems (Montoya et al., 2009; O'Gorman, Fitch, \&
Crowe, 2012; Bruder, Frainer, Rota, \& Primicerio, 2019; Polazzo,
Marina, Crettaz-Minaglia, \& Rico, 2022). Environmental and
anthropogenic stressor effects in the southwest and the Atlantic sector
of the Southern Ocean have been mostly assessed individually and at the
organism and/or population, i.e., at the node level (Table 2), with one
exception: the effect of fisheries in the San Jorge Gulf food web (see
section below for more details). To address the most plausible stressors
effects on the selected food webs, given the current information, we
built hypotheses for each study area. To this aim, we developed a
theoretical framework considering the following: a) stressor(s), b)
parameter(s) affected, c) node-level properties of the affected species,
and d) network-level properties of the food web. We considered that a
stressor will affect one of the following species' characteristics or
parameters: metabolism, biomass, distribution, and diet (Figure 2).
`Metabolism' refers to any change related to metabolic rate, such as
reproduction, hatching, larval development, growth and mortality, and
contamination due to pollutants (e.g.~growth effect in filter-feeders
due to sediment in water column in Potter Cove; endocrine disruption in
fish due to urban pollutants in Beagle Channel). `Biomass' indicates an
effect at the population level, where the density/abundance is being
impacted (e.g.~abundance decreases in macrobenthos due to iceberg
scouring in Weddell Sea). `Distribution' entails a change at the
population level in the geographic space occupied by a species,
e.g.~southward contraction of Antarctic krill due to sea warming of the
Scotia Sea. `Diet' includes alterations in the prey items of a species
at the population level, e.g, due to prey switching, having a direct
effect on the structure of the food web, e.g.~new prey item (discards)
for seabirds due to fishery activities in Burdwood Bank. Next, we
considered node- and network-level properties relevant to the
hypothesized stressor effects on the food webs, and which have been
previously calculated for the studied food webs (Table 3). At the
node-level, we included: a) degree, b) trophic position, c) omnivory
index, and d) relative abundance (see Table S1 in Supporting Information
for properties of stressed nodes). At the network-level, we considered:
a) connectance, b) path length, c) mean trophic level, and d) omnivory
(Table 1).

\begin{figure}
\centering
\includegraphics[width=6.27in,height=3.528in]{Figures/Figure2_StressorDiagram.png}
\caption{Conceptual diagram: from species' stressors to food web
effects. See text for explanation.}
\end{figure}

\scriptsize

Table 3. Node and network-level properties used to build hypotheses on
the stressors' effects.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1963}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2617}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3925}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1495}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Property}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Definition}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Relevance for stressor effects}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Reference}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\emph{Node-level} & & & \\
Degre & Number of feeding interactions in which the species participates
as prey and/or predator. & Perturbations to high-degree species may have
more significant effects on the food web robustness to perturbations
than low-degree species. & Dunne, Williams, \& Martinez (2002a); Jordán,
Benedek, \& Podani (2007) \\
Trophic position & Place in the food web relative to the basal resources
that support the community. Classifies species in: basal, intermediate
and top. & Perturbations on basal resources, intermediate species and
top predators are expected to have large effects on the rest of their
communities if ecosystem control is bottom-up, wasp-waist or top-down,
respectively. & Williams \& Martinez (2000); Thompson, Hemberg,
Starzomski, \& Shurin (2007) \\
Omnivory & Consumer resource use across trophic levels. & High-omnivore
species (generalists) are more flexible than low-omnivore species to
diet changes. & Thompson et al. (2007) \\
Relative abundance & Species' density in proportion to the other species
of the food web. & Perturbations on abundant (dominant) species are
expected to have large effects on the stability and energy flux of the
food and ecosystem, respectively. & Nilsson \& McCann (2016) \\
\emph{Network-level} & & & \\
Conectance & Proportion of actual interactions among possible ones. &
Estimator of community sensitivity to stressors. High connectance gives
resistance and resilience to the food web. & Dunne, Williams, \&
Martinez (2002b) \\
Path length & Average distance, accounted by the number of interactions,
between any pair of species. & Short distances enhance rapid and broad
propagation of perturbations. & Albert \& Barabási (2002) \\
Mean TL & Average of all species' trophic position contained in the food
web. & Influences the magnitude and efficiency of trophic transfer. A
higher mean food chain length reflects increased energy availability and
productivity. & Duffy et al. (2007); Olivier et al. (2019) \\
Omnivory & Proportion of species that feed at different trophic levels.
& It provides trophic flexibility to an ecosystem. Reduces probability
of trophic cascades. & Kratina, LeCraw, Ingram, \& Anholt (2012) \\
\end{longtable}

\normalsize

\hypertarget{warming-effects-on-food-webs-in-a-southwest-atlantic---antarctic-gradient}{%
\subsubsection{4.1 Warming effects on food webs in a southwest Atlantic
- Antarctic
gradient}\label{warming-effects-on-food-webs-in-a-southwest-atlantic---antarctic-gradient}}

The most common stressor reported along the southwest Atlantic -
Antarctic gradient is global warming, except for Beagle Channel and
Burdwood Bank, which are more influenced by the introduction of an alien
species and fisheries, respectively (section 3). The main
characteristics of global warming in the region, and the most plausible
drivers of change, are: sea warming, glacial retreat, elevated sediment
input in the water column, and reduction of the sea ice extent. These
drivers act in different ways and magnitudes in the studied locations
along the latitudinal gradient. The above does not mean that there are
no other stressors interacting with global warming, potentially
buffering the overall effect on the food web (e.g.~sea warming and
fishery in San Jorge Gulf). Climate change has led to several
well-documented impacts on marine species regarding distributional
shifts induced by warming of marine currents (Wu et al., 2012;
Poloczanska et al., 2013; Vergés et al., 2014; Vergés et al., 2019).
Furthermore, warmer temperatures increase species metabolic rates
(Brown, Gillooly, Allen, Savage, \& West, 2004). Changes in metabolic
rates can subsequently translate into shifts in species traits (body
size, Vucic-Pestic, Ehnes, Rall, \& Brose (2011); Klein, Hill, Hinke,
Phillips, \& Watters (2018){]}, biomass (Perry et al., 2020) and
distribution (Kortsch, Primicerio, Fossheim, Dolgov, \& Aschan, 2015).
Alterations in the species body size and distributions have ripple
effects on feeding interactions, for example, it can introduce new
feeding interactions (Vergés et al., 2014; Pecuchet et al., 2020),
modify existing ones, and shorten energy pathways (Bartley et al., 2019;
O'Gorman et al., 2019), and reduce trophic efficiencies (Vucic-Pestic et
al., 2011).

In recent years, several new fish (Galván et al., 2022) and
macroinvertebrates species (Vinuesa, 2005; López-Gappa, 2022) were
registered in Patagonia, mostly in San Jorge Gulf in relation to the
southward range shift of warm-temperate species. This distributional
change is driven by the tropicalization of temperate waters caused by
sea warming (Vergés et al., 2014; Vergés et al., 2019). Because of its
location in the ecotone between two biogeographic provinces, the
Argentine (30°S - 44°S) and the Magellanic (43°S - 55°S) (Balech \&
Ehrlich, 2008), the San Jorge Gulf is prone to changes in species
composition. We hypothesize that sea warming will alter the food web
structure topologically, by increasing the number of species and
interactions. Newcomers are, in general, mid-trophic level species with
generalist diets, hence an increase in food web connectance may be
expected (Bartley et al., 2019). In another temperate ecosystem, an
increase in the number of fish species, led to an increase in functional
diversity and predation rate (Sgarlatta, 2023); consequences that may
also be expected in San Jorge Gulf. Given the short path length of the
San Jorge Gulf food web, the disturbances from the listed stressors are
expected to spread to many species of the food web. However, it has to
be acknowledged that the increase in functional diversity driven by the
range expansion of warm-temperate species is contrary to the process of
homogenization and loss of functional diversity in the area driven by
trawl fisheries (Rincón-Díaz et al., 2021).

In the middle of the latitudinal gradient (considered in this study),
the Scotia Sea has experienced one of the largest levels of sea warming
of any polar region (Whitehouse et al., 2008; Atkinson et al., 2019).
López-López et al. (2022) suggested that the southward distributional
change of generalist predators from the northern towards southern Scotia
Sea increases network connectance of the latter, while decreasing its
modularity. The lower modularity may increase the probability of
perturbations spreading through the network (Stouffer \& Bascompte,
2011), which may be offset by increased connectance enhancing robustness
to species loss (Dunne et al., 2002a). In the northern Scotia Sea around
South Georgia Island, we suggest that the declining krill biomass driven
by sea warming (Atkinson et al., 2019), and ocean acidification and
pollution synergy (Rowlands et al., 2021), will reduce the energy
transfer to top predators like seabirds and marine mammals. However,
this may be buffered since the dominant copepod species have maintained
its distribution (Tarling et al., 2018), but most importantly, showed an
abundance increase in recent decades likely due to reduced predation and
competition for food (Ward, Tarling, \& Thorpe, 2018). All this is
significant for the structure of the food web given the central role of
krill and copepods, characterized by high degree and mid-trophic
position (Table S1). Overall, the food web's inherent resilience, marked
by high connectance and omnivory, added to the potential compensation
for the krill decrease due to a copepod increase, may buffer against
structural changes.

In Potter Cove, a fjord-like Antarctic ecosystem, the impacts of climate
change affect many species within the food web. Potter Cove has recently
experienced frequent events of marine heatwaves, i.e.~prolonged periods
of anomalously high sea surface temperatures (Oliver et al., 2018;
Latorre et al., 2023). This has led to decreases in biomasses of
different planktonic functional groups (Garcia et al., 2019; Latorre et
al., 2023). Given the relatively low abundance of phytoplankton,
zooplankton's low degree, and the modular configuration of the food web,
we hypothesize that changes in these nodes, due to increased sea
temperatures, will be retained in the basal pelagic compartment of the
food web and will not expand to higher trophic levels. Benthic primary
producers in Potter Cove are being influence by the decrease in winter
sea ice cover (higher light availability), the increased levels of
sediments in the water column due to glacial melt run-off (lower light
penetration) and the newly free-ice areas available for colonization
associated to glacier retreat. The overall local effect of climate
change on macroalgae is a net increase in their production (Dolores
Deregibus et al., 2023). It has been proposed that larger diversity in
primary producers can support a more diverse food web with more
specialized consumers (Iken et al., 2023). We expect to see differential
effects of climate change on hard and soft bottom associated food webs
(Cordone et al., 2020). Given the high relative abundance and the high
degree of the macroalgae functional group (Table S1), we expect a longer
hard bottom food web, wider consumer trophic niches, and that it will
become more stable as sea ice cover decreases and the glacier retreats
due to global warming. In soft bottom areas of the Cove, multiple food
web nodes are being affected by ongoing warming effects: decrease in net
primary production of benthic microalgae (Hoffmann et al., 2019), and
changes in the benthic community biomass, distribution and composition
(Pasotti et al., 2015; Sahade et al., 2015). Given that the Potter Cove
food web's present low connectance and omnivory, we suggest fragility
and potential trophic cascade effects (Marina et al., 2018) with
pronounced changes in energy fluxes.

The southernmost location of the latitudinal gradient is the Weddell
Sea, where the main effect of global warming is the decrease in sea ice
extent, with reported anomalies in the past summer seasons (Fretwell et
al., 2023). Declining sea ice extent has reduced the abundance of krill
(Atkinson et al., 2004; Flores et al., 2012), and produced an increase
in phytoplankton productivity (Pinkerton et al., 2021; Isla, 2023),
altering the plankton community structure, and benefiting cryptophytes
over diatoms (Lin et al., 2021). Moreover, habitat loss from sea ice
decline will reduce the foraging success and breeding sites of seabirds
(e.g., snow petrel Pagodroma nivea and emperor penguin Aptenodytes
forsteri), decreasing their biomassess and modifying their
distributions. The projected rise in iceberg scouring is expected to
significantly alter the biomass and community structure of macrobenthos,
which in turn will impact mid-trophic level predators such as demersal
fish (Julian Gutt, 2001; Mintenbeck et al., 2012). While we do not
anticipate large-scale topological changes in the food web, local
extinctions could lead to such changes, particularly affecting benthic
species (Julian Gutt \& Piepenburg, 2003). Given that the impacted
species---whether individually like krill, or collectively like
macrobenthos---present a mid-trophic position, high biomass and high
degree (Table S1), we hypothesize that significant shifts in energy
fluxes will occur. Additionally, the food web's low proportion of
omnivores suggests reduced system resilience (Table 1), increasing the
likelihood of regime changes (Julian Gutt et al., 2015).

The two ecosystems in the subantarctic region, the Beagle Channel and
Burdwood Bank, are more affected by other anthropogenic stressors than
warming. Although these areas might be impacted by global warming
effects due to studies at a regional scale (Franco et al., 2020), to
date no studies exist for Beagle Channel and Burdwood Bank ecosystems.
In the Beagle Channel, the introduction of chinook salmon, a non-native
species, poses a significant risk to the existing food web (Fernández et
al., 2010). We hypothesize that chinook salmon's predation on Fuegian
sprat and black southern cod will disrupt the established patterns of
interaction within the food web. Both of these prey species are crucial
due to their mid-trophic positions and relatively high abundance (Table
S1). Moreover, we expect that changes in the black southern cod
population will have a more significant impact on the food web than
changes in the Fuegian sprat population, as the black southern cod has a
higher degree (Table S1). Overall, these disruptions could have
far-reaching effects on the ecosystem. This is particularly concerning
given the short path length of the food web, which means that changes
can quickly propagate through the system, affecting many species and
potentially destabilizing the entire network. This phenomenon is further
heightened by the ecosystem's inherent vulnerability to changes at
mid-trophic levels, often referred to as wasp-waist control. In the
Burdwood Bank region, fishing activities may be the main stressor
causing shifts in the food web (Table 2). We hypothesize that a
combination of factors will destabilize the already fragile ecosystem,
characterized by low connectance and low omnivory (Table 1). These
factors include a decline in the biomass of the Patagonian toothfish -a
key, highly-connected species (Table S1)- as well as smaller changes in
the biomass of four mid-level fish species, five top-level seabird
species, and over 30 types of benthic macroinvertebrates due to bycatch
(Gaitán \& Marí, 2016; Patricia A. Martínez et al., 2022; Tamini et al.,
2023). Additionally, alterations in the diets of six seabird species,
caused by discarded catch (Tamini et al., 2023), are expected to disrupt
the energy flow and further reduce the stability of the food web.

\hypertarget{gaps-and-future-perspectives}{%
\subsection{5. Gaps and future
perspectives}\label{gaps-and-future-perspectives}}

Since this review deals with qualitative data regarding, not only
prey-predator interactions but also the stressor effect, adding
quantitative data to the food webs (i.e.~interaction strength
estimations) and/or to the stressor effect (i.e.~magnitude) will provide
a better comprehension of how species' stressors might be translated
into food web effects. In this context, quantitative food webs will be
useful if they reflect energy fluxes among species (Nilsson \& McCann,
2016). Emerging methods have been proposed in this sense promising ways
to estimate shifts in species interactions within food webs (Gauzens et
al., 2019; Gellner, McCann, \& Hastings, 2023). These shifts, in turn,
could lead to changes in energy fluxes within the ecosystem.

As summarized in Table 2, in the selected study areas of the southwest
Atlantic - Antarctic latitudinal gradient several stressors might be
directly affecting consumer's diets, triggered by modified environmental
conditions (sea warming, reduced sea ice extent), new habitats (due to
distributional shifts) and/or new species (alien species). Moreover, the
population trend of important species is also changing (e.g., Funes
(2020), Constable et al. (2014), Pineda-Metz (2019), Hindell et al.
(2020), Woods et al. (2023)) driving changes in their roles as either
predators or prey (e.g. Belleggia et al. (2017), Pasti et al. (2021)).
These diet and biomass updates, either considered as a short or
long-term effect, should be contemplated in order to generate reliable
predictions of food web responses to multiple stressors in the southwest
Atlantic - Antarctic region. Examples of these interactions between
diet, relative biomass, and their effects on food webs at different
temporal scales include biological invasions and trawl fisheries. Both
stressors, in the short term, increase the complexity of food webs by
adding generalist predators and/or new prey (e.g., Cordone, Galván, \&
Momo (2023)) or by enabling discard consumption (e.g., Funes et al.
(2022)). However, in the long term, both stressors may lead to the
biological extinction of certain prey and competitors (e.g., Anton et
al. (2019)) or a significant reduction in target and incidental catch
species (e.g., Dulvy et al. (2014)), thereby promoting food web
simplification.

Regarding data gaps, we should mention the cases for Beagle Channel and
Burdwood Bank ecosystems, where almost no information exists for the
impact of global warming effects (sea warming, glacial retreat, ocean
acidification) in natural populations/communities, though warming
effects have been shown at a regional scale (Franco et al., 2020). It is
worth mentioning that in Beagle Channel recent experimental studies have
tested the tolerance of fishes to scenarios of sea warming and/or
acidification suggesting potential metabolic impacts (María Eugenia
Lattuca, Boy, Vanella, Barrantes, \& Fernández, 2018; María E. Lattuca
et al., 2023).

Analyzing the impact of multiple stressors through observational studies
can be challenging (Julian Gutt et al., 2021). This complexity arises
from the potential for antagonistic effects, where impacts cancel each
other out, or synergistic effects, where the combined impact is greater
than the sum of individual effects (Crain et al., 2008; Boyd, Lennartz,
Glover, \& Doney, 2015; Côté et al., 2016). These effects are further
complicated to handle in the framework of complex food webs. The number
of pathways of effect through which a species may affect or be affected
by other species, and through which stressors may permeate communities,
increases exponentially with the number of species and interactions in a
network (Menge, 1995). In this sense, Beauchesne, Cazelles, Archambault,
Dee, \& Gravel (2021) have developed a theory-grounded approach using
motifs (i.e.~groups of species that, when put together, construct whole
food webs) to simplify food webs; methodology that could be applied to
our food web study cases.

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

In the light of cumulative knowledge on how global warming and emerging
stressors are affecting species inhabiting marine ecosystems in the
Southern Hemisphere (e.g. Julian Gutt et al. (2021), Galván et al.
(2022)), there is a need to better understand these changes in a more
real framework where species interact and, consequently, might spread
the stressor effect to other species in complex ways. To this end, we
have reviewed the main environmental and anthropogenic stressors acting
in six different areas along a large-scale latitudinal gradient, from
temperate to Antarctic ecosystems. Using a theoretical framework that
combines node (species) and network (food web) level data, we show how
warming effects might affect food web structure and functioning. These
qualitative predictions are intended to serve as the basis to studies
that quantify how much the ongoing stressors are affecting food web
properties, such as energy fluxes and stability, in marine ecosystems of
the Southern Hemisphere.

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Acha2004}{}}%
Acha, E. M., Mianzan, H. W., Guerrero, R. A., Favero, M., \& Bava, J.
(2004). Marine fronts at the continental shelves of austral {South
America}: {Physical} and ecological processes. \emph{Journal of Marine
Systems}, \emph{44}(1), 83--105.
doi:\href{https://doi.org/10.1016/j.jmarsys.2003.09.005}{10.1016/j.jmarsys.2003.09.005}

\leavevmode\vadjust pre{\hypertarget{ref-APN2022}{}}%
Administración de Parques Nacionales. (2022). {Plan de gestión AMP
Namuncurá Banco Burdwood}. {Dirección Nacional de Áreas Marinas
Protegidas (DNAMP), Argentina}.

\leavevmode\vadjust pre{\hypertarget{ref-Albert2002}{}}%
Albert, R., \& Barabási, A.-L. (2002). Statistical mechanics of complex
networks. \emph{Reviews of Modern Physics}, \emph{74}(1), 47--97.
doi:\href{https://doi.org/10.1103/RevModPhys.74.47}{10.1103/RevModPhys.74.47}

\leavevmode\vadjust pre{\hypertarget{ref-Allega2019}{}}%
Allega, Lucrecia, Braverman, M. S., Cabreira, A. G., Campodónico, S.,
Colonello, J. H., Derisio, C. M., \ldots{} Verón, E. M. (2019).
\emph{{Estado del conocimiento biológico pesquero de los principales
recursos vivos y su ambiente, con relación a la exploración
hidrocarburífera en la Zona Económica Exclusiva Argentina y
adyacencias}}. {INIDEP}.

\leavevmode\vadjust pre{\hypertarget{ref-Allega2020}{}}%
Allega, L., Braverman, M., Campodónico, S., Carozza, C. R., Cepeda, G.
D., Colonello, J. H., \ldots{} Verón, E. (2020). \emph{{Estado del
conocimiento biológico pesquero de los principales recursos vivos y su
ambiente, con relación a la exploración hidrocarburífera en la Zona
Económica Exclusiva argentina y sus adyacencias}}. {Mar del Plata}:
{Instituto Nacional de Investigación y Desarrollo Pesquero (INIDEP)}.

\leavevmode\vadjust pre{\hypertarget{ref-Anton2019}{}}%
Anton, A., Geraldi, N. R., Lovelock, C. E., Apostolaki, E. T., Bennett,
S., Cebrian, J., \ldots{} Duarte, C. M. (2019). Global ecological
impacts of marine exotic species. \emph{Nature Ecology \& Evolution},
\emph{3}(5), 787--800.
doi:\href{https://doi.org/10.1038/s41559-019-0851-0}{10.1038/s41559-019-0851-0}

\leavevmode\vadjust pre{\hypertarget{ref-Antoni2020}{}}%
Antoni, J. S., Almandoz, G. O., Ferrario, M. E., Hernando, M. P.,
Varela, D. E., Rozema, P. D., \ldots{} Schloss, I. R. (2020). Response
of a natural {Antarctic} phytoplankton assemblage to changes in
temperature and salinity. \emph{Journal of Experimental Marine Biology
and Ecology}, \emph{532}, 151444.
doi:\href{https://doi.org/10.1016/j.jembe.2020.151444}{10.1016/j.jembe.2020.151444}

\leavevmode\vadjust pre{\hypertarget{ref-Atkinson2019}{}}%
Atkinson, A., Hill, S. L., Pakhomov, E. A., Siegel, V., Reiss, C. S.,
Loeb, V. J., \ldots{} Sailley, S. F. (2019). Krill ({Euphausia} superba)
distribution contracts southward during rapid regional warming.
\emph{Nature Climate Change}, \emph{9}(2), 142--147.
doi:\href{https://doi.org/10.1038/s41558-018-0370-z}{10.1038/s41558-018-0370-z}

\leavevmode\vadjust pre{\hypertarget{ref-Atkinson2004}{}}%
Atkinson, A., Siegel, V., Pakhomov, E., \& Rothery, P. (2004). Long-term
decline in krill stock and increase in salps within the {Southern
Ocean}. \emph{Nature}, \emph{432}(7013), 100--103.
doi:\href{https://doi.org/10.1038/nature02996}{10.1038/nature02996}

\leavevmode\vadjust pre{\hypertarget{ref-Baiser2012}{}}%
Baiser, B., Gotelli, N. J., Buckley, H. L., Miller, T. E., \& Ellison,
A. M. (2012). Geographic variation in network structure of a nearctic
aquatic food web. \emph{Global Ecology and Biogeography}, \emph{21}(5),
579--591.
doi:\href{https://doi.org/10.1111/j.1466-8238.2011.00705.x}{10.1111/j.1466-8238.2011.00705.x}

\leavevmode\vadjust pre{\hypertarget{ref-Balech2008}{}}%
Balech, E., \& Ehrlich, M. D. (2008). {Esquema biogeográfico del Mar
Argentino}.

\leavevmode\vadjust pre{\hypertarget{ref-Ballerini2014}{}}%
Ballerini, T., Hofmann, E. E., Ainley, D. G., Daly, K., Marrari, M.,
Ribic, C. A., \ldots{} Steele, J. H. (2014). Productivity and linkages
of the food web of the southern region of the western {Antarctic
Peninsula} continental shelf. \emph{Progress in Oceanography},
\emph{122}, 10--29.
doi:\href{https://doi.org/10.1016/j.pocean.2013.11.007}{10.1016/j.pocean.2013.11.007}

\leavevmode\vadjust pre{\hypertarget{ref-Bartley2019}{}}%
Bartley, T. J., McCann, K. S., Bieg, C., Cazelles, K., Granados, M.,
Guzzo, M. M., \ldots{} McMeans, B. C. (2019). Food web rewiring in a
changing world. \emph{Nature Ecology \& Evolution}, \emph{3}(3),
345--354.
doi:\href{https://doi.org/10.1038/s41559-018-0772-3}{10.1038/s41559-018-0772-3}

\leavevmode\vadjust pre{\hypertarget{ref-Bauer2022}{}}%
Bauer, B., Berti, E., Ryser, R., Gauzens, B., Hirt, M. R., Rosenbaum,
B., \ldots{} Brose, U. (2022). Biotic filtering by species' interactions
constrains food-web variability across spatial and abiotic gradients.
\emph{Ecology Letters}, \emph{25}(5), 1225--1236.
doi:\href{https://doi.org/10.1111/ele.13995}{10.1111/ele.13995}

\leavevmode\vadjust pre{\hypertarget{ref-Beauchesne2021}{}}%
Beauchesne, D., Cazelles, K., Archambault, P., Dee, L. E., \& Gravel, D.
(2021). On the sensitivity of food webs to multiple stressors.
\emph{Ecology Letters}, \emph{24}(10), 2219--2237.
doi:\href{https://doi.org/10.1111/ele.13841}{10.1111/ele.13841}

\leavevmode\vadjust pre{\hypertarget{ref-Belgrano2005}{}}%
Belgrano, A., Scharler, S. E. R. C. U. M., Scharler, U. M., Dunne, J.,
\& Ulanowicz, R. E. (2005). \emph{Aquatic {Food Webs}: {An Ecosystem
Approach}}. {OUP Oxford}.

\leavevmode\vadjust pre{\hypertarget{ref-Belleggia2017}{}}%
Belleggia, M., Giberto, D., \& Bremec, C. (2017). Adaptation of diet in
a changed environment: {Increased} consumption of lobster krill {Munida}
gregaria ({Fabricius}, 1793) by {Argentine} hake. \emph{Marine Ecology},
\emph{38}(4), e12445.
doi:\href{https://doi.org/10.1111/maec.12445}{10.1111/maec.12445}

\leavevmode\vadjust pre{\hypertarget{ref-Bers2013}{}}%
Bers, A. V., Momo, F., Schloss, I. R., \& Abele, D. (2013). Analysis of
trends and sudden changes in long-term environmental data from {King
George Island} ({Antarctica}): Relationships between global climatic
oscillations and local system response. \emph{Climatic Change},
\emph{116}(3), 789--803.
doi:\href{https://doi.org/10.1007/s10584-012-0523-4}{10.1007/s10584-012-0523-4}

\leavevmode\vadjust pre{\hypertarget{ref-Bogazzi2005}{}}%
Bogazzi, E., Baldoni, A., Rivas, A., Martos, P., Reta, R., Orensanz, J.
M. (lobo)., \ldots{} Werner, F. (2005). Spatial correspondence between
areas of concentration of {Patagonian} scallop ({Zygochlamys}
patagonica) and frontal systems in the southwestern {Atlantic}.
\emph{Fisheries Oceanography}, \emph{14}(5), 359--376.
doi:\href{https://doi.org/10.1111/j.1365-2419.2005.00340.x}{10.1111/j.1365-2419.2005.00340.x}

\leavevmode\vadjust pre{\hypertarget{ref-Bovcon2013}{}}%
Bovcon, N. D., Góngora, M. E., Marinao, C., \& González-Zevallos, D.
(2013). Composición de las capturas y descartes generados en la pesca de
merluza común {Merluccius} hubbsi y langostino patagónico {Pleoticus}
muelleri: Un caso de estudio en la flota fresquera de altura del {Golfo
San Jorge}, {Chubut}, {Argentina}. \emph{Revista de Biología Marina y
Oceanografía}, \emph{48}(2), 303--319.
doi:\href{https://doi.org/10.4067/S0718-19572013000200010}{10.4067/S0718-19572013000200010}

\leavevmode\vadjust pre{\hypertarget{ref-Boyd2015}{}}%
Boyd, P. W., Lennartz, S. T., Glover, D. M., \& Doney, S. C. (2015).
Biological ramifications of climate-change-mediated oceanic
multi-stressors. \emph{Nature Climate Change}, \emph{5}(1), 71--79.
doi:\href{https://doi.org/10.1038/nclimate2441}{10.1038/nclimate2441}

\leavevmode\vadjust pre{\hypertarget{ref-Braeckman2021}{}}%
Braeckman, U., Pasotti, F., Hoffmann, R., Vázquez, S., Wulff, A.,
Schloss, I. R., \ldots{} Vanreusel, A. (2021). Glacial melt disturbance
shifts community metabolism of an {Antarctic} seafloor ecosystem from
net autotrophy to heterotrophy. \emph{Communications Biology},
\emph{4}(1), 148.
doi:\href{https://doi.org/10.1038/s42003-021-01673-6}{10.1038/s42003-021-01673-6}

\leavevmode\vadjust pre{\hypertarget{ref-Braithwaite2015}{}}%
Braithwaite, J. E., Meeuwig, J. J., Letessier, T. B., Jenner, K. C. S.,
\& Brierley, A. S. (2015). From sea ice to blubber: Linking whale
condition to krill abundance using historical whaling records.
\emph{Polar Biology}, \emph{38}(8), 1195--1202.
doi:\href{https://doi.org/10.1007/s00300-015-1685-0}{10.1007/s00300-015-1685-0}

\leavevmode\vadjust pre{\hypertarget{ref-Brown2004}{}}%
Brown, J. H., Gillooly, J. F., Allen, A. P., Savage, V. M., \& West, G.
B. (2004). Toward a {Metabolic Theory} of {Ecology}. \emph{Ecology},
\emph{85}(7), 1771--1789.
doi:\href{https://doi.org/10.1890/03-9000}{10.1890/03-9000}

\leavevmode\vadjust pre{\hypertarget{ref-Bruder2019}{}}%
Bruder, A., Frainer, A., Rota, T., \& Primicerio, R. (2019). The
{Importance} of {Ecological Networks} in {Multiple-Stressor Research}
and {Management}. \emph{Frontiers in Environmental Science}, \emph{7}.

\leavevmode\vadjust pre{\hypertarget{ref-Bruno2023}{}}%
Bruno, D. O., Riccialdelli, L., Acha, E. M., \& Fernández, D. A. (2023).
Seasonal variation of autochthonous and allochthonous carbon sources for
the first levels of the {Beagle Channel} food web. \emph{Journal of
Marine Systems}, \emph{239}, 103859.
doi:\href{https://doi.org/10.1016/j.jmarsys.2023.103859}{10.1016/j.jmarsys.2023.103859}

\leavevmode\vadjust pre{\hypertarget{ref-Buskey2016}{}}%
Buskey, E. J., White, H. K., \& Esbaugh, A. J. (2016). Impact of {Oil
Spills} on {Marine Life} in the {Gulf} of {Mexico}: {EFFECTS ON
PLANKTON}, {NEKTON}, {AND DEEP-SEA BENTHOS}. \emph{Oceanography},
\emph{29}(3), 174--181. Retrieved from
\url{https://www.jstor.org/stable/24862719}

\leavevmode\vadjust pre{\hypertarget{ref-Campana2018}{}}%
Campana, G. L., Zacher, K., Deregibus, D., Momo, F. R., Wiencke, C., \&
Quartino, M. L. (2018). Succession of {Antarctic} benthic algae ({Potter
Cove}, {South Shetland Islands}): Structural patterns and glacial impact
over a four-year period. \emph{Polar Biology}, \emph{41}(2), 377--396.
doi:\href{https://doi.org/10.1007/s00300-017-2197-x}{10.1007/s00300-017-2197-x}

\leavevmode\vadjust pre{\hypertarget{ref-Chown2022}{}}%
Chown, S. L., Leihy, R. I., Naish, T. R., Brooks, C. M., Convey, P.,
Henley, B. J., \ldots{} Grant, S. M. (2022). Antarctic climate change
and the environment: A decadal synopsis and recommendations for action.

\leavevmode\vadjust pre{\hypertarget{ref-Ciancio2008}{}}%
Ciancio, J. E., Pascual, M. A., Botto, F., Frere, E., \& Iribarne, O.
(2008). Trophic relationships of exotic anadromous salmonids in the
southern {Patagonian Shelf} as inferred from stable isotopes.
\emph{Limnology and Oceanography}, \emph{53}(2), 788--798.
doi:\href{https://doi.org/10.4319/lo.2008.53.2.0788}{10.4319/lo.2008.53.2.0788}

\leavevmode\vadjust pre{\hypertarget{ref-Ciancio2010}{}}%
Ciancio, J., Beauchamp, D. A., \& Pascual, M. (2010). Marine effect of
introduced salmonids: {Prey} consumption by exotic steelhead and
anadromous brown trout in the {Patagonian Continental Shelf}.
\emph{Limnology and Oceanography}, \emph{55}(5), 2181--2192.
doi:\href{https://doi.org/10.4319/lo.2010.55.5.2181}{10.4319/lo.2010.55.5.2181}

\leavevmode\vadjust pre{\hypertarget{ref-Cirtwill2015}{}}%
Cirtwill, A. R., Stouffer, D. B., \& Romanuk, T. N. (2015). Latitudinal
gradients in biotic niche breadth vary across ecosystem types.
\emph{Proceedings of the Royal Society B: Biological Sciences},
\emph{282}(1819), 20151589.
doi:\href{https://doi.org/10.1098/rspb.2015.1589}{10.1098/rspb.2015.1589}

\leavevmode\vadjust pre{\hypertarget{ref-Clark2013}{}}%
Clark, G. F., Stark, J. S., Johnston, E. L., Runcie, J. W., Goldsworthy,
P. M., Raymond, B., \& Riddle, M. J. (2013). Light-driven tipping points
in polar ecosystems. \emph{Global Change Biology}, \emph{19}(12),
3749--3761.
doi:\href{https://doi.org/10.1111/gcb.12337}{10.1111/gcb.12337}

\leavevmode\vadjust pre{\hypertarget{ref-Collins2010}{}}%
Collins, M. A., Brickle, P., Brown, J., \& Belchier, M. (2010). Chapter
{Four} - {The Patagonian Toothfish}: {Biology}, {Ecology} and {Fishery}.
In M. Lesser (Ed.), \emph{Advances in {Marine Biology}} (Vol. 58, pp.
227--300). {Academic Press}.
doi:\href{https://doi.org/10.1016/B978-0-12-381015-1.00004-6}{10.1016/B978-0-12-381015-1.00004-6}

\leavevmode\vadjust pre{\hypertarget{ref-Commendatore2012}{}}%
Commendatore, M. G., Nievas, M. L., Amin, O., \& Esteves, J. L. (2012).
Sources and distribution of aliphatic and polyaromatic hydrocarbons in
coastal sediments from the {Ushuaia Bay} ({Tierra} del {Fuego},
{Patagonia}, {Argentina}). \emph{Marine Environmental Research},
\emph{74}, 20--31.
doi:\href{https://doi.org/10.1016/j.marenvres.2011.11.010}{10.1016/j.marenvres.2011.11.010}

\leavevmode\vadjust pre{\hypertarget{ref-Constable2014}{}}%
Constable, A. J., Melbourne-Thomas, J., Corney, S. P., Arrigo, K. R.,
Barbraud, C., Barnes, D. K. A., \ldots{} Ziegler, P. (2014). Climate
change and {Southern Ocean} ecosystems {I}: How changes in physical
habitats directly affect marine biota. \emph{Global Change Biology},
\emph{20}(10), 3004--3025.
doi:\href{https://doi.org/10.1111/gcb.12623}{10.1111/gcb.12623}

\leavevmode\vadjust pre{\hypertarget{ref-Cordone2023}{}}%
Cordone, G., Galván, D. E., \& Momo, F. R. (2023). Impacts of an
invasion by green crab {Carcinus} maenas on the intertidal food web of a
{Patagonian} rocky shore, {Argentina}. \emph{Marine Ecology Progress
Series}, \emph{713}, 97--115.
doi:\href{https://doi.org/10.3354/meps14336}{10.3354/meps14336}

\leavevmode\vadjust pre{\hypertarget{ref-Cordone2020}{}}%
Cordone, G., Salinas, V., Marina, T. I., Doyle, S. R., Pasotti, F.,
Saravia, L. A., \& Momo, F. R. (2020). Green vs brown food web:
{Effects} of habitat type on multidimensional stability proxies for a
highly-resolved {Antarctic} food web. \emph{Food Webs}, \emph{25},
e00166.
doi:\href{https://doi.org/10.1016/j.fooweb.2020.e00166}{10.1016/j.fooweb.2020.e00166}

\leavevmode\vadjust pre{\hypertarget{ref-Correa2008}{}}%
Correa, C., \& Gross, M. R. (2008). Chinook salmon invade southern
{South America}. \emph{Biological Invasions}, \emph{10}(5), 615--639.
doi:\href{https://doi.org/10.1007/s10530-007-9157-2}{10.1007/s10530-007-9157-2}

\leavevmode\vadjust pre{\hypertarget{ref-Cossi2021}{}}%
Cossi, P. F., Ojeda, M., Chiesa, I. L., Rimondino, G. N., Fraysse, C.,
Calcagno, J., \& Pérez, A. F. (2021). First evidence of microplastics in
the {Marine Protected Area Namuncurá} at {Burdwood Bank}, {Argentina}: A
study on {Henricia} obesa and {Odontaster} penicillatus
({Echinodermata}: {Asteroidea}). \emph{Polar Biology}, \emph{44}(12),
2277--2287.
doi:\href{https://doi.org/10.1007/s00300-021-02959-5}{10.1007/s00300-021-02959-5}

\leavevmode\vadjust pre{\hypertarget{ref-Cote2016}{}}%
Côté, I. M., Darling, E. S., \& Brown, C. J. (2016). Interactions among
ecosystem stressors and their importance in conservation.
\emph{Proceedings of the Royal Society B: Biological Sciences},
\emph{283}(1824), 20152592.
doi:\href{https://doi.org/10.1098/rspb.2015.2592}{10.1098/rspb.2015.2592}

\leavevmode\vadjust pre{\hypertarget{ref-Crain2008}{}}%
Crain, C. M., Kroeker, K., \& Halpern, B. S. (2008). Interactive and
cumulative effects of multiple human stressors in marine systems.
\emph{Ecology Letters}, \emph{11}(12), 1304--1315.
doi:\href{https://doi.org/10.1111/j.1461-0248.2008.01253.x}{10.1111/j.1461-0248.2008.01253.x}

\leavevmode\vadjust pre{\hypertarget{ref-DAlelio2016}{}}%
D'Alelio, D., Libralato, S., Wyatt, T., \& Ribera d'Alcalà, M. (2016).
Ecological-network models link diversity, structure and function in the
plankton food-web. \emph{Scientific Reports}, \emph{6}(1), 21806.
doi:\href{https://doi.org/10.1038/srep21806}{10.1038/srep21806}

\leavevmode\vadjust pre{\hypertarget{ref-Dayton1990}{}}%
Dayton, P. K. (1990). Polar {Benthos}. In \emph{Polar {Oceanography Part
B}: {Chemistry}, {Biology} and {Geology}} (W.O. Smith, Jr., pp.
631--685). {London}: {Academic Press}.

\leavevmode\vadjust pre{\hypertarget{ref-deHaro2022}{}}%
de Haro, J. C., Perez Orsi, H., Cané, S., Di Pangracio, A., Falabella,
V., \& Sapoznikow, A. (2022). {Informe colaborativo sobre el Estado de
situación. Riesgos e impactos de la prospección sísmica en el Mar
Argentino}. {Foro para la Conservación del Mar Patagónico y Áreas de
Influencia}.

\leavevmode\vadjust pre{\hypertarget{ref-Deppeler2020}{}}%
Deppeler, S., Schulz, K. G., Hancock, A., Pascoe, P., McKinlay, J., \&
Davidson, A. (2020). Ocean acidification reduces growth and grazing
impact of {Antarctic} heterotrophic nanoflagellates.
\emph{Biogeosciences}, \emph{17}(16), 4153--4171.
doi:\href{https://doi.org/10.5194/bg-17-4153-2020}{10.5194/bg-17-4153-2020}

\leavevmode\vadjust pre{\hypertarget{ref-Deregibus2023}{}}%
Deregibus, Dolores, Campana, G. L., Neder, C., Barnes, D. K. A., Zacher,
K., Piscicelli, J. M., \ldots{} Quartino, M. L. (2023). Potential
macroalgal expansion and blue carbon gains with northern {Antarctic
Peninsula} glacial retreat. \emph{Marine Environmental Research},
\emph{189}, 106056.
doi:\href{https://doi.org/10.1016/j.marenvres.2023.106056}{10.1016/j.marenvres.2023.106056}

\leavevmode\vadjust pre{\hypertarget{ref-Deregibus2016}{}}%
Deregibus, D., Quartino, M. L., Campana, G. L., Momo, F. R., Wiencke,
C., \& Zacher, K. (2016). Photosynthetic light requirements and vertical
distribution of macroalgae in newly ice-free areas in {Potter Cove},
{South Shetland Islands}, {Antarctica}. \emph{Polar Biology},
\emph{39}(1), 153--166.
doi:\href{https://doi.org/10.1007/s00300-015-1679-y}{10.1007/s00300-015-1679-y}

\leavevmode\vadjust pre{\hypertarget{ref-Deregibus2017}{}}%
Deregibus, D., Quartino, M. L., Zacher, K., Campana, G. L., \& Barnes,
D. K. A. (2017). Understanding the link between sea ice, ice scour and
{Antarctic} benthic biodiversity\textendash the need for cross-station
and international collaboration. \emph{Polar Record}, \emph{53}(2),
143--152.
doi:\href{https://doi.org/10.1017/S0032247416000875}{10.1017/S0032247416000875}

\leavevmode\vadjust pre{\hypertarget{ref-Deregibus2020}{}}%
Deregibus, Dolores, Zacher, K., Bartsch, I., Campana, G. L., Momo, F.
R., Wiencke, C., \ldots{} Quartino, M. L. (2020). Carbon {Balance Under}
a {Changing Light Environment}. In I. Gómez \& P. Huovinen (Eds.),
\emph{Antarctic {Seaweeds}: {Diversity}, {Adaptation} and {Ecosystem
Services}} (pp. 173--191). {Cham}: {Springer International Publishing}.
doi:\href{https://doi.org/10.1007/978-3-030-39448-6_9}{10.1007/978-3-030-39448-6\_9}

\leavevmode\vadjust pre{\hypertarget{ref-DiMauro2022}{}}%
Di Mauro, R., Castillo, S., Pérez, A., Iachetti, C. M., Silva, L.,
Tomba, J. P., \& Chiesa, I. L. (2022). Anthropogenic microfibers are
highly abundant at the {Burdwood Bank} seamount, a protected
sub-{Antarctic} environment in the {Southwestern Atlantic Ocean}.
\emph{Environmental Pollution}, \emph{306}, 119364.
doi:\href{https://doi.org/10.1016/j.envpol.2022.119364}{10.1016/j.envpol.2022.119364}

\leavevmode\vadjust pre{\hypertarget{ref-Dinniman2023}{}}%
Dinniman, M. S., St-Laurent, P., Arrigo, K. R., Hofmann, E. E., \& van
Dijken, G. L. (2023). Sensitivity of the {Relationship Between Antarctic
Ice Shelves} and {Iron Supply} to {Projected Changes} in the
{Atmospheric Forcing}. \emph{Journal of Geophysical Research: Oceans},
\emph{128}(2), e2022JC019210.
doi:\href{https://doi.org/10.1029/2022JC019210}{10.1029/2022JC019210}

\leavevmode\vadjust pre{\hypertarget{ref-Dodino2022}{}}%
Dodino, S., Riccialdelli, L., Polito, M. J., Pütz, K., Brasso, R. L., \&
Raya Rey, A. (2022). Mercury exposure driven by geographic and trophic
factors in {Magellanic} penguins from {Tierra} del {Fuego}. \emph{Marine
Pollution Bulletin}, \emph{174}, 113184.
doi:\href{https://doi.org/10.1016/j.marpolbul.2021.113184}{10.1016/j.marpolbul.2021.113184}

\leavevmode\vadjust pre{\hypertarget{ref-Duarte2011}{}}%
Duarte, C. A., Giarratano, E., Amin, O. A., \& Comoglio, L. I. (2011).
Heavy metal concentrations and biomarkers of oxidative stress in native
mussels ({Mytilus} edulis chilensis) from {Beagle Channel} coast
({Tierra} del {Fuego}, {Argentina}). \emph{Marine Pollution Bulletin},
\emph{62}(8), 1895--1904.
doi:\href{https://doi.org/10.1016/j.marpolbul.2011.05.031}{10.1016/j.marpolbul.2011.05.031}

\leavevmode\vadjust pre{\hypertarget{ref-Duffy2007}{}}%
Duffy, J. E., Cardinale, B. J., France, K. E., McIntyre, P. B.,
Thébault, E., \& Loreau, M. (2007). The functional role of biodiversity
in ecosystems: Incorporating trophic complexity. \emph{Ecology Letters},
\emph{10}(6), 522--538.
doi:\href{https://doi.org/10.1111/j.1461-0248.2007.01037.x}{10.1111/j.1461-0248.2007.01037.x}

\leavevmode\vadjust pre{\hypertarget{ref-Dulvy2014}{}}%
Dulvy, N. K., Fowler, S. L., Musick, J. A., Cavanagh, R. D., Kyne, P.
M., Harrison, L. R., \ldots{} White, W. T. (2014). Extinction risk and
conservation of the world's sharks and rays. \emph{eLife}, \emph{3},
e00590.
doi:\href{https://doi.org/10.7554/eLife.00590}{10.7554/eLife.00590}

\leavevmode\vadjust pre{\hypertarget{ref-Dunne2002a}{}}%
Dunne, J. A., Williams, R. J., \& Martinez, N. D. (2002a). Food-web
structure and network theory: {The} role of connectance and size.
\emph{Proceedings of the National Academy of Sciences}, \emph{99}(20),
12917--12922.
doi:\href{https://doi.org/10.1073/pnas.192407699}{10.1073/pnas.192407699}

\leavevmode\vadjust pre{\hypertarget{ref-Dunne2002}{}}%
Dunne, J. A., Williams, R. J., \& Martinez, N. D. (2002b). Network
structure and biodiversity loss in food webs: Robustness increases with
connectance. \emph{Ecology Letters}, \emph{5}(4), 558--567.
doi:\href{https://doi.org/10.1046/j.1461-0248.2002.00354.x}{10.1046/j.1461-0248.2002.00354.x}

\leavevmode\vadjust pre{\hypertarget{ref-Fernandez2010}{}}%
Fernández, D. A., Ciancio, J., Ceballos, S. G., Riva-Rossi, C., \&
Pascual, M. A. (2010). Chinook salmon ({Oncorhynchus} tshawytscha,
{Walbaum} 1792) in the {Beagle Channel}, {Tierra} del {Fuego}: The onset
of an invasion. \emph{Biological Invasions}, \emph{12}(9), 2991--2997.
doi:\href{https://doi.org/10.1007/s10530-010-9731-x}{10.1007/s10530-010-9731-x}

\leavevmode\vadjust pre{\hypertarget{ref-Ferreira2021}{}}%
Ferreira, M. F., Lo Nostro, F. L., Fernández, D. A., \& Genovese, G.
(2021). Endocrine disruption in the sub {Antarctic} fish
{Patagonotothen} tessellata ({Perciformes}, {Notothenidae}) from {Beagle
Channel} associated to anthropogenic impact. \emph{Marine Environmental
Research}, \emph{171}, 105478.
doi:\href{https://doi.org/10.1016/j.marenvres.2021.105478}{10.1016/j.marenvres.2021.105478}

\leavevmode\vadjust pre{\hypertarget{ref-Fioramonti2022}{}}%
Fioramonti, N. E., Ribeiro Guevara, S., Becker, Y. A., \& Riccialdelli,
L. (2022). Mercury transfer in coastal and oceanic food webs from the
{Southwest Atlantic Ocean}. \emph{Marine Pollution Bulletin},
\emph{175}, 113365.
doi:\href{https://doi.org/10.1016/j.marpolbul.2022.113365}{10.1016/j.marpolbul.2022.113365}

\leavevmode\vadjust pre{\hypertarget{ref-Flores2012}{}}%
Flores, H., Atkinson, A., Kawaguchi, S., Krafft, B., Milinevsky, G.,
Nicol, S., \ldots{} Werner, T. (2012). Impact of climate change on
{Antarctic} krill. \emph{Marine Ecology Progress Series}, \emph{458},
1--19. doi:\href{https://doi.org/10.3354/meps09831}{10.3354/meps09831}

\leavevmode\vadjust pre{\hypertarget{ref-Franco2020}{}}%
Franco, B. C., Combes, V., \& González Carman, V. (2020). Subsurface
{Ocean Warming Hotspots} and {Potential Impacts} on {Marine Species}:
{The Southwest South Atlantic Ocean Case Study}. \emph{Frontiers in
Marine Science}, \emph{7}.

\leavevmode\vadjust pre{\hypertarget{ref-Frelat2022}{}}%
Frelat, R., Kortsch, S., Kröncke, I., Neumann, H., Nordström, M. C.,
Olivier, P. E. N., \& Sell, A. F. (2022). Food web structure and
community composition: A comparison across space and time in the {North
Sea}. \emph{Ecography}, \emph{2022}(2), ecog.05945.
doi:\href{https://doi.org/10.1111/ecog.05945}{10.1111/ecog.05945}

\leavevmode\vadjust pre{\hypertarget{ref-Fretwell2023}{}}%
Fretwell, P. T., Boutet, A., \& Ratcliffe, N. (2023). Record low 2022
{Antarctic} sea ice led to catastrophic breeding failure of emperor
penguins. \emph{Communications Earth \& Environment}, \emph{4}(1), 1--6.
doi:\href{https://doi.org/10.1038/s43247-023-00927-x}{10.1038/s43247-023-00927-x}

\leavevmode\vadjust pre{\hypertarget{ref-Fuentes2016}{}}%
Fuentes, V., Alurralde, G., Meyer, B., Aguirre, G. E., Canepa, A.,
Wölfl, A.-C., \ldots{} Schloss, I. R. (2016). Glacial melting: An
overlooked threat to {Antarctic} krill. \emph{Scientific Reports},
\emph{6}(1), 27234.
doi:\href{https://doi.org/10.1038/srep27234}{10.1038/srep27234}

\leavevmode\vadjust pre{\hypertarget{ref-Funes2020}{}}%
Funes, M. (2020, March). \emph{Efectos de la pesca de arrastre sobre la
estructura trófica del norte del {Golfo San Jorge}} (PhD thesis).

\leavevmode\vadjust pre{\hypertarget{ref-Funes2019}{}}%
Funes, M., Marinao, C., \& Galván, D. E. (2019). Does trawl fisheries
affect the diet of fishes? {A} stable isotope analysis approach*.
\emph{Isotopes in Environmental and Health Studies}, \emph{55}(4),
327--343.
doi:\href{https://doi.org/10.1080/10256016.2019.1626381}{10.1080/10256016.2019.1626381}

\leavevmode\vadjust pre{\hypertarget{ref-Funes2022}{}}%
Funes, M., Saravia, L. A., Cordone, G., Iribarne, O. O., \& Galván, D.
E. (2022). Network analysis suggests changes in food web stability
produced by bottom trawl fishery in {Patagonia}. \emph{Scientific
Reports}, \emph{12}(1), 10876.
doi:\href{https://doi.org/10.1038/s41598-022-14363-y}{10.1038/s41598-022-14363-y}

\leavevmode\vadjust pre{\hypertarget{ref-Gaitan2016}{}}%
Gaitán, E., \& Marí, N. (2016). {Análisis de las comunidades bentónicas
asociadas a capturas de la flota comercial dirigida a Macruronus
magellanicus}. {Instituto Nacional de Investigación y Desarrollo
Pesquero (INIDEP)}.

\leavevmode\vadjust pre{\hypertarget{ref-Galvan2022}{}}%
Galván, D. E., Bovcon, N. D., Cochia, P. D., González, R. A., Lattuca,
M. E., Reinaldo, M. O., \ldots{} Svendsen, G. M. (2022). Changes in the
{Specific} and {Biogeographic Composition} of {Coastal Fish Assemblages}
in {Patagonia}, {Driven} by {Climate Change}, {Fishing}, and {Invasion}
by {Alien Species}. In E. W. Helbling, M. A. Narvarte, R. A. González,
\& V. E. Villafañe (Eds.), \emph{Global {Change} in {Atlantic Coastal
Patagonian Ecosystems}: {A Journey Through Time}} (pp. 205--231).
{Cham}: {Springer International Publishing}.
doi:\href{https://doi.org/10.1007/978-3-030-86676-1_9}{10.1007/978-3-030-86676-1\_9}

\leavevmode\vadjust pre{\hypertarget{ref-Garcia2019}{}}%
Garcia, M. D., Fernández Severini, M. D., Spetter, C., López Abbate, M.
C., Tartara, M. N., Nahuelhual, E. G., \ldots{} Hoffmeyer, M. S. (2019).
Effects of glacier melting on the planktonic communities of two
{Antarctic} coastal areas ({Potter Cove} and {Hope Bay}) in summer.
\emph{Regional Studies in Marine Science}, \emph{30}, 100731.
doi:\href{https://doi.org/10.1016/j.rsma.2019.100731}{10.1016/j.rsma.2019.100731}

\leavevmode\vadjust pre{\hypertarget{ref-Garcia2016}{}}%
Garcia, M. D., Hoffmeyer, M. S., Abbate, M. C. L., Barría De Cao, M. S.,
Pettigrosso, R. E., Almandoz, G. O., \ldots{} Schloss, I. R. (2016).
Micro- and mesozooplankton responses during two contrasting summers in a
coastal {Antarctic} environment. \emph{Polar Biology}, \emph{39}(1),
123--137.
doi:\href{https://doi.org/10.1007/s00300-015-1678-z}{10.1007/s00300-015-1678-z}

\leavevmode\vadjust pre{\hypertarget{ref-Garcia-Borboroglu2008}{}}%
García-Borboroglu, P., Boersma, P. D., Reyes, L., \& Skewgar, E. (2008).
Petroleum {Pollution} and {Penguins}: {Marine Conservation Tools} to
{Reduce} the {Problem}. In \emph{Marine {Pollution}: {New} research}
(Hofer, T.N., pp. 339--356). {New York, USA}: {Nova Science Publishers
Inc.}

\leavevmode\vadjust pre{\hypertarget{ref-Garrott1993}{}}%
Garrott, R. A., Eberhardt, L. L., \& Burn, D. M. (1993). Mortality of
{Sea Otters} in {Prince William Sound Following} the {Exxon Valdez Oil
Spill}. \emph{Marine Mammal Science}, \emph{9}(4), 343--359.
doi:\href{https://doi.org/10.1111/j.1748-7692.1993.tb00468.x}{10.1111/j.1748-7692.1993.tb00468.x}

\leavevmode\vadjust pre{\hypertarget{ref-Gauzens2019}{}}%
Gauzens, B., Barnes, A., Giling, D. P., Hines, J., Jochum, M., Lefcheck,
J. S., \ldots{} Brose, U. (2019). Fluxweb: {An R} package to easily
estimate energy fluxes in food webs. \emph{Methods in Ecology and
Evolution}, \emph{10}(2), 270--279.
doi:\href{https://doi.org/10.1111/2041-210X.13109}{10.1111/2041-210X.13109}

\leavevmode\vadjust pre{\hypertarget{ref-Gellner2023}{}}%
Gellner, G., McCann, K., \& Hastings, A. (2023). Stable diverse food
webs become more common when interactions are more biologically
constrained. \emph{Proceedings of the National Academy of Sciences},
\emph{120}(31), e2212061120.
doi:\href{https://doi.org/10.1073/pnas.2212061120}{10.1073/pnas.2212061120}

\leavevmode\vadjust pre{\hypertarget{ref-Giarratano2010}{}}%
Giarratano, E., \& Amin, O. A. (2010). Heavy metals monitoring in the
southernmost mussel farm of the world ({Beagle Channel}, {Argentina}).
\emph{Ecotoxicology and Environmental Safety}, \emph{73}(6), 1378--1384.
doi:\href{https://doi.org/10.1016/j.ecoenv.2010.06.023}{10.1016/j.ecoenv.2010.06.023}

\leavevmode\vadjust pre{\hypertarget{ref-Gil2011}{}}%
Gil, M. N., Torres, A. I., Amin, O., \& Esteves, J. L. (2011).
Assessment of recent sediment influence in an urban polluted
subantarctic coastal ecosystem. {Beagle Channel} ({Southern Argentina}).
\emph{Marine Pollution Bulletin}, \emph{62}(1), 201--207.
doi:\href{https://doi.org/10.1016/j.marpolbul.2010.10.004}{10.1016/j.marpolbul.2010.10.004}

\leavevmode\vadjust pre{\hypertarget{ref-Glembocki2015}{}}%
Glembocki, N. G., Williams, G. N., Góngora, M. E., Gagliardini, D. A.,
\& Orensanz, J. M. (Lobo). (2015). Synoptic oceanography of {San Jorge
Gulf} ({Argentina}): {A} template for {Patagonian} red shrimp
({Pleoticus} muelleri) spatial dynamics. \emph{Journal of Sea Research},
\emph{95}, 22--35.
doi:\href{https://doi.org/10.1016/j.seares.2014.10.011}{10.1016/j.seares.2014.10.011}

\leavevmode\vadjust pre{\hypertarget{ref-Glorioso1995}{}}%
Glorioso, P. D., \& Flather, R. A. (1995). A barotropic model of the
currents off {SE South America}. \emph{Journal of Geophysical Research:
Oceans}, \emph{100}(C7), 13427--13440.
doi:\href{https://doi.org/10.1029/95JC00942}{10.1029/95JC00942}

\leavevmode\vadjust pre{\hypertarget{ref-Gongora2009}{}}%
Góngora, María E., Bovcon, N. D., \& Cochia, P. D. (2009). Ictiofauna
capturada incidentalmente en la pesquería de langostino patagónico
{Pleoticus} muelleri {Bate}, 1888. \emph{Revista de Biología Marina y
Oceanografía}, \emph{44}(3), 583--593.
doi:\href{https://doi.org/10.4067/S0718-19572009000300006}{10.4067/S0718-19572009000300006}

\leavevmode\vadjust pre{\hypertarget{ref-Gongora2012}{}}%
Góngora, María Eva, González-Zevallos, D., Pettovello, A., \& Mendía, L.
(2012). Caracterización de las principales pesquerías del golfo {San
Jorge Patagonia}, {Argentina}. \emph{Latin American Journal of Aquatic
Research}, \emph{40}(1), 1--11.

\leavevmode\vadjust pre{\hypertarget{ref-Gonzalez-Zevallos2006}{}}%
González-Zevallos, D., \& Yorio, P. (2006). Seabird use of discards and
incidental captures at the {Argentine} hake trawl fishery in the {Golfo
San Jorge}, {Argentina}. \emph{Marine Ecology Progress Series},
\emph{316}, 175--183.
doi:\href{https://doi.org/10.3354/meps316175}{10.3354/meps316175}

\leavevmode\vadjust pre{\hypertarget{ref-Gonzalez-Zevallos2011}{}}%
González-Zevallos, D., \& Yorio, P. (2011). Consumption of discards and
interactions between {Black-browed Albatrosses} ({Thalassarche}
melanophrys) and {Kelp Gulls} ({Larus} dominicanus) at trawl fisheries
in {Golfo San Jorge}, {Argentina}. \emph{Journal of Ornithology},
\emph{152}(4), 827--838.
doi:\href{https://doi.org/10.1007/s10336-011-0657-6}{10.1007/s10336-011-0657-6}

\leavevmode\vadjust pre{\hypertarget{ref-Gorini2021}{}}%
Gorini, F. L., Lukaszewicz, G., \& Giussi, A. R. (2021).
\emph{{Actualización de la estadística pesquera de peces demersales
australes en el Atlántico Sudoccidental (período 2008-2020)}} (\{Informe
T\textbackslash\textquotesingle ecnico Oficial\}) (p. 64). {Mar del
Plata}: {Instituto Nacional de Investigación y Desarrollo Pesquero
(INIDEP)}.

\leavevmode\vadjust pre{\hypertarget{ref-Guihou2020}{}}%
Guihou, K., Piola, A. R., Palma, E. D., \& Chidichimo, M. P. (2020).
Dynamical connections between large marine ecosystems of austral {South
America} based on numerical simulations. \emph{Ocean Science},
\emph{16}(2), 271--290.
doi:\href{https://doi.org/10.5194/os-16-271-2020}{10.5194/os-16-271-2020}

\leavevmode\vadjust pre{\hypertarget{ref-Gutt2001a}{}}%
Gutt, Julian. (2001). On the direct impact of ice on marine benthic
communities, a review. \emph{Polar Biology}, \emph{24}(8), 553--564.
doi:\href{https://doi.org/10.1007/s003000100262}{10.1007/s003000100262}

\leavevmode\vadjust pre{\hypertarget{ref-Gutt2015}{}}%
Gutt, Julian, Bertler, N., Bracegirdle, T. J., Buschmann, A., Comiso,
J., Hosie, G., \ldots{} Xavier, J. C. (2015). The {Southern Ocean}
ecosystem under multiple climate change stresses - an integrated
circumpolar assessment. \emph{Global Change Biology}, \emph{21}(4),
1434--1453.
doi:\href{https://doi.org/10.1111/gcb.12794}{10.1111/gcb.12794}

\leavevmode\vadjust pre{\hypertarget{ref-Gutt2021}{}}%
Gutt, Julian, Isla, E., Xavier, J. C., Adams, B. J., Ahn, I.-Y., Cheng,
C.-H. C., \ldots{} Wall, D. H. (2021). Antarctic ecosystems in
transition \textendash{} life between stresses and opportunities.
\emph{Biological Reviews}, \emph{96}(3), 798--821.
doi:\href{https://doi.org/10.1111/brv.12679}{10.1111/brv.12679}

\leavevmode\vadjust pre{\hypertarget{ref-Gutt2003}{}}%
Gutt, Julian, \& Piepenburg, D. (2003). Scale-dependent impact on
diversity of {Antarctic} benthos caused by grounding of icebergs.
\emph{Marine Ecology Progress Series}, \emph{253}, 77--83.
doi:\href{https://doi.org/10.3354/meps253077}{10.3354/meps253077}

\leavevmode\vadjust pre{\hypertarget{ref-Gutt2001}{}}%
Gutt, Julian, \& Starmans, A. (2001). Quantification of iceberg impact
and benthic recolonisation patterns in the {Weddell Sea} ({Antarctica}).
\emph{Polar Biology}, \emph{24}(8), 615--619.
doi:\href{https://doi.org/10.1007/s003000100263}{10.1007/s003000100263}

\leavevmode\vadjust pre{\hypertarget{ref-Gutt1996}{}}%
Gutt, J., Starmans, A., \& Dieckmann, G. (1996). Impact of iceberg
scouring on polar benthic habitats. \emph{Marine Ecology Progress
Series}, \emph{137}, 311--316.
doi:\href{https://doi.org/10.3354/meps137311}{10.3354/meps137311}

\leavevmode\vadjust pre{\hypertarget{ref-Hendry2018}{}}%
Hendry, K. R., Meredith, M. P., \& Ducklow, H. W. (2018). The marine
system of the {West Antarctic Peninsula}: Status and strategy for
progress. \emph{Philosophical Transactions of the Royal Society A:
Mathematical, Physical and Engineering Sciences}, \emph{376}(2122),
20170179.
doi:\href{https://doi.org/10.1098/rsta.2017.0179}{10.1098/rsta.2017.0179}

\leavevmode\vadjust pre{\hypertarget{ref-Hernandez2019}{}}%
Hernández, E. A., Lopez, J. L., Piquet, A. M.-T., Mac Cormack, W. P., \&
Buma, A. G. J. (2019). Changes in salinity and temperature drive marine
bacterial communities' structure at {Potter Cove}, {Antarctica}.
\emph{Polar Biology}, \emph{42}(12), 2177--2191.
doi:\href{https://doi.org/10.1007/s00300-019-02590-5}{10.1007/s00300-019-02590-5}

\leavevmode\vadjust pre{\hypertarget{ref-Hilborn2017}{}}%
Hilborn, R., Amoroso, R. O., Bogazzi, E., Jensen, O. P., Parma, A. M.,
Szuwalski, C., \& Walters, C. J. (2017). When does fishing forage
species affect their predators? \emph{Fisheries Research}, \emph{191},
211--221.
doi:\href{https://doi.org/10.1016/j.fishres.2017.01.008}{10.1016/j.fishres.2017.01.008}

\leavevmode\vadjust pre{\hypertarget{ref-Hill2012}{}}%
Hill, S. L., Keeble, K., Atkinson, A., \& Murphy, E. J. (2012). A
foodweb model to explore uncertainties in the {South Georgia} shelf
pelagic ecosystem. \emph{Deep Sea Research Part II: Topical Studies in
Oceanography}, \emph{59--60}, 237--252.
doi:\href{https://doi.org/10.1016/j.dsr2.2011.09.001}{10.1016/j.dsr2.2011.09.001}

\leavevmode\vadjust pre{\hypertarget{ref-Hill2013}{}}%
Hill, S. L., Phillips, T., \& Atkinson, A. (2013). Potential {Climate
Change Effects} on the {Habitat} of {Antarctic Krill} in the {Weddell
Quadrant} of the {Southern Ocean}. \emph{PLOS ONE}, \emph{8}(8), e72246.
doi:\href{https://doi.org/10.1371/journal.pone.0072246}{10.1371/journal.pone.0072246}

\leavevmode\vadjust pre{\hypertarget{ref-Hindell2020}{}}%
Hindell, M. A., Reisinger, R. R., Ropert-Coudert, Y., Hückstädt, L. A.,
Trathan, P. N., Bornemann, H., \ldots{} Raymond, B. (2020). Tracking of
marine predators to protect {Southern Ocean} ecosystems. \emph{Nature},
\emph{580}(7801), 87--92.
doi:\href{https://doi.org/10.1038/s41586-020-2126-y}{10.1038/s41586-020-2126-y}

\leavevmode\vadjust pre{\hypertarget{ref-Hoffmann2019}{}}%
Hoffmann, R., Al-Handal, A. Y., Wulff, A., Deregibus, D., Zacher, K.,
Quartino, M. L., \ldots{} Braeckman, U. (2019). Implications of {Glacial
Melt-Related Processes} on the {Potential Primary Production} of a
{Microphytobenthic Community} in {Potter Cove} ({Antarctica}).
\emph{Frontiers in Marine Science}, \emph{6}, 655.
doi:\href{https://doi.org/10.3389/fmars.2019.00655}{10.3389/fmars.2019.00655}

\leavevmode\vadjust pre{\hypertarget{ref-Iken2023}{}}%
Iken, K., Amsler, C. D., Gorman, K. B., Klein, A. G., Galloway, A. W.
E., Amsler, M. O., \ldots{} McClintock, J. B. (2023). Macroalgal input
into the coastal food web along a gradient of seasonal sea ice cover
along the {Western Antarctic Peninsula}. \emph{Marine Ecology Progress
Series}, \emph{718}, 1--22.
doi:\href{https://doi.org/10.3354/meps14388}{10.3354/meps14388}

\leavevmode\vadjust pre{\hypertarget{ref-Irons2000}{}}%
Irons, D. B., Kendall, S. J., Erickson, W. P., McDonald, L. L., \&
Lance, B. K. (2000). Nine {Years After} the {Exxon Valdez Oil Spill}:
{Effects} on {Marine Bird Populations} in {Prince William Sound},
{Alaska}. \emph{The Condor}, \emph{102}(4), 723--737.
doi:\href{https://doi.org/10.1093/condor/102.4.723}{10.1093/condor/102.4.723}

\leavevmode\vadjust pre{\hypertarget{ref-Isla2023}{}}%
Isla, E. (2023). Animal\textendash{{Energy Relationships}} in a
{Changing Ocean}: {The Case} of {Continental Shelf Macrobenthic
Communities} on the {Weddell Sea} and the {Vicinity} of the {Antarctic
Peninsula}. \emph{Biology}, \emph{12}(5), 659.
doi:\href{https://doi.org/10.3390/biology12050659}{10.3390/biology12050659}

\leavevmode\vadjust pre{\hypertarget{ref-Jacob2011}{}}%
Jacob, U., Thierry, A., Brose, U., Arntz, W. E., Berg, S., Brey, T.,
\ldots{} Dunne, J. A. (2011). The {Role} of {Body Size} in {Complex Food
Webs}: {A Cold Case}. In A. Belgrano (Ed.), \emph{Advances in
{Ecological Research}} (Vol. 45, pp. 181--223). {Academic Press}.
doi:\href{https://doi.org/10.1016/B978-0-12-386475-8.00005-8}{10.1016/B978-0-12-386475-8.00005-8}

\leavevmode\vadjust pre{\hypertarget{ref-Jerosch2018}{}}%
Jerosch, K., Pehlke, H., Monien, P., Scharf, F., Weber, L., Kuhn, G.,
\ldots{} Abele, D. (2018). Benthic meltwater fjord habitats formed by
rapid glacier recession on {King George Island}, {Antarctica}.
\emph{Philosophical Transactions of the Royal Society A: Mathematical,
Physical and Engineering Sciences}, \emph{376}(2122), 20170178.
doi:\href{https://doi.org/10.1098/rsta.2017.0178}{10.1098/rsta.2017.0178}

\leavevmode\vadjust pre{\hypertarget{ref-Jordan2007}{}}%
Jordán, F., Benedek, Z., \& Podani, J. (2007). Quantifying positional
importance in food webs: {A} comparison of centrality indices.
\emph{Ecological Modelling}, \emph{205}(1), 270--275.
doi:\href{https://doi.org/10.1016/j.ecolmodel.2007.02.032}{10.1016/j.ecolmodel.2007.02.032}

\leavevmode\vadjust pre{\hypertarget{ref-Kaminskyinprep}{}}%
Kaminsky, J., Bagur, M., Boraso, A., Schloss, I. R., Buschmann, A. H.,
\& Quartino, M. L. (in prep.). Kelp recruitment declines in a
sub-{Antarctic} ecosystem with high sediment inputs and changes in the
understory algae.

\leavevmode\vadjust pre{\hypertarget{ref-Kawaguchi2013}{}}%
Kawaguchi, S., Ishida, A., King, R., Raymond, B., Waller, N., Constable,
A., \ldots{} Ishimatsu, A. (2013). Risk maps for {Antarctic} krill under
projected {Southern Ocean} acidification. \emph{Nature Climate Change},
\emph{3}(9), 843--847.
doi:\href{https://doi.org/10.1038/nclimate1937}{10.1038/nclimate1937}

\leavevmode\vadjust pre{\hypertarget{ref-Kawaguchi2010}{}}%
Kawaguchi, So, Kurihara, H., King, R., Hale, L., Berli, T., Robinson, J.
P., \ldots{} Ishimatsu, A. (2010). Will krill fare well under {Southern
Ocean} acidification? \emph{Biology Letters}, \emph{7}(2), 288--291.
doi:\href{https://doi.org/10.1098/rsbl.2010.0777}{10.1098/rsbl.2010.0777}

\leavevmode\vadjust pre{\hypertarget{ref-Kawaguchi2009}{}}%
Kawaguchi, So, Nicol, S., \& Press, A. (2009). Direct effects of climate
change on the {Antarctic} krill fishery.

\leavevmode\vadjust pre{\hypertarget{ref-Klein2018}{}}%
Klein, E. S., Hill, S. L., Hinke, J. T., Phillips, T., \& Watters, G. M.
(2018). Impacts of rising sea temperature on krill increase risks for
predators in the {Scotia Sea}. \emph{PLOS ONE}, \emph{13}(1), e0191011.
doi:\href{https://doi.org/10.1371/journal.pone.0191011}{10.1371/journal.pone.0191011}

\leavevmode\vadjust pre{\hypertarget{ref-Kortsch2019}{}}%
Kortsch, S., Primicerio, R., Aschan, M., Lind, S., Dolgov, A. V., \&
Planque, B. (2019). Food-web structure varies along environmental
gradients in a high-latitude marine ecosystem. \emph{Ecography},
\emph{42}(2), 295--308.
doi:\href{https://doi.org/10.1111/ecog.03443}{10.1111/ecog.03443}

\leavevmode\vadjust pre{\hypertarget{ref-Kortsch2015}{}}%
Kortsch, S., Primicerio, R., Fossheim, M., Dolgov, A. V., \& Aschan, M.
(2015). Climate change alters the structure of arctic marine food webs
due to poleward shifts of boreal generalists. \emph{Proceedings of the
Royal Society B: Biological Sciences}, \emph{282}(1814), 20151546.
doi:\href{https://doi.org/10.1098/rspb.2015.1546}{10.1098/rspb.2015.1546}

\leavevmode\vadjust pre{\hypertarget{ref-Kratina2012}{}}%
Kratina, P., LeCraw, R. M., Ingram, T., \& Anholt, B. R. (2012).
Stability and persistence of food webs with omnivory: {Is} there a
general pattern? \emph{Ecosphere}, \emph{3}(6), art50.
doi:\href{https://doi.org/10.1890/ES12-00121.1}{10.1890/ES12-00121.1}

\leavevmode\vadjust pre{\hypertarget{ref-Lagger2018}{}}%
Lagger, C., Nime, M., Torre, L., Servetto, N., Tatián, M., \& Sahade, R.
(2018). Climate change, glacier retreat and a new ice-free island offer
new insights on {Antarctic} benthic responses. \emph{Ecography},
\emph{41}(4), 579--591.
doi:\href{https://doi.org/10.1111/ecog.03018}{10.1111/ecog.03018}

\leavevmode\vadjust pre{\hypertarget{ref-Lagger2017}{}}%
Lagger, C., Servetto, N., Torre, L., \& Sahade, R. (2017). Benthic
colonization in newly ice-free soft-bottom areas in an {Antarctic}
fjord. \emph{PLOS ONE}, \emph{12}(11), e0186756.
doi:\href{https://doi.org/10.1371/journal.pone.0186756}{10.1371/journal.pone.0186756}

\leavevmode\vadjust pre{\hypertarget{ref-Latorre2023}{}}%
Latorre, M. P., Iachetti, C. M., Schloss, I. R., Antoni, J., Malits, A.,
De La Rosa, F., \ldots{} Hernando, M. (2023). Summer heatwaves affect
coastal {Antarctic} plankton metabolism and community structure.
\emph{Journal of Experimental Marine Biology and Ecology}, \emph{567},
151926.
doi:\href{https://doi.org/10.1016/j.jembe.2023.151926}{10.1016/j.jembe.2023.151926}

\leavevmode\vadjust pre{\hypertarget{ref-Lattuca2018}{}}%
Lattuca, María Eugenia, Boy, C. C., Vanella, F. A., Barrantes, M. E., \&
Fernández, D. A. (2018). Thermal responses of three native fishes from
estuarine areas of the {Beagle Channel}, and their implications for
climate change. \emph{Hydrobiologia}, \emph{808}(1), 235--249.
doi:\href{https://doi.org/10.1007/s10750-017-3424-8}{10.1007/s10750-017-3424-8}

\leavevmode\vadjust pre{\hypertarget{ref-Lattuca2023}{}}%
Lattuca, María E., Vanella, F. A., Malanga, G., Rubel, M. D., Manríquez,
P. H., Torres, R., \ldots{} Fernández, D. A. (2023). Ocean acidification
and seasonal temperature extremes combine to impair the thermal
physiology of a sub-{Antarctic} fish. \emph{Science of The Total
Environment}, \emph{856}, 159284.
doi:\href{https://doi.org/10.1016/j.scitotenv.2022.159284}{10.1016/j.scitotenv.2022.159284}

\leavevmode\vadjust pre{\hypertarget{ref-Lin2021}{}}%
Lin, Y., Moreno, C., Marchetti, A., Ducklow, H., Schofield, O., Delage,
E., \ldots{} Cassar, N. (2021). Decline in plankton diversity and carbon
flux with reduced sea ice extent along the {Western Antarctic
Peninsula}. \emph{Nature Communications}, \emph{12}(1), 4948.
doi:\href{https://doi.org/10.1038/s41467-021-25235-w}{10.1038/s41467-021-25235-w}

\leavevmode\vadjust pre{\hypertarget{ref-Llorca2012}{}}%
Llorca, M., Farré, M., Tavano, M. S., Alonso, B., Koremblit, G., \&
Barceló, D. (2012). Fate of a broad spectrum of perfluorinated compounds
in soils and biota from {Tierra} del {Fuego} and {Antarctica}.
\emph{Environmental Pollution}, \emph{163}, 158--166.
doi:\href{https://doi.org/10.1016/j.envpol.2011.10.027}{10.1016/j.envpol.2011.10.027}

\leavevmode\vadjust pre{\hypertarget{ref-Lopez-Gappa2022}{}}%
López-Gappa, J. (2022). The {Impact} of {Global Change} on {Marine
Benthic Invertebrates}. In E. W. Helbling, M. A. Narvarte, R. A.
González, \& V. E. Villafañe (Eds.), \emph{Global {Change} in {Atlantic
Coastal Patagonian Ecosystems}: {A Journey Through Time}} (pp.
177--204). {Cham}: {Springer International Publishing}.
doi:\href{https://doi.org/10.1007/978-3-030-86676-1_8}{10.1007/978-3-030-86676-1\_8}

\leavevmode\vadjust pre{\hypertarget{ref-Lopez-Lopez2022}{}}%
López-López, L., Genner, M. J., Tarling, G. A., Saunders, R. A., \&
O'Gorman, E. J. (2022). Ecological {Networks} in the {Scotia Sea}:
{Structural Changes Across Latitude} and {Depth}. \emph{Ecosystems},
\emph{25}(2), 457--470.
doi:\href{https://doi.org/10.1007/s10021-021-00665-1}{10.1007/s10021-021-00665-1}

\leavevmode\vadjust pre{\hypertarget{ref-Manno2022}{}}%
Manno, C., Peck, V. L., Corsi, I., \& Bergami, E. (2022). Under
pressure: {Nanoplastics} as a further stressor for sub-{Antarctic}
pteropods already tackling ocean acidification. \emph{Marine Pollution
Bulletin}, \emph{174}, 113176.
doi:\href{https://doi.org/10.1016/j.marpolbul.2021.113176}{10.1016/j.marpolbul.2021.113176}

\leavevmode\vadjust pre{\hypertarget{ref-Marina2018}{}}%
Marina, T. I., Salinas, V., Cordone, G., Campana, G., Moreira, E.,
Deregibus, D., \ldots{} Momo, F. R. (2018). The {Food Web} of {Potter
Cove} ({Antarctica}): Complexity, structure and function.
\emph{Estuarine, Coastal and Shelf Science}, \emph{200}, 141--151.
doi:\href{https://doi.org/10.1016/j.ecss.2017.10.015}{10.1016/j.ecss.2017.10.015}

\leavevmode\vadjust pre{\hypertarget{ref-Marinainrev}{}}%
Marina, T. I., Schloss, I. R., Lovrich, G. A., Boy, C. C., Bruno, D. O.,
Capitanio, F. L., \ldots{} Riccialdelli, L. (in rev.). The complex
network of trophic interactions in a {Subantarctic} oceanic {Marine
Protected Area}. \emph{Marine Ecology Progress Series}.

\leavevmode\vadjust pre{\hypertarget{ref-Martinez2022}{}}%
Martínez, Patricia A., Wöhler, O. C., Troccoli, G. H., Di Marco, E., \&
Maydana, L. (2022). \emph{{Descripción de las capturas incidentales de
granaderos presentes en los lances dirigidos a merluza negra en la
pesquería argentina de arrastre: período 2010-2019.}} {Instituto
Nacional de Investigación y Desarrollo Pesquero (INIDEP)}.

\leavevmode\vadjust pre{\hypertarget{ref-Martinez2015}{}}%
Martínez, P. A., Wöhler, O. G., \& Troccoli, G. H. (2015). \emph{{La
evolución de la pesquería de merluza negra (Dissostichus eleginoides) en
el espacio marítimo argentino. Periodo 2003- 2014}} (\{Technical
report\} No. 11/15) (p. 12). {Instituto Nacional de Investigación y
Desarrollo Pesquero (INIDEP), Argentina}.

\leavevmode\vadjust pre{\hypertarget{ref-Matano2019}{}}%
Matano, Ricardo P., Palma, E. D., \& Combes, V. (2019). The {Burdwood
Bank Circulation}. \emph{Journal of Geophysical Research: Oceans},
\emph{124}(10), 6904--6926.
doi:\href{https://doi.org/10.1029/2019JC015001}{10.1029/2019JC015001}

\leavevmode\vadjust pre{\hypertarget{ref-Matano2010}{}}%
Matano, R. P., Palma, E. D., \& Piola, A. R. (2010). The influence of
the {Brazil} and {Malvinas Currents} on the {Southwestern Atlantic
Shelf} circulation. \emph{Ocean Science}, \emph{6}(4), 983--995.
doi:\href{https://doi.org/10.5194/os-6-983-2010}{10.5194/os-6-983-2010}

\leavevmode\vadjust pre{\hypertarget{ref-McCann1998}{}}%
McCann, K., Hastings, A., \& Huxel, G. R. (1998). Weak trophic
interactions and the balance of nature. \emph{Nature}, \emph{395}(6704),
794--798. doi:\href{https://doi.org/10.1038/27427}{10.1038/27427}

\leavevmode\vadjust pre{\hypertarget{ref-Menge1995}{}}%
Menge, B. A. (1995). Indirect {Effects} in {Marine Rocky Intertidal
Interaction Webs}: {Patterns} and {Importance}. \emph{Ecological
Monographs}, \emph{65}(1), 21--74.
doi:\href{https://doi.org/10.2307/2937158}{10.2307/2937158}

\leavevmode\vadjust pre{\hypertarget{ref-Meredith2018}{}}%
Meredith, M. P., Falk, U., Bers, A. V., Mackensen, A., Schloss, I. R.,
Ruiz Barlett, E., \ldots{} Abele, D. (2018). Anatomy of a glacial
meltwater discharge event in an {Antarctic} cove. \emph{Philosophical
Transactions of the Royal Society A: Mathematical, Physical and
Engineering Sciences}, \emph{376}(2122), 20170163.
doi:\href{https://doi.org/10.1098/rsta.2017.0163}{10.1098/rsta.2017.0163}

\leavevmode\vadjust pre{\hypertarget{ref-Meyer2017}{}}%
Meyer, B., Freier, U., Grimm, V., Groeneveld, J., Hunt, B. P. V.,
Kerwath, S., \ldots{} Yilmaz, N. I. (2017). The winter pack-ice zone
provides a sheltered but food-poor habitat for larval {Antarctic} krill.
\emph{Nature Ecology \& Evolution}, \emph{1}(12), 1853--1861.
doi:\href{https://doi.org/10.1038/s41559-017-0368-3}{10.1038/s41559-017-0368-3}

\leavevmode\vadjust pre{\hypertarget{ref-Michel2019}{}}%
Michel, L. N., Danis, B., Dubois, P., Eleaume, M., Fournier, J., Gallut,
C., \ldots{} Lepoint, G. (2019). Increased sea ice cover alters food web
structure in {East Antarctica}. \emph{Scientific Reports}, \emph{9}(1),
8062.
doi:\href{https://doi.org/10.1038/s41598-019-44605-5}{10.1038/s41598-019-44605-5}

\leavevmode\vadjust pre{\hypertarget{ref-Mintenbeck2012}{}}%
Mintenbeck, K., Barrera-Oro, E. R., Brey, T., Jacob, U., Knust, R.,
Mark, F. C., \ldots{} Arntz, W. E. (2012). Impact of {Climate Change} on
{Fishes} in {Complex Antarctic Ecosystems}. In U. Jacob \& G. Woodward
(Eds.), \emph{Advances in {Ecological Research}} (Vol. 46, pp.
361--426). {Burlington}: {Academic Press}.
doi:\href{https://doi.org/10.1016/B978-0-12-396992-7.00006-X}{10.1016/B978-0-12-396992-7.00006-X}

\leavevmode\vadjust pre{\hypertarget{ref-Montoya2009}{}}%
Montoya, JoséM., Woodward, G., Emmerson, M. C., \& Solé, R. V. (2009).
Press perturbations and indirect effects in real food webs.
\emph{Ecology}, \emph{90}(9), 2426--2433.
doi:\href{https://doi.org/10.1890/08-0657.1}{10.1890/08-0657.1}

\leavevmode\vadjust pre{\hypertarget{ref-Moore1974}{}}%
Moore, S. F., \& Dwyer, R. L. (1974). Effects of oil on marine
organisms: {A} critical assessment of published data. \emph{Water
Research}, \emph{8}(10), 819--827.
doi:\href{https://doi.org/10.1016/0043-1354(74)90028-1}{10.1016/0043-1354(74)90028-1}

\leavevmode\vadjust pre{\hypertarget{ref-Moy2009}{}}%
Moy, A. D., Howard, W. R., Bray, S. G., \& Trull, T. W. (2009). Reduced
calcification in modern {Southern Ocean} planktonic foraminifera.
\emph{Nature Geoscience}, \emph{2}(4), 276--280.
doi:\href{https://doi.org/10.1038/ngeo460}{10.1038/ngeo460}

\leavevmode\vadjust pre{\hypertarget{ref-Murphy2007}{}}%
Murphy, E. J., Trathan, P. N., Watkins, J. L., Reid, K., Meredith, M.
P., Forcada, J., \ldots{} Rothery, P. (2007). Climatically driven
fluctuations in {Southern Ocean} ecosystems. \emph{Proceedings of the
Royal Society B: Biological Sciences}, \emph{274}(1629), 3057--3067.
doi:\href{https://doi.org/10.1098/rspb.2007.1180}{10.1098/rspb.2007.1180}

\leavevmode\vadjust pre{\hypertarget{ref-Murphy2006}{}}%
Murphy, E. j., Watkins, J. l., Trathan, P. n., Reid, K., Meredith, M.
p., Thorpe, S. e., \ldots{} Fleming, A. h. (2006). Spatial and temporal
operation of the {Scotia Sea} ecosystem: A review of large-scale links
in a krill centred food web. \emph{Philosophical Transactions of the
Royal Society B: Biological Sciences}, \emph{362}(1477), 113--148.
doi:\href{https://doi.org/10.1098/rstb.2006.1957}{10.1098/rstb.2006.1957}

\leavevmode\vadjust pre{\hypertarget{ref-Nardi2019}{}}%
Nardi, C. F., Fernández, D. A., Vanella, F. A., \& Chalde, T. (2019).
The expansion of exotic {Chinook} salmon ({Oncorhynchus} tshawytscha) in
the extreme south of {Patagonia}: An environmental {DNA} approach.
\emph{Biological Invasions}, \emph{21}(4), 1415--1425.
doi:\href{https://doi.org/10.1007/s10530-018-1908-8}{10.1007/s10530-018-1908-8}

\leavevmode\vadjust pre{\hypertarget{ref-Neder2023}{}}%
Neder, C. (2023). \emph{{El bentos antártico y su respuesta al cambio
climático: una aproximación usando modelos de distribución de especies
como caso de estudio en caleta Potter}} (PhD thesis). Universidad
Nacional de Córdoba, {Córdoba, Argentina}.

\leavevmode\vadjust pre{\hypertarget{ref-Neder2022}{}}%
Neder, C., Fofonova, V., Androsov, A., Kuznetsov, I., Abele, D., Falk,
U., \ldots{} Jerosch, K. (2022). Modelling suspended particulate matter
dynamics at an {Antarctic} fjord impacted by glacier melt. \emph{Journal
of Marine Systems}, \emph{231}, 103734.
doi:\href{https://doi.org/10.1016/j.jmarsys.2022.103734}{10.1016/j.jmarsys.2022.103734}

\leavevmode\vadjust pre{\hypertarget{ref-Nilsson2016}{}}%
Nilsson, K. A., \& McCann, K. S. (2016). Interaction strength
revisited\textemdash clarifying the role of energy flux for food web
stability. \emph{Theoretical Ecology}, \emph{9}(1), 59--71.
doi:\href{https://doi.org/10.1007/s12080-015-0282-8}{10.1007/s12080-015-0282-8}

\leavevmode\vadjust pre{\hypertarget{ref-Nowacek2015}{}}%
Nowacek, D. P., Clark, C. W., Mann, D., Miller, P. J., Rosenbaum, H. C.,
Golden, J. S., \ldots{} Southall, B. L. (2015). Marine seismic surveys
and ocean noise: Time for coordinated and prudent planning.
\emph{Frontiers in Ecology and the Environment}, \emph{13}(7), 378--386.
doi:\href{https://doi.org/10.1890/130286}{10.1890/130286}

\leavevmode\vadjust pre{\hypertarget{ref-OGorman2012}{}}%
O'Gorman, E. J., Fitch, J. E., \& Crowe, T. P. (2012). Multiple
anthropogenic stressors and the structural properties of food webs.
\emph{Ecology}, \emph{93}(3), 441--448.
doi:\href{https://doi.org/10.1890/11-0982.1}{10.1890/11-0982.1}

\leavevmode\vadjust pre{\hypertarget{ref-OGorman2019}{}}%
O'Gorman, E. J., Petchey, O. L., Faulkner, K. J., Gallo, B., Gordon, T.
A. C., Neto-Cerejeira, J., \ldots{} Woodward, G. (2019). A simple model
predicts how warming simplifies wild food webs. \emph{Nature Climate
Change}, \emph{9}(8), 611--616.
doi:\href{https://doi.org/10.1038/s41558-019-0513-x}{10.1038/s41558-019-0513-x}

\leavevmode\vadjust pre{\hypertarget{ref-Ojeda2021}{}}%
Ojeda, M., Cossi, P. F., Rimondino, G. N., Chiesa, I. L., Boy, C. C., \&
Pérez, A. F. (2021). Microplastics pollution in the intertidal limpet,
{Nacella} magellanica, from {Beagle Channel} ({Argentina}).
\emph{Science of The Total Environment}, \emph{795}, 148866.
doi:\href{https://doi.org/10.1016/j.scitotenv.2021.148866}{10.1016/j.scitotenv.2021.148866}

\leavevmode\vadjust pre{\hypertarget{ref-Oliver2018}{}}%
Oliver, E. C. J., Donat, M. G., Burrows, M. T., Moore, P. J., Smale, D.
A., Alexander, L. V., \ldots{} Wernberg, T. (2018). Longer and more
frequent marine heatwaves over the past century. \emph{Nature
Communications}, \emph{9}(1), 1324.
doi:\href{https://doi.org/10.1038/s41467-018-03732-9}{10.1038/s41467-018-03732-9}

\leavevmode\vadjust pre{\hypertarget{ref-Olivier2019}{}}%
Olivier, P., Frelat, R., Bonsdorff, E., Kortsch, S., Kröncke, I.,
Möllmann, C., \ldots{} Nordström, M. C. (2019). Exploring the temporal
variability of a food web using long-term biomonitoring data.
\emph{Ecography}, \emph{42}(12), 2107--2121.
doi:\href{https://doi.org/10.1111/ecog.04461}{10.1111/ecog.04461}

\leavevmode\vadjust pre{\hypertarget{ref-Orgeira2021}{}}%
Orgeira, J. L., Alvarez, F., \& Salvó, C. S. (2021). The same pathway to
the {Weddell Sea} birdlife, after 65 years: Similarities in the species
composition, richness and abundances. \emph{Czech Polar Reports},
\emph{11}(2), 291--304.
doi:\href{https://doi.org/10.5817/CPR2021-2-20}{10.5817/CPR2021-2-20}

\leavevmode\vadjust pre{\hypertarget{ref-Orr2020}{}}%
Orr, J. A., Vinebrooke, R. D., Jackson, M. C., Kroeker, K. J., Kordas,
R. L., Mantyka-Pringle, C., \ldots{} Piggott, J. J. (2020). Towards a
unified study of multiple stressors: Divisions and common goals across
research disciplines. \emph{Proceedings of the Royal Society B:
Biological Sciences}, \emph{287}(1926), 20200421.
doi:\href{https://doi.org/10.1098/rspb.2020.0421}{10.1098/rspb.2020.0421}

\leavevmode\vadjust pre{\hypertarget{ref-Paine1996}{}}%
Paine, R. T., Ruesink, J. L., Sun, A., Soulanille, E. L., Wonham, M. J.,
Harley, C. D. G., \ldots{} Secord, D. L. (1996). {TROUBLE ON OILED
WATERS}: {Lessons} from the {\emph{Exxon Valdez}} {Oil Spill}.
\emph{Annual Review of Ecology and Systematics}, \emph{27}(1), 197--235.
doi:\href{https://doi.org/10.1146/annurev.ecolsys.27.1.197}{10.1146/annurev.ecolsys.27.1.197}

\leavevmode\vadjust pre{\hypertarget{ref-Pallin2023}{}}%
Pallin, L. J., Kellar, N. M., Steel, D., Botero-Acosta, N., Baker, C.
S., Conroy, J. A., \ldots{} Friedlaender, A. S. (2023). A surplus no
more? {Variation} in krill availability impacts reproductive rates of
{Antarctic} baleen whales. \emph{Global Change Biology}, \emph{29}(8),
2108--2121.
doi:\href{https://doi.org/10.1111/gcb.16559}{10.1111/gcb.16559}

\leavevmode\vadjust pre{\hypertarget{ref-Pante2023}{}}%
Pante, E., Simon-Bouhet, B., \& Irisson, J. (2023). Marmap: {Import},
{Plot} and {Analyze Bathymetric} and {Topographic Data}.

\leavevmode\vadjust pre{\hypertarget{ref-Pasotti2015}{}}%
Pasotti, F., Saravia, L. A., Troch, M. D., Tarantelli, M. S., Sahade,
R., \& Vanreusel, A. (2015). Benthic {Trophic Interactions} in an
{Antarctic Shallow Water Ecosystem Affected} by {Recent Glacier
Retreat}. \emph{PLOS ONE}, \emph{10}(11), e0141742.
doi:\href{https://doi.org/10.1371/journal.pone.0141742}{10.1371/journal.pone.0141742}

\leavevmode\vadjust pre{\hypertarget{ref-Pasti2021}{}}%
Pasti, A. T., Bovcon, N. D., Ruibal-Núñez, J., Navoa, X., Jacobi, K. J.,
\& Galván, D. E. (2021). The diet of {Mustelus} schmitti in areas with
and without commercial bottom trawling ({Central Patagonia},
{Southwestern Atlantic}): {Is} it evidence of trophic interaction with
the {Patagonian} shrimp fishery? \emph{Food Webs}, \emph{29}, e00214.
doi:\href{https://doi.org/10.1016/j.fooweb.2021.e00214}{10.1016/j.fooweb.2021.e00214}

\leavevmode\vadjust pre{\hypertarget{ref-Pauli2021}{}}%
Pauli, N.-C., Flintrop, C. M., Konrad, C., Pakhomov, E. A., Swoboda, S.,
Koch, F., \ldots{} Iversen, M. H. (2021). Krill and salp faecal pellets
contribute equally to the carbon flux at the {Antarctic Peninsula}.
\emph{Nature Communications}, \emph{12}(1), 7168.
doi:\href{https://doi.org/10.1038/s41467-021-27436-9}{10.1038/s41467-021-27436-9}

\leavevmode\vadjust pre{\hypertarget{ref-Pecuchet2022}{}}%
Pecuchet, L., Jørgensen, L. L., Dolgov, A. V., Eriksen, E., Husson, B.,
Skern-Mauritzen, M., \& Primicerio, R. (2022). Spatio-temporal turnover
and drivers of bentho-demersal community and food web structure in a
high-latitude marine ecosystem. \emph{Diversity and Distributions},
\emph{28}(12), 2503--2520.
doi:\href{https://doi.org/10.1111/ddi.13580}{10.1111/ddi.13580}

\leavevmode\vadjust pre{\hypertarget{ref-Pecuchet2020}{}}%
Pecuchet, L., Lindegren, M., Kortsch, S., Całkiewicz, J., Jurgensone,
I., Margonski, P., \ldots{} Nordström, M. C. (2020). Spatio-temporal
dynamics of multi-trophic communities reveal ecosystem-wide functional
reorganization. \emph{Ecography}, \emph{43}(2), 197--208.
doi:\href{https://doi.org/10.1111/ecog.04643}{10.1111/ecog.04643}

\leavevmode\vadjust pre{\hypertarget{ref-Perez2021}{}}%
Pérez, A. F., Ojeda, M., Cossi, P. F., Rimondino, G. N., Chiesa, I. L.,
Fraysse, C., \& Calcagno, J. (2021). {Microplásticos en organismos
marinos de altas latitudes}. In \emph{{I Jornadas de la Red Científica
Argentina para el estudio de los plásticos y sus impactos en el ambiente
(SEPIA)}}.

\leavevmode\vadjust pre{\hypertarget{ref-Perez2020}{}}%
Pérez, Analía F., Ojeda, M., Rimondino, G. N., Chiesa, I. L., Di Mauro,
R., Boy, C. C., \& Calcagno, J. A. (2020). First report of microplastics
presence in the mussel {Mytilus} chilensis from {Ushuaia Bay} ({Beagle
Channel}, {Tierra} del {Fuego}, {Argentina}). \emph{Marine Pollution
Bulletin}, \emph{161}, 111753.
doi:\href{https://doi.org/10.1016/j.marpolbul.2020.111753}{10.1016/j.marpolbul.2020.111753}

\leavevmode\vadjust pre{\hypertarget{ref-Perry2020}{}}%
Perry, F. A., Kawaguchi, S., Atkinson, A., Sailley, S. F., Tarling, G.
A., Mayor, D. J., \ldots{} Cooper, A. (2020).
Temperature\textendash{{Induced Hatch Failure}} and {Nauplii
Malformation} in {Antarctic Krill}. \emph{Frontiers in Marine Science},
\emph{7}.

\leavevmode\vadjust pre{\hypertarget{ref-Peterson2001}{}}%
Peterson, C. H. (2001). The {`{Exxon Valdez}'} oil spill in {Alaska}:
{Acute}, indirect and chronic effects on the ecosystem. In
\emph{Advances in {Marine Biology}} (Vol. 39, pp. 1--103). {Academic
Press}.
doi:\href{https://doi.org/10.1016/S0065-2881(01)39008-9}{10.1016/S0065-2881(01)39008-9}

\leavevmode\vadjust pre{\hypertarget{ref-Pineda-Metz2019a}{}}%
Pineda-Metz, S. E. A. (2019). \emph{Benthic communities of the {Weddell
Sea}: Past, present and future} (PhD thesis). University of Bremen,
{Germany}.

\leavevmode\vadjust pre{\hypertarget{ref-Pineda-Metz2020}{}}%
Pineda-Metz, S. E. A., Gerdes, D., \& Richter, C. (2020). Benthic fauna
declined on a whitening {Antarctic} continental shelf. \emph{Nature
Communications}, \emph{11}(1), 2226.
doi:\href{https://doi.org/10.1038/s41467-020-16093-z}{10.1038/s41467-020-16093-z}

\leavevmode\vadjust pre{\hypertarget{ref-Pinkerton2021}{}}%
Pinkerton, M. H., Boyd, P. W., Deppeler, S., Hayward, A., Höfer, J., \&
Moreau, S. (2021). Evidence for the {Impact} of {Climate Change} on
{Primary Producers} in the {Southern Ocean}. \emph{Frontiers in Ecology
and Evolution}, \emph{9}.

\leavevmode\vadjust pre{\hypertarget{ref-Pinones2016}{}}%
Piñones, A., \& Fedorov, A. V. (2016). Projected changes of {Antarctic}
krill habitat by the end of the 21st century. \emph{Geophysical Research
Letters}, \emph{43}(16), 8580--8589.
doi:\href{https://doi.org/10.1002/2016GL069656}{10.1002/2016GL069656}

\leavevmode\vadjust pre{\hypertarget{ref-Piola2009}{}}%
Piola, A. R., \& Falabella, V. (2009). {El mar Patagónico}. In
\emph{{Atlas del Mar Patagónico. Especies y Espacios}} (Falabella, V.,
Campagna, C. y Croxall, J. Ed., pp. 54--75). {Buenos Aires, Argentina}:
{Wildlife Conservation Society y Birdlife Internacional}.

\leavevmode\vadjust pre{\hypertarget{ref-Polazzo2022}{}}%
Polazzo, F., Marina, T. I., Crettaz-Minaglia, M., \& Rico, A. (2022).
Food web rewiring drives long-term compositional differences and
late-disturbance interactions at the community level. \emph{Proceedings
of the National Academy of Sciences}, \emph{119}(17), e2117364119.
doi:\href{https://doi.org/10.1073/pnas.2117364119}{10.1073/pnas.2117364119}

\leavevmode\vadjust pre{\hypertarget{ref-Poloczanska2013}{}}%
Poloczanska, E. S., Brown, C. J., Sydeman, W. J., Kiessling, W.,
Schoeman, D. S., Moore, P. J., \ldots{} Richardson, A. J. (2013). Global
imprint of climate change on marine life. \emph{Nature Climate Change},
\emph{3}(10), 919--925.
doi:\href{https://doi.org/10.1038/nclimate1958}{10.1038/nclimate1958}

\leavevmode\vadjust pre{\hypertarget{ref-Quartino2008}{}}%
Quartino, M. L., \& Boraso de Zaixso, A. L. (2008). Summer macroalgal
biomass in {Potter Cove}, {South Shetland Islands}, {Antarctica}: Its
production and flux to the ecosystem. \emph{Polar Biology},
\emph{31}(3), 281--294.
doi:\href{https://doi.org/10.1007/s00300-007-0356-1}{10.1007/s00300-007-0356-1}

\leavevmode\vadjust pre{\hypertarget{ref-Quartino2013}{}}%
Quartino, María Liliana, Deregibus, D., Campana, G. L., Latorre, G. E.
J., \& Momo, F. R. (2013). Evidence of {Macroalgal Colonization} on
{Newly Ice-Free Areas} following {Glacial Retreat} in {Potter Cove}
({South Shetland Islands}), {Antarctica}. \emph{PLoS ONE}, \emph{8}(3),
e58223.
doi:\href{https://doi.org/10.1371/journal.pone.0058223}{10.1371/journal.pone.0058223}

\leavevmode\vadjust pre{\hypertarget{ref-Raymond2011}{}}%
Raymond, B. (2011). \emph{A circumpolar pelagic regionalisation of the
{Southern Ocean}. {CCAMLR WS-MPA}, {Brest}, {France}, 29
{Aug}\textendash 2 {Sep} 2011. {Document WS-MPA-11}/6.
{http://data.aad.gov.au/regionalisation}}.

\leavevmode\vadjust pre{\hypertarget{ref-Riccialdelli2020}{}}%
Riccialdelli, L., Becker, Y. A., Fioramonti, N. E., Torres, M., Bruno,
D. O., Rey, A. R., \& Fernández, D. A. (2020). Trophic structure of
southern marine ecosystems: A comparative isotopic analysis from the
{Beagle Channel} to the oceanic {Burdwood Bank} area under a wasp-waist
assumption. \emph{Marine Ecology Progress Series}, \emph{655}, 1--27.
doi:\href{https://doi.org/10.3354/meps13524}{10.3354/meps13524}

\leavevmode\vadjust pre{\hypertarget{ref-Riccialdelli2013}{}}%
Riccialdelli, L., Newsome, S. D., Dellabianca, N. A., Bastida, R.,
Fogel, M. L., \& Goodall, R. N. P. (2013). Ontogenetic diet shift in
{Commerson}'s dolphin ({Cephalorhynchus} commersonii commersonii) off
{Tierra} del {Fuego}. \emph{Polar Biology}, \emph{36}(5), 617--627.
doi:\href{https://doi.org/10.1007/s00300-013-1289-5}{10.1007/s00300-013-1289-5}

\leavevmode\vadjust pre{\hypertarget{ref-Riccialdelli2017}{}}%
Riccialdelli, L., Newsome, S. D., Fogel, M. L., \& Fernández, D. A.
(2017). Trophic interactions and food web structure of a subantarctic
marine food web in the {Beagle Channel}: {Bahía Lapataia}, {Argentina}.
\emph{Polar Biology}, \emph{40}(4), 807--821.
doi:\href{https://doi.org/10.1007/s00300-016-2007-x}{10.1007/s00300-016-2007-x}

\leavevmode\vadjust pre{\hypertarget{ref-Rincon-Diaz2021}{}}%
Rincón-Díaz, M. P., Bovcon, N. D., Cochia, P. D., Góngora, M. E., \&
Galván, D. E. (2021). Fish functional diversity as an indicator of
resilience to industrial fishing in {Patagonia Argentina}. \emph{Journal
of Fish Biology}, \emph{99}(5), 1650--1667.
doi:\href{https://doi.org/10.1111/jfb.14873}{10.1111/jfb.14873}

\leavevmode\vadjust pre{\hypertarget{ref-Rodriguez2022}{}}%
Rodriguez, I. D., Marina, T. I., Schloss, I. R., \& Saravia, L. A.
(2022). Marine food webs are more complex but less stable in
sub-{Antarctic} ({Beagle Channel}, {Argentina}) than in {Antarctic}
({Potter Cove}, {Antarctic Peninsula}) regions. \emph{Marine
Environmental Research}, \emph{174}, 105561.
doi:\href{https://doi.org/10.1016/j.marenvres.2022.105561}{10.1016/j.marenvres.2022.105561}

\leavevmode\vadjust pre{\hypertarget{ref-Rossi2019}{}}%
Rossi, L., Sporta Caputi, S., Calizza, E., Careddu, G., Oliverio, M.,
Schiaparelli, S., \& Costantini, M. L. (2019). Antarctic food web
architecture under varying dynamics of sea ice cover. \emph{Scientific
Reports}, \emph{9}(1), 12454.
doi:\href{https://doi.org/10.1038/s41598-019-48245-7}{10.1038/s41598-019-48245-7}

\leavevmode\vadjust pre{\hypertarget{ref-Rowlands2021}{}}%
Rowlands, E., Galloway, T., Cole, M., Lewis, C., Peck, V., Thorpe, S.,
\& Manno, C. (2021). The {Effects} of {Combined Ocean Acidification} and
{Nanoplastic Exposures} on the {Embryonic Development} of {Antarctic
Krill}. \emph{Frontiers in Marine Science}, \emph{8}.

\leavevmode\vadjust pre{\hypertarget{ref-Ruckamp2011}{}}%
Rückamp, M., Braun, M., Suckro, S., \& Blindow, N. (2011). Observed
glacial changes on the {King George Island} ice cap, {Antarctica}, in
the last decade. \emph{Global and Planetary Change}, \emph{79}(1-2),
99--109.
doi:\href{https://doi.org/10.1016/j.gloplacha.2011.06.009}{10.1016/j.gloplacha.2011.06.009}

\leavevmode\vadjust pre{\hypertarget{ref-RuibalNunez2020}{}}%
Ruibal Nuñez, J. (2020). \emph{{Impacto ecológico de la actividad
pesquera en las poblaciones de condrictios en el litoral de la Provincia
de Chubut y Golfo San Jorge}} (PhD thesis). Universidad Nacional de Mar
del Plata (UNMdP), {Argentina}.

\leavevmode\vadjust pre{\hypertarget{ref-Sahade2015}{}}%
Sahade, R., Lagger, C., Torre, L., Momo, F., Monien, P., Schloss, I.,
\ldots{} Abele, D. (2015). Climate change and glacier retreat drive
shifts in an {Antarctic} benthic ecosystem. \emph{Science Advances},
\emph{1}(10), e1500050.
doi:\href{https://doi.org/10.1126/sciadv.1500050}{10.1126/sciadv.1500050}

\leavevmode\vadjust pre{\hypertarget{ref-Sailley2013}{}}%
Sailley, S. F., Ducklow, H. W., Moeller, H. V., Fraser, W. R.,
Schofield, O. M., Steinberg, D. K., \ldots{} Doney, S. C. (2013). Carbon
fluxes and pelagic ecosystem dynamics near two western {Antarctic
Peninsula Adélie} penguin colonies: An inverse model approach.
\emph{Marine Ecology Progress Series}, \emph{492}, 253--272.
doi:\href{https://doi.org/10.3354/meps10534}{10.3354/meps10534}

\leavevmode\vadjust pre{\hypertarget{ref-Saravia2021}{}}%
Saravia, J., Paschke, K., Oyarzún-Salazar, R., Cheng, C.-H. C., Navarro,
J. M., \& Vargas-Chacoff, L. (2021). Effects of warming rates on
physiological and molecular components of response to {CTMax} heat
stress in the {Antarctic} fish {Harpagifer} antarcticus. \emph{Journal
of Thermal Biology}, \emph{99}, 103021.
doi:\href{https://doi.org/10.1016/j.jtherbio.2021.103021}{10.1016/j.jtherbio.2021.103021}

\leavevmode\vadjust pre{\hypertarget{ref-Saunders2019}{}}%
Saunders, R. A., Hill, S. L., Tarling, G. A., \& Murphy, E. J. (2019).
Myctophid {Fish} ({Family Myctophidae}) {Are Central Consumers} in the
{Food Web} of the {Scotia Sea} ({Southern Ocean}). \emph{Frontiers in
Marine Science}, \emph{6}.

\leavevmode\vadjust pre{\hypertarget{ref-Schejter2021}{}}%
Schejter, L., \& Albano, M. (2021). Benthic communities at the marine
protected area {Namuncurá}/{Burdwood} bank, {SW Atlantic Ocean}:
Detection of vulnerable marine ecosystems and contributions to the
assessment of the rezoning process. \emph{Polar Biology}, \emph{44}(10),
2023--2037.
doi:\href{https://doi.org/10.1007/s00300-021-02936-y}{10.1007/s00300-021-02936-y}

\leavevmode\vadjust pre{\hypertarget{ref-Schloss2012}{}}%
Schloss, I. R., Abele, D., Moreau, S., Demers, S., Bers, A. V.,
González, O., \& Ferreyra, G. A. (2012). Response of phytoplankton
dynamics to 19-year (1991\textendash 2009) climate trends in {Potter
Cove} ({Antarctica}). \emph{Journal of Marine Systems}, \emph{92}(1),
53--66.
doi:\href{https://doi.org/10.1016/j.jmarsys.2011.10.006}{10.1016/j.jmarsys.2011.10.006}

\leavevmode\vadjust pre{\hypertarget{ref-Schloss2023}{}}%
Schloss, I. R., Pizarro, G., Cadaillon, A. M., Giesecke, R., Hernando,
M. P., Almandoz, G. O., \ldots{} Ruiz, C. (2023). Alexandrium catenella
dynamics and paralytic shellfish toxins distribution along the {Beagle
Channel} (southern {Patagonia}). \emph{Journal of Marine Systems},
\emph{239}, 103856.
doi:\href{https://doi.org/10.1016/j.jmarsys.2022.103856}{10.1016/j.jmarsys.2022.103856}

\leavevmode\vadjust pre{\hypertarget{ref-Schmidt2011}{}}%
Schmidt, K., Atkinson, A., Steigenberger, S., Fielding, S., Lindsay, M.
C. M., Pond, D. W., \ldots{} Achterberg, E. P. (2011). Seabed foraging
by {Antarctic} krill: {Implications} for stock assessment,
bentho-pelagic coupling, and the vertical transfer of iron.
\emph{Limnology and Oceanography}, \emph{56}(4), 1411--1428.
doi:\href{https://doi.org/10.4319/lo.2011.56.4.1411}{10.4319/lo.2011.56.4.1411}

\leavevmode\vadjust pre{\hypertarget{ref-Seco2021}{}}%
Seco, J., Aparício, S., Brierley, A. S., Bustamante, P., Ceia, F. R.,
Coelho, J. P., \ldots{} Xavier, J. C. (2021). Mercury biomagnification
in a {Southern Ocean} food web. \emph{Environmental Pollution},
\emph{275}, 116620.
doi:\href{https://doi.org/10.1016/j.envpol.2021.116620}{10.1016/j.envpol.2021.116620}

\leavevmode\vadjust pre{\hypertarget{ref-Sgarlatta2023}{}}%
Sgarlatta, M. (2023). \emph{Mechanisms driving tropicalization of
temperate reef fish assemblages} (Thesis). UNSW Sydney.
doi:/\href{https://doi.org/10.26190/unsworks/25104}{10.26190/unsworks/25104}

\leavevmode\vadjust pre{\hypertarget{ref-Siniff2008}{}}%
Siniff, D. B., Garrott, R. A., Rotella, J. J., Fraser, W. R., \& Ainley,
D. G. (2008). Opinion: {Projecting} the effects of environmental change
on {Antarctic} seals. \emph{Antarctic Science}, \emph{20}(5), 425--435.
doi:\href{https://doi.org/10.1017/S0954102008001351}{10.1017/S0954102008001351}

\leavevmode\vadjust pre{\hypertarget{ref-Smale2008}{}}%
Smale, Dan A., \& Barnes, D. K. A. (2008). Likely responses of the
{Antarctic} benthos to climate-related changes in physical disturbance
during the 21st century, based primarily on evidence from the {West
Antarctic Peninsula} region. \emph{Ecography}, \emph{31}(3), 289--305.
doi:\href{https://doi.org/10.1111/j.0906-7590.2008.05456.x}{10.1111/j.0906-7590.2008.05456.x}

\leavevmode\vadjust pre{\hypertarget{ref-Smale2008a}{}}%
Smale, Dan A., Brown, K. M., Barnes, D. K. A., Fraser, K. P. P., \&
Clarke, A. (2008). Ice {Scour Disturbance} in {Antarctic Waters}.
\emph{Science}, \emph{321}(5887), 371--371.
doi:\href{https://doi.org/10.1126/science.1158647}{10.1126/science.1158647}

\leavevmode\vadjust pre{\hypertarget{ref-Soeffker2022}{}}%
Soeffker, M., Hollyman, P. R., Collins, M. A., Hogg, O. T., Riley, A.,
Laptikhovsky, V., \ldots{} Darby, C. (2022). Contrasting life-history
traits of two toothfish ({Dissostichus} spp.) species at their range
edge around the {South Sandwich Islands}. \emph{Deep Sea Research Part
II: Topical Studies in Oceanography}, \emph{201}, 105098.
doi:\href{https://doi.org/10.1016/j.dsr2.2022.105098}{10.1016/j.dsr2.2022.105098}

\leavevmode\vadjust pre{\hypertarget{ref-Souza2018}{}}%
Souza, M. R. D. P. de, Herrerias, T., Zaleski, T., Forgati, M.,
Kandalski, P. K., Machado, C., \ldots{} Donatti, L. (2018). Heat stress
in the heart and muscle of the {Antarctic} fishes {Notothenia} rossii
and {Notothenia} coriiceps: {Carbohydrate} metabolism and antioxidant
defence. \emph{Biochimie}, \emph{146}, 43--55.
doi:\href{https://doi.org/10.1016/j.biochi.2017.11.010}{10.1016/j.biochi.2017.11.010}

\leavevmode\vadjust pre{\hypertarget{ref-Steele2010}{}}%
Steele, J. H., Thorpe, S. A., \& Turekian, K. K. (2010). \emph{Ocean
{Currents}}. {Academic Press}.

\leavevmode\vadjust pre{\hypertarget{ref-Stouffer2011}{}}%
Stouffer, D. B., \& Bascompte, J. (2011). Compartmentalization increases
food-web persistence. \emph{Proceedings of the National Academy of
Sciences}, \emph{108}(9), 3648--3652.
doi:\href{https://doi.org/10.1073/pnas.1014353108}{10.1073/pnas.1014353108}

\leavevmode\vadjust pre{\hypertarget{ref-Strobel2013}{}}%
Strobel, A., Leo, E., Pörtner, H. O., \& Mark, F. C. (2013). Elevated
temperature and {PCO2} shift metabolic pathways in differentially
oxidative tissues of {Notothenia} rossii. \emph{Comparative Biochemistry
and Physiology Part B: Biochemistry and Molecular Biology},
\emph{166}(1), 48--57.
doi:\href{https://doi.org/10.1016/j.cbpb.2013.06.006}{10.1016/j.cbpb.2013.06.006}

\leavevmode\vadjust pre{\hypertarget{ref-Strydom2021}{}}%
Strydom, T., Catchen, M. D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., \ldots{} Poisot, T. (2021). A roadmap towards
predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences}, \emph{376}(1837), 20210063.
doi:\href{https://doi.org/10.1098/rstb.2021.0063}{10.1098/rstb.2021.0063}

\leavevmode\vadjust pre{\hypertarget{ref-Tamini2023}{}}%
Tamini, L. L., Dellacasa, R. F., Chavez, L. N., Marinao, C. J., Góngora,
M. E., Crawford, R., \& Frere, E. (2023). Bird scaring lines reduce
seabird mortality in mid-water and bottom trawlers in {Argentina}.
\emph{ICES Journal of Marine Science}, fsad109.
doi:\href{https://doi.org/10.1093/icesjms/fsad109}{10.1093/icesjms/fsad109}

\leavevmode\vadjust pre{\hypertarget{ref-Tarling2018}{}}%
Tarling, G. A., Ward, P., \& Thorpe, S. E. (2018). Spatial distributions
of {Southern Ocean} mesozooplankton communities have been resilient to
long-term surface warming. \emph{Global Change Biology}, \emph{24}(1),
132--142.
doi:\href{https://doi.org/10.1111/gcb.13834}{10.1111/gcb.13834}

\leavevmode\vadjust pre{\hypertarget{ref-Teixido2002}{}}%
Teixidó, N., Garrabou, J., \& Arntz, W. E. (2002). Spatial pattern
quantification of {Antarctic} benthic communities using landscape
indices. \emph{Marine Ecology Progress Series}, \emph{242}, 1--14.
doi:\href{https://doi.org/10.3354/meps242001}{10.3354/meps242001}

\leavevmode\vadjust pre{\hypertarget{ref-Thompson2012}{}}%
Thompson, R. M., Brose, U., Dunne, J. A., Hall, R. O., Hladyz, S.,
Kitching, R. L., \ldots{} Tylianakis, J. M. (2012). Food webs:
Reconciling the structure and function of biodiversity. \emph{Trends in
Ecology \& Evolution}, \emph{27}(12), 689--697.
doi:\href{https://doi.org/10.1016/j.tree.2012.08.005}{10.1016/j.tree.2012.08.005}

\leavevmode\vadjust pre{\hypertarget{ref-Thompson2007}{}}%
Thompson, R. M., Hemberg, M., Starzomski, B. M., \& Shurin, J. B.
(2007). Trophic {Levels} and {Trophic Tangles}: {The Prevalence} of
{Omnivory} in {Real Food Webs}. \emph{Ecology}, \emph{88}(3), 612--617.
doi:\href{https://doi.org/10.1890/05-1454}{10.1890/05-1454}

\leavevmode\vadjust pre{\hypertarget{ref-Tombesi2020}{}}%
Tombesi, M. L., Rabufetti, F., \& Lovrich, G. A. (2020). Las áreas
marinas protegidas en la {Argentina}. \emph{La Lupa. Colección Fueguina
De Divulgación Científica}, \emph{16}, 2--7.
doi:\url{https://www.coleccionlalupa.com.ar/index.php/lalupa/article/view/79}

\leavevmode\vadjust pre{\hypertarget{ref-Trathan2023}{}}%
Trathan, Philip N. (2023). The future of the {South Georgia} and {South
Sandwich Islands} marine protected area in a changing environment: {The}
choice between industrial fisheries, or ecosystem protection.
\emph{Marine Policy}, \emph{155}, 105773.
doi:\href{https://doi.org/10.1016/j.marpol.2023.105773}{10.1016/j.marpol.2023.105773}

\leavevmode\vadjust pre{\hypertarget{ref-Trathan2021}{}}%
Trathan, P. N., Fielding, S., Hollyman, P. R., Murphy, E. J.,
Warwick-Evans, V., \& Collins, M. A. (2021). Enhancing the ecosystem
approach for the fishery for {Antarctic} krill within the complex,
variable, and changing ecosystem at {South Georgia}. \emph{ICES Journal
of Marine Science}, \emph{78}(6), 2065--2081.
doi:\href{https://doi.org/10.1093/icesjms/fsab092}{10.1093/icesjms/fsab092}

\leavevmode\vadjust pre{\hypertarget{ref-Trebilco2020}{}}%
Trebilco, R., Melbourne-Thomas, J., \& Constable, A. J. (2020). The
policy relevance of {Southern Ocean} food web structure: {Implications}
of food web change for fisheries, conservation and carbon sequestration.
\emph{Marine Policy}, \emph{115}, 103832.
doi:\href{https://doi.org/10.1016/j.marpol.2020.103832}{10.1016/j.marpol.2020.103832}

\leavevmode\vadjust pre{\hypertarget{ref-Turner2020}{}}%
Turner, J., Guarino, M. V., Arnatt, J., Jena, B., Marshall, G. J.,
Phillips, T., \ldots{} Cavanagh, R. (2020). Recent {Decrease} of {Summer
Sea Ice} in the {Weddell Sea}, {Antarctica}. \emph{Geophysical Research
Letters}, \emph{47}(11), e2020GL087127.
doi:\href{https://doi.org/10.1029/2020GL087127}{10.1029/2020GL087127}

\leavevmode\vadjust pre{\hypertarget{ref-Verges2019}{}}%
Vergés, A., McCosker, E., Mayer-Pinto, M., Coleman, M. A., Wernberg, T.,
Ainsworth, T., \& Steinberg, P. D. (2019). Tropicalisation of temperate
reefs: {Implications} for ecosystem functions and management actions.
\emph{Functional Ecology}, \emph{33}(6), 1000--1013.
doi:\href{https://doi.org/10.1111/1365-2435.13310}{10.1111/1365-2435.13310}

\leavevmode\vadjust pre{\hypertarget{ref-Verges2014}{}}%
Vergés, A., Steinberg, P. D., Hay, M. E., Poore, A. G. B., Campbell, A.
H., Ballesteros, E., \ldots{} Wilson, S. K. (2014). The tropicalization
of temperate marine ecosystems: Climate-mediated changes in herbivory
and community phase shifts. \emph{Proceedings of the Royal Society B:
Biological Sciences}, \emph{281}(1789), 20140846.
doi:\href{https://doi.org/10.1098/rspb.2014.0846}{10.1098/rspb.2014.0846}

\leavevmode\vadjust pre{\hypertarget{ref-Vinuesa2005}{}}%
Vinuesa, J. H. (2005). Distribución de crustáceos decápodos y
estomatópodos del golfo {San Jorge}, {Argentina}. \emph{Revista de
Biología Marina y Oceanografía}, \emph{40}(1), 7--21.
doi:\href{https://doi.org/10.4067/S0718-19572005000100002}{10.4067/S0718-19572005000100002}

\leavevmode\vadjust pre{\hypertarget{ref-Vinuesa2007}{}}%
Vinuesa, J. H., \& Varisco, M. (2007). Trophic ecology of the lobster
krill {Munida} gregaria in {San Jorge Gulf}, {Argentina}.
\emph{Investigaciones Marinas}, \emph{35}(2), 25--34.
doi:\href{https://doi.org/10.4067/S0717-71782007000200003}{10.4067/S0717-71782007000200003}

\leavevmode\vadjust pre{\hypertarget{ref-Vucic-Pestic2011}{}}%
Vucic-Pestic, O., Ehnes, R. B., Rall, B. C., \& Brose, U. (2011).
Warming up the system: Higher predator feeding rates but lower energetic
efficiencies. \emph{Global Change Biology}, \emph{17}(3), 1301--1310.
doi:\href{https://doi.org/10.1111/j.1365-2486.2010.02329.x}{10.1111/j.1365-2486.2010.02329.x}

\leavevmode\vadjust pre{\hypertarget{ref-Ward2018}{}}%
Ward, P., Tarling, G. A., \& Thorpe, S. E. (2018). Temporal changes in
abundances of large calanoid copepods in the {Scotia Sea}: Comparing the
1930s with contemporary times. \emph{Polar Biology}, \emph{41}(11),
2297--2310.
doi:\href{https://doi.org/10.1007/s00300-018-2369-3}{10.1007/s00300-018-2369-3}

\leavevmode\vadjust pre{\hypertarget{ref-Wege2021}{}}%
Wege, M., Salas, L., \& LaRue, M. (2021). Ice matters: {Life-history}
strategies of two {Antarctic} seals dictate climate change eventualities
in the {Weddell Sea}. \emph{Global Change Biology}, \emph{27}(23),
6252--6262.
doi:\href{https://doi.org/10.1111/gcb.15828}{10.1111/gcb.15828}

\leavevmode\vadjust pre{\hypertarget{ref-Whitehouse2008}{}}%
Whitehouse, M. J., Meredith, M. P., Rothery, P., Atkinson, A., Ward, P.,
\& Korb, R. E. (2008). Rapid warming of the ocean around {South
Georgia}, {Southern Ocean}, during the 20th century: {Forcings},
characteristics and implications for lower trophic levels. \emph{Deep
Sea Research Part I: Oceanographic Research Papers}, \emph{55}(10),
1218--1228.
doi:\href{https://doi.org/10.1016/j.dsr.2008.06.002}{10.1016/j.dsr.2008.06.002}

\leavevmode\vadjust pre{\hypertarget{ref-Whitworth1980}{}}%
Whitworth, T. (1980). Zonation and geostrophic flow of the {Antarctic}
circumpolar current at {Drake Passage}. \emph{Deep Sea Research Part A.
Oceanographic Research Papers}, \emph{27}(7), 497--507.
doi:\href{https://doi.org/10.1016/0198-0149(80)90036-9}{10.1016/0198-0149(80)90036-9}

\leavevmode\vadjust pre{\hypertarget{ref-Williams2000}{}}%
Williams, R. J., \& Martinez, N. D. (2000). Simple rules yield complex
food webs. \emph{Nature}, \emph{404}(6774), 180--183.
doi:\href{https://doi.org/10.1038/35004572}{10.1038/35004572}

\leavevmode\vadjust pre{\hypertarget{ref-Wood2015}{}}%
Wood, S. A., Russell, R., Hanson, D., Williams, R. J., \& Dunne, J. A.
(2015). Effects of spatial scale of sampling on food web structure.
\emph{Ecology and Evolution}, \emph{5}(17), 3769--3782.
doi:\href{https://doi.org/10.1002/ece3.1640}{10.1002/ece3.1640}

\leavevmode\vadjust pre{\hypertarget{ref-Woods2023}{}}%
Woods, B. L., Van de Putte, A. P., Hindell, M. A., Raymond, B.,
Saunders, R. A., Walters, A., \& Trebilco, R. (2023). Species
distribution models describe spatial variability in mesopelagic fish
abundance in the {Southern Ocean}. \emph{Frontiers in Marine Science},
\emph{9}.

\leavevmode\vadjust pre{\hypertarget{ref-Woodward2010}{}}%
Woodward, G., Benstead, J. P., Beveridge, O. S., Blanchard, J., Brey,
T., Brown, L. E., \ldots{} Yvon-Durocher, G. (2010). Ecological
{Networks} in a {Changing Climate}. In G. Woodward (Ed.), \emph{Advances
in {Ecological Research}} (Vol. 42, pp. 71--138). {Academic Press}.
doi:\href{https://doi.org/10.1016/B978-0-12-381363-3.00002-2}{10.1016/B978-0-12-381363-3.00002-2}

\leavevmode\vadjust pre{\hypertarget{ref-Wu2012}{}}%
Wu, L., Cai, W., Zhang, L., Nakamura, H., Timmermann, A., Joyce, T.,
\ldots{} Giese, B. (2012). Enhanced warming over the global subtropical
western boundary currents. \emph{Nature Climate Change}, \emph{2}(3),
161--166.
doi:\href{https://doi.org/10.1038/nclimate1353}{10.1038/nclimate1353}

\leavevmode\vadjust pre{\hypertarget{ref-Yorio2009}{}}%
Yorio, P. (2009). Marine protected areas, spatial scales, and
governance: Implications for the conservation of breeding seabirds.
\emph{Conservation Letters}, \emph{2}(4), 171--178.
doi:\href{https://doi.org/10.1111/j.1755-263X.2009.00062.x}{10.1111/j.1755-263X.2009.00062.x}

\leavevmode\vadjust pre{\hypertarget{ref-Yorio2017}{}}%
Yorio, P., González-Zevallos, D., Gatto, A., Biagioni, O., \& Castillo,
J. (2017). Relevance of forage fish in the diet of {Magellanic} penguins
breeding in northern {Patagonia}, {Argentina}. \emph{Marine Biology
Research}, \emph{13}(6), 603--617.
doi:\href{https://doi.org/10.1080/17451000.2016.1273529}{10.1080/17451000.2016.1273529}

\end{CSLReferences}

\end{document}