### 2023-12-05 manuscript.tex
Update citation format

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[vmargin=1in,hmargin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{pdflscape,booktabs}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage[running]{lineno}
\linenumbers
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\section{The response of trophic interaction networks to multiple
stressors in a marine latitudinal gradient of the Southern
Hemisphere}\label{the-response-of-trophic-interaction-networks-to-multiple-stressors-in-a-marine-latitudinal-gradient-of-the-southern-hemisphere}

\subsection{Running Title: Marine stressors in the Southern
Hemisphere}\label{running-title-marine-stressors-in-the-southern-hemisphere}

Tomás I. Marina\textsuperscript{1}, Leonardo A.
Saravia\textsuperscript{1,2,*}, Iara D. Rodriguez\textsuperscript{3},
Manuela Funes\textsuperscript{4}, Georgina Cordone\textsuperscript{5},
Santiago R. Doyle\textsuperscript{3,6}, Anahí
Silvestro\textsuperscript{6}, David E. Galván\textsuperscript{5},
Susanne Kortsch\textsuperscript{7} \& Fernando Momo\textsuperscript{3,6}

\textsuperscript{1} Centro Austral de Investigaciones Científicas
(CADIC-CONICET), Ushuaia, Argentina;

\textsuperscript{2} Instituto de Ciencias Polares, Ambiente y Recursos
Naturales, Universidad Nacional de Tierra del Fuego (UNTdF), Ushuaia,
Argentina;

\textsuperscript{3} Instituto de Ciencias, Universidad Nacional de
General Sarmiento (UNGS), Los Polvorines, Argentina;

\textsuperscript{4} Instituto de Investigaciones Marinas y Costeras
(IIMyC-CONICET), Mar del Plata, Argentina;

\textsuperscript{5} Centro Para el Estudio de Sistemas Marinos
(CESIMAR-CONICET), Puerto Madryn, Argentina;

\textsuperscript{6} Instituto de Ecología y Desarrollo Sustentable
(INEDES-CONICET-UNLu), Luján, Argentina;

\textsuperscript{7} Tvärminne Zoological Station, University of
Helsinki, Hanko, Finland.

* corresponding author: Leonardo A. Saravia. Centro Austral de
Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina.
\href{mailto:lasaravia@untdf.edu.ar}{\nolinkurl{lasaravia@untdf.edu.ar}}

\subsection{Abstract}\label{abstract}

Ecological networks offer valuable insights into community structure,
key species identification, and ecosystem management for biodiversity
conservation. Understanding how these networks react to environmental
and anthropogenic stressors, especially along geographical gradients, is
of increasing interest. This review presents a pioneering analysis of
stressor responses in marine food webs from the southwest Atlantic to
the Antarctic (45 - 78ºS), encompassing areas such as San Jorge Gulf,
Beagle Channel, Burdwood Bank, Scotia Sea, Potter Cove, and the Weddell
Sea in Antarctica. Our objectives are to: 1) describe the structure of
marine food webs along this axis using a network approach; 2) identify
predominant environmental and anthropogenic stressors affecting each
ecosystem; and 3) summarize observed food web changes and hypothesize on
stressor impacts. Our collaborative team, comprising regional experts
and global authorities on high-latitude marine food webs and stressor
effects, ensures a comprehensive and credible literature review. We
assessed the effects of stressors primarily at the species level, with
notable exceptions like fisheries in San Jorge Gulf. Hypotheses for each
study area were formulated considering: a) stressors; b) impacted
parameters; c) node-level species properties; and d) network-level food
web properties. Global warming emerges as the most common stressor
across the gradient, except in the Beagle Channel and Burdwood Bank,
where alien species introduction and fisheries are more influential,
respectively. We offer specific hypotheses on how warming may affect
food webs. Our findings highlight the benefits of a network approach in
understanding and predicting stressor effects in Southern Hemisphere
marine ecosystems. This approach provides a holistic understanding of
ecological networks, enhances our ability to identify key species and
interactions, and offers insights for ecosystem management and
conservation in the face of various stressors.

Keywords: anthropogenic stressors, environmental stressors, food webs,
latitudinal gradient, Southern Hemisphere

\subsection{1. Introduction}\label{introduction}

The application of a network perspective has emerged as a powerful tool
to tackle the complexity of species interactions, facilitating a better
understanding of the structure and functioning of ecosystems (Belgrano
et al. 2005; Thompson et al. 2012). Trophic networks (or food webs)
allow identifying properties and key species that may be crucial for
ecosystem stability, and hence important for ecosystem management and
biodiversity conservation (Thompson et al. 2012). There is a growing
interest in understanding how ecological networks respond to
environmental and anthropogenic stressors along geographical gradients
(Cirtwill et al. 2015; Bauer et al. 2022). Yet, only a few studies have
described variation in food web structure along latitudinal gradients in
marine ecosystems. The few that have come from the Global North (Wood et
al. 2015; Kortsch et al. 2019; Pecuchet et al. 2022), whereas no
studies, nor meta-analyses, on geographical variation in marine food
webs exist for the Global South (Southern Hemisphere).

Here we review for the first time the state-of-the-art knowledge on
stressor response of marine food webs along the southwest Atlantic to
Antarctic gradient (45 - 78ºS, Figure 1). We focus on proven and
expected changes in food webs driven by stressors in selected areas
along this large-scale latitudinal gradient. We recruited food web
experts from different marine systems of Argentina and the world (see
co-authors' list). Throughout the year 2023 we maintained regular
discussion meetings, typically held every two or three weeks. The aim of
this review is threefold: 1) describe the complexity and structure of
marine food webs along the southwest Atlantic to Antarctic axis from a
network perspective; 2) identify the ongoing environmental and
anthropogenic stressors for each marine ecosystem containing the food
webs; and 3) resume proven food web changes and elaborate hypotheses on
how the identified stressors might affect food web features (e.g.~energy
flow, stability), combining information on node- and network-level
properties. Finally, we suggest which additional data and analyses are
needed to gain insights into the stressors' effects on food web
properties in the southwest Atlantic to Antarctic region.

\subsection{2. The structure of marine food webs in the Southwest
Atlantic - Antarctic
region}\label{the-structure-of-marine-food-webs-in-the-southwest-atlantic---antarctic-region}

Together, the southwest and the Atlantic sector of the Southern Ocean
comprise one of the most biologically productive regions of the world's
oceans (Acha et al. 2004; Latorre et al. 2023). The referred region
extends from San Jorge Gulf (45ºS) in the Patagonian shelf to the
Weddell Sea (78ºS) in the Southern Ocean, and covers a well-connected
oceanic latitudinal gradient (Matano et al. 2010; Guihou et al. 2020).

Throughout this latitudinal gradient, many investigations have been
carried out addressing the trophic ecology of specific species and
prey-predator relationships (Vinuesa and Varisco 2007; Pasotti et al.
2015; Saunders et al. 2019; Riccialdelli et al. 2020), but few studies
have considered the complexity of the ecosystem in terms of a high
resolution of species and their prey-predator interactions (but see
Jacob et al. (2011), Marina et al. (2018), Funes et al. (2022),
López-López et al. (2022), Rodriguez et al. (2022), Marina et al. (in
rev.)). Neglecting this complexity might lead to a misunderstanding of
the structure and functioning of the ecosystems, and ultimately reduce
the ability to predict ecosystem responses to perturbations (Montoya et
al. 2009).

In this review, we consider marine areas in the Southern hemisphere for
which highly-resolved empirical food webs, in terms of species diversity
and trophic interactions, have been previously described. These areas
include: (1) San Jorge Gulf (45 - 47ºS, 65 - 68ºW), (2) Beagle Channel
(\textasciitilde54ºS, 68ºW), (3) Burdwood Bank (\textasciitilde54ºS,
59ºW), (4) Scotia Sea (48 - 58ºS, 50ºW), (5) Potter Cove (62ºS, 58ºW,
Antarctica), and (6) Weddell Sea (74 - 78ºS, 30ºW, Antarctica) (Figure
1). The geographic locations of these marine biomes span from temperate
to Antarctic, and are exposed to both environmental (sea warming,
glacial retreat) and anthropogenic (fishery, pollution) stressors.

\begin{figure}
\centering
\includegraphics[width=6.27in,height=3.528in]{Figures/Map_final.jpg}
\caption{Map of the study areas along the southwest Atlantic - Antarctic
latitudinal gradient. The areas are marked with numbers from one to six.
Smaller areas (1. San Jorge Gulf, 2. Beagle Channel, 3. Burdwood Bank,
and 5. Potter Cove) are shown on the panels to the left, whereas the
larger areas (4. North and South Scotia Sea, and 6. Weddell Sea) are
marked with a red rectangle on the map. The map was drawn using the
`marmap' R package (Pante et al. 2023). Continental contour shapefiles
were obtained from www.ign.gob.ar.}
\end{figure}

San Jorge Gulf is the northernmost study area considered in this review.
It is a partially enclosed basin spanning approximately 34,000 km2 and
\textasciitilde100 m of maximum depths, located from Cabo dos Bahias
(44°55'S) to Cabo tres Puntas (47°06'S) (Figure 1). The shallower
northern and southern ends of the Gulf (with depths \textasciitilde40 m)
present two prominent frontal systems. These systems are the areas of
highest productivity (Glembocki et al. 2015). The Gulf's productivity
supports large invertebrate and vertebrate fisheries (Góngora et al.
2012), as well as marine mammal and seabird populations (Yorio 2009).
The San Jorge Gulf food web comprises 165 nodes and 1015 trophic
interactions with a connectance of 0.04. The percentage of top predators
is 16\%, 78\% of intermediate nodes and 6\% of basal nodes; 60\% of
predators are omnivorous (Table 1). The most connected nodes are: the
Argentine red shrimp Pleoticus muelleri, the squat lobster Gimothea
gregaria, squids (Illex argentinus as dominant species) and Amphipoda.
Notably, these nodes present mid-trophic positions in the food web (3.0,
2.5, 3.6 and 2, respectively) (Funes et al. 2022).

At a southern latitude, the Beagle Channel (54°52'S, 68°8'W) is an
interoceanic passage located at the southernmost tip of South America,
spanning 240 km in length and 5 km in width (\textasciitilde1200 km2),
and a range depth of 0 - 140 m (Figure 1). It features complex
coastlines, varying bathymetry, a prevailing west-to-east circulation
pattern, and a significant longitudinal gradient of glacial freshwater
discharge (Schloss et al. 2023). Bruno et al. (2023) suggests that
locally produced suspended particulate organic matter (mostly composed
by phytoplankton) and organic matter accumulated in the sediments
(macroalgae-originated detritus) are the primary food sources for the
marine ecosystem, as opposed to allochthonous materials. Beagle Channel
food web includes 145 nodes and 1115 trophic interactions with a
connectance of 0.05 (Table 1). The food web is suggested to have a
wasp-waist structure, where the following species play a crucial role in
the dynamics of the ecosystem: Fueguian sprat Sprattus fuegensis,
longtail southern cod Patagonotothen ramsayi, black southern cod P.
tessellata, frogmouth Cottoperca trigloides, and squat lobster Grimothea
gregaria (Riccialdelli et al. 2020). Moreover, the squat lobster has
been identified as a species responsible for linking modules and
connecting the entire food web (Rodriguez et al. 2022).

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II, meaning the shallow submarine
plateau named Burdwood Bank with a 200 m isobath boundary, and a deep
slope that reaches 4000 m in depth, respectively (Administración de
Parques Nacionales 2022) (Figure 1). Physical features in the plateau
are fairly stable, with salinity averaging 34 all year round and
temperature ranging between 4 and 8ºC (Acha et al. 2004). The plateau is
surrounded by steep flanks of up to 4000 m depth, protected by the
Marine Protected Area Namuncurá - Burdwood Bank II (32,000 km2;
\textasciitilde55º - 56ºS, \textasciitilde58º - 62ºW). Intense upwelling
and mixing occur in relation to the slope, entraining deep nutrient-rich
waters into the photic layer (Matano et al. 2019), and resulting in a
fairly homogeneous water column both spatially and temporally (Matano et
al. 2019). The Burdwood Bank food web comprises 379 nodes and 1788
interactions, with a connectance of 0.01, and an asymmetric degree
distribution (i.e.~most of the species have a relatively low number of
interactions and few species concentrate most of them). Almost half of
the consumers are omnivores (0.48), and the network displays a
small-world pattern (Marina et al. in rev.) (Table 1).

The Scotia Sea is a deep-sea basin, delimited by the Drake Passage to
the West and by the island complex of the Scotia Arc to the North, East,
and South, with an approximate extension of 1.5 x 106 km2 and a depth
range of 0 - 3000 m (Murphy et al. 2006) (Figure 1). Its oceanography is
dominated by the Antarctic Circumpolar Current, which is spatially
structured by frontal systems (Whitworth 1980). The South Antarctic
Circumpolar Current Front subdivides the Scotia Sea into two
biogeographic regions: the Northern Scotia Sea is characterized by
higher and more variable temperatures, and the Southern Scotia Sea by
lower and more stable temperatures and influenced by seasonal sea ice
(Raymond 2011). The analysis of the Northern and Southern Scotia Sea
food webs shows that the former is relatively more complex than the
latter: with higher species richness (218 vs 192) and interactions
(10008 vs 7241) and a higher connection overall (0.21 vs 0.20). As
expected from a more complex network, the path length is shorter in the
Northern Scotia Sea food web. In the same sense, the Southern Scotia Sea
network displays a greater proportion of omnivores and a lower mean
trophic level (López-López et al. 2022) (Table 1).

In the Antarctic realm, Potter Cove is a \textasciitilde9 km2 fjord with
a depth range of 0 - 200 m located at 25 de Mayo/King George Island
(62°14'S, 58°38'W, South Shetland Islands) on the West Antarctic
Peninsula (Figure 1). The cove, bordered by the Fourcade Glacier, is
divided into three areas: a) the internal cove, a high
glacier-influenced, soft sediment zone with a 50 m maximum depth; b) the
central cove, a mixed substrate area with low meltwater influence and an
80 m maximum depth; and c) the external cove, ice-free for 60 years with
a 185 m maximum depth and rocky substrate (Jerosch et al. 2018). Potter
Cove's high-latitude location results in variable environmental
conditions due to photoperiod length seasonality. Sea ice often covers
this area in winter (Schloss et al. 2012). With low phytoplankton
biomass, macroalgae, and microphytobenthos are likely the primary food
sources for secondary benthic production (Quartino and Boraso de Zaixso
2008). The Potter Cove food web includes 110 nodes and 649 interactions,
with an asymmetric degree distribution, and a connectance value of 0.05
(Table 1). It presents a modular structure (groups of species interact
more strongly with each other than with species belonging to other
groups), that's positively associated with stability, since perturbation
can be retained within modules (Rodriguez et al. 2022).

Located between 74 and 78ºS, the high Antarctic Weddell Sea shelf spans
approximately 450 km from East to West (Jacob et al. 2011) (Figure 1).
The water depth ranges from 200 to 500 m, with shallower regions being
covered by continental ice that forms the coastline along the eastern
and southern parts of the Weddell Sea. Within this shelf area, exists a
complex three-dimensional benthic habitat characterized by substantial
benthic biomasses and an intermediate to high diversity when compared to
benthic boreal communities (Teixidó et al. 2002). The Weddell Sea food
web exhibits a high level of network complexity, featuring the greatest
number of nodes (490) and trophic interactions (16041) among the
analyzed food webs in this review (Table 1). Its connectance (0.07) and
other structural properties are intermediate compared to the other food
webs. Recently, the interaction strengths of this food web were
estimated, revealing the presence of numerous weak and few strong
interactions, which is consistent with findings in other complex food
webs (Nilsson and McCann 2016). Notably, this asymmetric distribution of
interaction strength is likely to promote community persistence.

\newpage
\footnotesize

Table 1. Complexity and structure properties of the marine food webs
considered in the present review. Refer to Table 2 for definition of
properties. mean TL: mean trophic level. Food webs are ordered by
increasing latitude.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1538}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0839}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0839}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1259}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1259}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0979}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1049}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.2238}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Food web}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Nodes}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Links}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Connectance}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Path length}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{mean TL}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Omnivory}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Reference}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
San Jorge Gulf & 165 & 1015 & 0.04 & 2.17 & 3.02 & 0.63 & Funes et al.
(2022) \\
Beagle Channel & 145 & 1115 & 0.05 & 2.12 & 2.37 & 0.55 & Rodriguez et
al. (2022) \\
Burdwood Bank & 379 & 1788 & 0.01 & 2.99 & 2.52 & 0.49 & Marina et al.
(in rev.) \\
Potter Cove & 110 & 649 & 0.05 & 2.33 & 2.22 & 0.46 & Marina et al.
(2018); Rodriguez et al. (2022) \\
N Scotia Sea & 218 & 10008 & 0.21 & 1.87 & 3.29 & 0.73 & López-López et
al. (2022) \\
S Scotia Sea & 192 & 7241 & 0.20 & 1.90 & 3.21 & 0.71 & López-López et
al. (2022) \\
Weddell Sea & 490 & 16041 & 0.07 & 2.19 & 2.62 & 0.51 & Jacob et al.
(2011) \\
\end{longtable}

\normalsize

\subsection{3. Environmental and anthropogenic stressors in the
Southwest Atlantic - Antarctic
region}\label{environmental-and-anthropogenic-stressors-in-the-southwest-atlantic---antarctic-region}

A stressor is any environmental or anthropogenic variable that causes a
quantifiable change, irrespective of its direction (increase or
decrease), in a biological response (Orr et al. 2020). In the light of
this, the southwest Atlantic - Antarctic marine biota has been and is
currently subjected to a variety of stressors (e.g.~sea warming, glacial
retreat, ice changes, acidification, species invasion, fisheries, and
contamination). Although it is well-known that multiple stressors act in
concert at any given time (e.g.~warming and fishery; acidification and
contaminants) (Gutt et al. 2021), to date stressor assessments have been
performed individually. Moreover, the potential for interactive effects
of two or more stressors (synergy or antagonism) (Côté et al. 2016) is
almost unknown for the region (Rowlands et al. 2021).

In the following subsections, we describe the main environmental and
anthropogenic stressors, and the species (or trophic species) and
parameters affected reported for the marine ecosystems that contain the
food webs considered in this review. In Table S2 we provide an
exhaustive list of all stressors affecting species inhabiting each area,
considering: the type of stressor, species and parameter affected, and
locality.

\subsubsection{3.1 San Jorge Gulf}\label{san-jorge-gulf}

San Jorge Gulf experienced several environmental and anthropogenic
stressors (Table S2). Trawl fisheries discard several species and add
new trophic interactions to the food web which resulted in a decrease in
trait variability and the stability of the system (Rincón-Díaz et al.
2021; Funes et al. 2022). Moreover, it changed the availability of prey
to several predators. For example, Merluccius hubbsi one of the main
bycatch species, became prey item to non-diving seabirds, like the kelp
gull Larus dominicanus (González-Zevallos and Yorio 2006) and reef
fishes (Funes et al. 2019). Although juveniles of M. hubbsi are largely
the main bycatch item, 29 other cartilaginous and 69 bony local fish
species were also registered as incidental catch between 2005 and 2014
(Bovcon et al. 2013; Ruibal Nuñez 2020). This level of impact triggered
a shift in the functional diversity of the assemblages homogenizing the
trophic function of fishes (Rincón-Díaz et al. 2021). Other functional
changes were a decrease in the maximum sizes of individual fish,
together with a drop in elasmobranchs biomass and an increase in
crustaceans biomass (Funes 2020). The significant increase in crustacean
biomass was mostly due to the increase in Pleoticus muelleri and
Grimothea gregaria populations (Funes 2020). These species rapidly
became the most important prey for the most abundant fishes in the area:
M. hubbsi and Mustelus schmitti (Pasti et al. 2021). However, the
above-mentioned effects of trawl fisheries on the structure and function
of the San Jorge Gulf community may have changed again, since the
fisheries ceased activity in 2015 (Annex I, Resolution CFP No 7/2018),
remaining a small trawling artisanal fishery.

Sea warming is another important environmental stressor in the San Jorge
Gulf, because of the southward shifts of northern fish populations to
the Gulf (Galván et al. 2022). The San Jorge Gulf is especially prone to
be affected by climate-driven shifts in species ranges, because it is
located in the ecotone between two biogeographic provinces, the
Argentine (30°S - 44°S) and the Magellanic (43°S - 55°S) (Balech and
Ehrlich 2008). In addition to this tropicalization from northern fish
populations, alien species are also documented to affect the demersal
assemblage (fish and macroinvertebrates) in the Gulf (Galván et al.
2022). Finally, San Jorge Gulf is exposed to urban and industrial
pollution due to an oil monobuoy from which oil manipulation and general
oil transport along the Patagonian coast registered several oil spills
and chronic oil discharges (García-Borboroglu et al. 2008). Other marine
systems impacted by oil spill showed an important decrease in marine
bird populations (Irons et al. 2000), seals and macroalgae (e.g. Paine
et al. (1996)), with examples of cascading effects throughout the food
web (Peterson 2001).

\subsubsection{3.2 Beagle Channel}\label{beagle-channel}

A main anthropogenic stressor in the Beagle Channel is the introduction
of invasive species (Table S2). Salmonidae were introduced to Tierra del
Fuego in the 1930s. Especially Chinook salmon Oncorhynchus tshawytscha
causes concern. The detection of Chinook salmon in Tierra del Fuego
dates back to April 2006, and its population has been expanding since
then (Nardi et al. 2019). Being a top predator Chinook salmon can
compete with several native species in the Beagle Channel (Correa and
Gross 2008), and prey over native species such as Notothenioids
(Eleginops maclovinus, Patagonotothen tessellata, P. cornucola, P. sima,
Paranotothenia magellanica, Harpagifer bispinis), the Atherinidae
Odontesthes smitti and O. nigricans, the Fuegian sprat Sprattus
fuegensis, and larvae of king crabs (Lithodes antarcticus and Paralomis
granulosa) (Fernández et al. 2010). Ciancio et al. (2008) observed that
Chinook salmon in the Southern Patagonian Shelf Ecosystem Area primarily
feed on sprats and display trophic levels comparable to those of
intermediate-sized fish and cephalopod predators species, showing
significant dietary overlap with Magellanic penguins (Spheniscus
magellanicus). Another potential competitor for Chinook salmon in the
Beagle Channel is the Commerson's dolphin (Cephalorhynchus commersonii),
which shares a similar diet (Riccialdelli et al. 2013).

In the Beagle Channel, other anthropogenic stressors include
contaminants like metals, perfluorinated compounds, hydrocarbons, and
microplastics found in animal tissue and sediments. Some studies
reported varying heavy metal levels in sea mussel tissues and sediments,
indicating bioavailability differences (Duarte et al. 2011). Other
research found similar carbon and carbohydrate levels in Ushuaia Bay's
surface sediments to those in hypertrophic ecosystems, along with
hydrocarbons and heavy metals linked to port and industrial activities
(Gil et al. 2011). Commendatore et al. (2012) identified three distinct
hydrocarbon accumulation zones in the Bay. Llorca et al. (2012) found
high levels of perfluorinated compounds in algae, fish, and guano.
Ferreira et al. (2021) showed that black southern cod in Ushuaia Bay may
be exposed to endocrine-disrupting compounds from urban and industrial
pollution. Pérez et al. (2020) and Ojeda et al. (2021) found
microplastics in M. edulis chilensis and Nacella magellanica,
respectively. These studies link pollutants to Ushuaia (54°48′S 68°18′W)
and its port and industrial activities. We expect the Beagle Channel
area closest to the city to be most affected. Contaminants can spread
through the food web, magnifying the risk to higher-level organisms.
Fioramonti et al. (2022) studied mercury transfer in three southwest
Atlantic Ocean food webs and found biodilution in the Beagle Channel and
Atlantic coast of Tierra del Fuego webs. However, mercury concentrations
rose with benthivory in these webs, and higher mercury was found in
phytoplankton and Grimothea gregaria. Squat lobsters connect pelagic and
benthic habitats, so any disruption to them could disrupt the food web
(Rodriguez et al. 2022). Dodino et al. (2022) found the highest mercury
levels in Magellanic penguin feathers from offshore colonies in Tierra
del Fuego. Recently, Ushuaia's kelp forests have seen a decrease in
biodiversity and changes in macroalgae composition due to urban
pollution (Kaminsky et al. in prep.).

\subsubsection{3.3 Burdwood Bank}\label{burdwood-bank}

The stressors reported for the ecosystem of the Marine Protected Areas
Namuncurá within Burdwood Bank I and II are mostly anthropogenic; there
is a lack of studies focusing on the environmental stressors that are
known to be occurring at a regional scale (e.g.~sea surface warming,
Franco et al. (2020)) (Table S2). Several fisheries targeting demersal
fishes operate in the vicinity and within the ecosystem of the Marine
Protected Areas (i.e.~Marine National Reserve management category). The
Patagonian fisheries on the toothfish Dissostichus eleginoides has
gained prominence in recent years (Allega et al. 2020; Gorini et al.
2021). Although these are regulated by the Argentinean government,
incidental catches do occur, where demersal fishes of the genera
Coelorinchus and Macrourus, seabirds and benthic invertebrates (30+
taxa) are the most common bycatches (Gaitán and Marí 2016; Martínez et
al. 2022). Noteworthy, among the invertebrates caught, 8 species are
indicator taxa of vulnerable marine ecosystems (Gaitán and Marí 2016;
Schejter and Albano 2021). Independent assessments of these bycatches
suggest no significant impact on the communities (Gaitán and Marí 2016;
Martínez et al. 2022). Besides bycatch, species of seabirds are being
affected by the fishery due to discards, altering its diet; the most
frequently encountered species are Black-browed albatross Thalassarche
melanophris, Southern giant petrel Macronectes giganteus, Cape petrel
Daption capense, Southern royal albatross Diomedea epomophora, Northern
giant petrel M. halli, and White-chinned petrel Procellaria
aequinoctialis (Tamini et al. 2023). Nevertheless, there is a lack of
knowledge considering the species' role in the ecosystem and the
potential joint effect of both target fishes and bycatch in a broader
food web framework.

Anthropogenic contaminants such as microplastics and mercury are present
in the water column of the Burdwood Bank ecosystem (Cossi et al. 2021;
Fioramonti et al. 2022; Di Mauro et al. 2022). Microplastics are
distributed all along the water column, from surface to deep waters
(3-2450 m) (Di Mauro et al. 2022). More importantly, microplastics were
found in soft tissues of benthic macroinvertebrates (sea stars Henricia
obesa and Odontaster penicillatus) and benthopelagic fishes
(Patagonotothen guntheri and P. ramsayi), which not only incorporated
the contaminant from the environment through their filter-feeding system
but could also get it indirectly from prey organisms already containing
plastics in their tissues (Cossi et al. 2021). Notably, one of the
contaminated species, the long tail southern cod Patagonotothen ramsayi,
is part of the core group of species that drive the ecosystem through
the suggested wasp-waist control (Riccialdelli et al. 2020). Mercury
transfer and biomagnification are ongoing processes in the ecosystem,
which are occurring at a greater pace than near coastal areas (Beagle
Channel) (Fioramonti et al. 2022). It's important to note that the
Fuegian sprat Sprattus fuegensis, a pelagic fish with a mid-trophic
level in the food web, presented the highest levels of mercury
(Fioramonti et al. 2022). Considering the wasp-waist control of the
Fuegian sprat in the food web (Riccialdelli et al. 2020), a rapid and
widespread contamination to the top predators is expected (Fioramonti et
al. 2022).

In recent years, licenses for seismic studies and exploration of
hydrocarbon resources have been granted all along the northwestern limit
of the Marine Protected Areas (Secretaría de Gobierno de Energía Res. Nº
65/2018). Although effects from these types of surveys on marine mammals
and seabirds are well-known for other regions of the world (Nowacek et
al. 2015), there is no particular knowledge for this ecosystem. However,
several documents warn of the potential negative effects this may have
on the species inhabiting the Burdwood Bank area (Allega et al. 2019; de
Haro et al. 2022).

Despite evidence of warming at surface, mid-water and bottom layers (100
m) in Burdwood Bank (Franco et al. 2020), specific studies on the
oceanographic aspects of the system are lacking. In this sense, there is
a big question mark about the environmental stressors impact on the
species and trophic interactions in Burdwood Bank.

\subsubsection{3.4 Scotia Sea}\label{scotia-sea}

The Scotia Sea is a vast and heterogenous oceanic region, where
especially the areas around South Georgia island represents an area of
interest, here referred to as `Northern Scotia Sea'. The majority of
studies analyzing the stressors' effects come from this area.

The Scotia Sea has experienced one of the largest levels of warming
within the polar regions (Whitehouse et al. 2008). Together with the
Southern Annular Mode anomalies this has caused a long-term decrease in
krill abundance; more pronounced in the northern than in the southern
Scotia Sea (Murphy et al. 2007). Over the past 90 years, the krill also
showed an increase in mean body length (Atkinson et al. 2019), which may
also alter predator-prey interactions and allow reaching cooler feeding
grounds near the seabed, with the potential to link krill to unexpected
predators (Schmidt et al. 2011). Another consequence of the mentioned
environmental stressors, is a change in krill distribution due to a
southward contraction (Atkinson et al. 2019). It is noteworthy here,
that the distributional shift has not affected all species inhabiting
the Northern Scotia Sea, and most abundant calanoid copepods have
maintained their distribution (Tarling et al. 2018). Considering the
above, food web models of the Northern Scotia Sea have suggested shifts
in the structure and functioning of the food web, from krill-based to
non-krill-based, where myctophid fishes and squid are playing important
roles (Saunders et al. 2019).

The principal anthropogenic stressor of the Scotia Sea is commercial
fisheries. Krill fishery not only stresses the targeted species, but
also the many predators that depend on it as a food source (Hilborn et
al. 2017). Yet, data currently available from monitoring of krill and
its predators remain insufficient, hence identifying the potential
fishery impacts on the ecosystem is difficult (Trathan et al. 2021).
Apart from krill fishery, two other commercial fisheries operate in the
Scotia Sea, targeting Patagonian toothfish species Dissostichus
eleginoides and D. mawsoni. The D. eleginoides stock is linked to the
stock at South Georgia (`Northern Scotia Sea') (Collins et al. 2010),
while the D. mawsoni stock is linked to the Antarctic continental shelf
(`Southern Scotia Sea') (Soeffker et al. 2022). Despite the fact that
the mentioned fisheries are certified by the Marine Stewardship Council
standards, Trathan (2023) identified several concerns regarding aspects
of fisheries itself, other species' population status (i.e.~recovery of
baleen whales), and environmental stressors (i.e.~sea warming), which
could have synergistic effects on the food web structure and functioning
that are still unknown.

Mercury transfer and biomagnification are current processes occurring in
the Scotia Sea, where the total concentration of contaminants increase
with trophic level and are highest in notothenioid and myctophid fishes
(e.g.~Dissostichus eleginoides, Gymnoscopelus nicholsi), and seabirds
(Seco et al. 2021). During years of low Antarctic krill abundance,
predators must deal with both the stress of reduced prey availability
and the concurrent rise in mercury exposure (Seco et al. 2021).

\subsubsection{3.5 Potter Cove
(Antarctica)}\label{potter-cove-antarctica}

Regional warming in the last half century has been one of the main
factors driving changes in Potter Cove (Western Antarctic Peninsula)
(Chown et al. 2022). Sudden environmental changes have occurred, such as
sea surface temperature increase, salinity decrease, suspended
particulate matter loading and chlorophyll-a increase, all linked to
climatic cycles (Southern Annular Mode and El Niño Southern Oscillation)
(Schloss et al. 2012). Particularly, sea warming has produced drastic
environmental and biological transformations (e.g.~shifts in dominance
of benthic community) in the Potter Cove ecosystem (Schloss et al. 2012;
Quartino et al. 2013; Sahade et al. 2015), a system highly dependent on
sea-ice dynamics (Table S2). In Potter Cove, total sea ice cover has
decreased since 1991 (Schloss et al. 2012). Changes in the annual timing
of landfast ice formation and breakup of the sea ice cover has multiple
effects on species in the food web (Michel et al. 2019). Warmer winters
and springs result in earlier sea-ice melt, causing an abrupt increase
in the light available benthic primary producers (Deregibus et al.
2020). Sea ice also mediates physical disturbances to the benthos by
influencing sedimentation and iceberg scouring. These factors affect the
production of macroalgae, albeit in opposite ways (Deregibus et al.
2017), and microphytobenthos (Hoffmann et al. 2019). On the other hand,
sea ice is an important habitat for diatoms and its associated
consumers, including copepods and krill (Flores et al. 2012), and thus
important for bentho-pelagic nutrient and carbon cycling during winter.
Additionally, a decrease in winter sea ice cover produces an increase in
physical perturbation on benthic shallow communities in coastal shallows
due to ice scouring (Deregibus et al. 2017). The glacier surrounding
Potter Cove has been receding at an increasing rate since 1950 (Rückamp
et al. 2011), which has caused a massive discharge of sediment-laden
meltwater (Meredith et al. 2018). Large quantities of suspended
particles affect growth, survival and reproduction of benthic species.
This had led to a major shift in the benthic community structure, from a
filter feeders--ascidian domination to a mixed assemblage with
scavengers and opportunistic species (Sahade et al. 2015), and the
metabolic balance in benthos went from net autotrophy to heterotrophy
(Braeckman et al. 2021). Additionally, massive stranding events of the
tunicate Salpa thompsoni and the euphausiid Euphausia superba (krill)
linked to the presence of glacial meltwater have been reported (Fuentes
et al. 2016). Rising temperatures leading to ice and glacial melting has
also substantial impacts on pelagic primary productivity, since it
reduces water salinity, affecting water column stratification, light
penetration and nutrient availability for photosynthesis (Schloss et al.
2012). In Potter Cove, changes in biomass of most phytoplankton species
have been observed under heat wave conditions, resulting in a shift from
a microplankton to a nanoplankton dominated community (Antoni et al.
2020; Latorre et al. 2023). This means that in areas strongly affected
by glacier melt, the planktonic food web is dominated by the microbial
loop (ciliates and heterotrophic dinoflagellates preys upon
nanophytoplankton, which are sequentially available prey for small
omnivorous copepods), instead of being predominantly herbivorous (Garcia
et al. 2016, 2019). In addition, phytoplankton species under these
warming conditions showed a decrease in metabolic rates and in the
quality of the fatty acids composition (Latorre et al. 2023).

\subsubsection{3.6 Weddell Sea
(Antarctica)}\label{weddell-sea-antarctica}

In the Weddell Sea several stressors have already been triggered by
global warming effects (Table S2), from which spatial and temporal
reduction in sea ice is suggested to be driving changes in pelagic and
benthic communities (Constable et al. 2014; Gutt et al. 2021). Sea ice
extent has reached new record lows since the satellite era began in
1978. Sea warming has been substantial in recent years: in 2017 the mean
temperature for February reached 1.45°C, the highest monthly mean ever
recorded (0.56°C above the climatological mean) (Turner et al. 2020). In
this context, it has been suggested that Antarctic krill Euphausia
superba has already declined as a result of productivity changes caused
by sea ice declines (Atkinson et al. 2004). Declining sea ice cover
allows increased access to krill by predators (Kawaguchi et al. 2009),
which further contributes to decreased krill abundances. This in turn
can result in reduced carbon export due to decreased fecal pellets from
krill (Pauli et al. 2021).

On the other hand, sea ice loss increases light availability, triggering
primary production by phytoplankton in the short-term (Pineda-Metz et
al. 2020). Furthermore, a warmer and more stable water column with a
shallow upper mixed layer, is expected to enhance the dominance of
cryptophytes over diatoms which may favor salp populations over krill
populations; thereby reducing the magnitude of energy transfer to higher
trophic levels and the seabed (Isla 2023).

Higher trophic level species are expected to respond differently to
global warming in the Weddell Sea. In this regard, Notothenioid fishes
(e.g.~Pleuragramma antarcticum) face multiple stressors (sea warming,
sea ice decline, ocean acidification) that threaten their survival.
While some species exhibit physiological plasticity to compensate for
increased oxygen demand, most notothenioid fishes are stenothermal and
unable to adjust their metabolic functioning (Mintenbeck et al. 2012).
It is anticipated that the thermal habitat preferred by the Antarctic
toothfish Dissostichus mawsoni could undergo contraction in the coming
three decades (Constable et al. 2014), further emphasizing the potential
impact of global warming effects on these species. Marine mammals,
Weddell and crabeater seals (Leptonychotes weddellii and Lobodon
carcinophagus), exhibit varying levels of sensitivity. The latter
exhibits a preference for breeding in close proximity to krill swarms,
and thus is particularly susceptible to reductions in sea ice
concentrations and sea surface temperatures. Rescued sea ice can
significantly affect their post- and future breeding foraging success
(Wege et al. 2021). In addition, the Antarctic petrel Thalassoica
antarctica, which is one of the most abundant seabirds in the area, is a
sea ice dependent forager and breeder (Orgeira et al. 2021); therefore
the loss of this habitat will degrade its foraging success and
reproductive output over time. The snow petrel Pagodroma nivea also
relies heavily on sea ice for foraging and its colonies in the Weddell
Sea may be affected. The emperor penguin Aptenodytes forsteri uses sea
ice for breeding; any reductions could constrain its populations, as
evidenced by recent observations in the Bellingshausen Sea region
(Fretwell et al. 2023). The Arctic tern Sterna paradisaea spends the
summer in the Weddell Sea exploiting krill swarms under receding ice
edges, then declines in ice cover may reduce prey availability and
quality of molting habitats. Continued warming is thus expected to
gradually erode the abundance and distribution of these `primary
species' that dominate the avifauna of the Weddell Sea through loss of
critical sea ice habitat (Orgeira et al. 2021). Variations in sea ice
extent have implications for great whales, such as the Humpback whale
Megaptera novaeangliae, since their primary prey resource is the
Antarctic krill (Braithwaite et al. 2015). Decreased winter ice coverage
results in reduced suitable habitat and lower krill abundance
(Braithwaite et al. 2015), with flow-on effects for whale body condition
observed historically in harvest data. With climate projections
indicating ongoing sea ice losses, further threats to critical krill
populations pose risks to the energy intake and reproductive success and
long-term viability of humpback whale populations dependent on
consistent Antarctic feeding (Pallin et al. 2023).

Iceberg scouring is a major factor in the high biodiversity of benthic
communities in the Weddell Sea (Gutt and Starmans 2001). Even at 600 m
depths, iceberg scouring has a strong effect on the benthic environment,
disrupting the upper layers of the seabed and removing macrofauna. This
patchy disturbance and distribution pattern occurs roughly every 200
square meters on the Antarctic continental shelf. Global warming is
predicted to raise iceberg scouring frequency (Gutt 2001; Smale et al.
2008), disrupting the environment (Smale and Barnes 2008). Gutt et al.
(2015) expects tipping points are significant due to these effects and
lack of knowledge.

\subsection{4. From species' stressors to food web
effects}\label{from-species-stressors-to-food-web-effects}

A major challenge in contemporary ecology lies in predicting the effects
of stressors on complex multispecies systems, such as food webs. Network
analysis has proved to be a powerful tool to tackle this issue, since it
can capture the effects of individual and multiple stressors on
communities and ecosystems (Montoya et al. 2009; O'Gorman et al. 2012;
Bruder et al. 2019).

Environmental and anthropogenic stressor effects in the southwest and
the Atlantic sector of the Southern Ocean have been mostly assessed
individually and at the organism and/or population, i.e., at the node
level (Table S2), with one exception: the effect of fisheries in the San
Jorge Gulf food web (see section below for more details). To address the
most plausible stressors effects on the selected food webs, given the
current information, we built hypotheses for each study area. To this
aim, we developed a theoretical framework considering the following: a)
stressor(s), b) parameter(s) affected, c) node-level properties of the
affected species, and d) network-level properties of the food web.

We considered that a stressor will affect one of the following species'
characteristics or parameters: metabolism, biomass, distribution, and
diet (Figure 2). `Metabolism' refers to any change related to metabolic
rate, such as reproduction, hatching, larval development, growth and
mortality, and contamination due to pollutants (e.g.~growth effect in
filter-feeders due to sediment in water column in Potter Cove; endocrine
disruption in fish due to urban pollutants in Beagle Channel). `Biomass'
indicates an effect at the population level, where the density/abundance
is being impacted (e.g.~abundance decreases in macrobenthos due to
iceberg scouring in Weddell Sea). `Distribution' entails a change at the
population level in the geographic space occupied by a species,
e.g.~southward contraction of Antarctic krill due to sea warming of the
Scotia Sea. `Diet' includes alterations in the prey items of a species
at the population level, e.g, due to prey switching, having a direct
effect on the structure of the food web, e.g.~new prey item (discards)
for seabirds due to fishery activities in Burdwood Bank. Next, we
considered node- and network-level properties relevant to the
hypothesized stressor effects on the food webs, and which have been
previously calculated for the studied food webs (Table 2). At the
node-level, we included: a) degree, b) trophic position, c) omnivory
index, and d) relative abundance (see Table S1 in Supporting Information
for properties of stressed nodes). At the network-level, we considered:
a) connectance, b) path length, c) mean trophic level, and d) omnivory
(Table 1).

\begin{figure}
\centering
\includegraphics[width=6.27in,height=3.528in]{Figures/Figure2_StressorDiagram.png}
\caption{Conceptual diagram: from species' stressors to food web
effects. See text for explanation.}
\end{figure}

\newpage
\scriptsize

Table 2. Node and network-level properties used to build hypotheses on
the stressors' effects.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1963}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2617}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3925}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1495}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Property}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Definition}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Relevance for stressor effects}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Reference}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\emph{Node-level} & & & \\
Degree & Number of feeding interactions in which the species
participates as prey and/or predator. & Perturbations to high-degree
species may have more significant effects on the food web robustness to
perturbations than low-degree species. & Dunne et al. (2002b); Jordán et
al. (2007) \\
Trophic position & Place in the food web relative to the basal resources
that support the community. Classifies species in: basal, intermediate
and top. & Perturbations on basal resources, intermediate species and
top predators are expected to have large effects on the rest of their
communities if ecosystem control is bottom-up, wasp-waist or top-down,
respectively. & Williams and Martinez (2000); Thompson et al. (2007) \\
Omnivory & Consumer resource use across trophic levels. & High-omnivore
species (generalists) are more flexible than low-omnivore species to
diet changes. & Thompson et al. (2007) \\
Relative abundance & Species' density in proportion to the other species
of the food web. & Perturbations on abundant (dominant) species are
expected to have large effects on the stability and energy flux of the
food and ecosystem, respectively. & Nilsson and McCann (2016) \\
\emph{Network-level} & & & \\
Conectance & Proportion of actual interactions among possible ones. &
Estimator of community sensitivity to stressors. High connectance gives
resistance and resilience to the food web. & Dunne et al. (2002a) \\
Path length & Average distance, accounted by the number of interactions,
between any pair of species. & Short distances enhance rapid and broad
propagation of perturbations. & Albert and Barabási (2002) \\
Mean TL & Average of all species' trophic position contained in the food
web. & Influences the magnitude and efficiency of trophic transfer. A
higher mean food chain length reflects increased energy availability and
productivity. & Duffy et al. (2007); Olivier et al. (2019) \\
Omnivory & Proportion of species that feed at different trophic levels.
& It provides trophic flexibility to an ecosystem. Reduces probability
of trophic cascades. & Kratina et al. (2012) \\
\end{longtable}

\normalsize

\subsubsection{4.1 Main stressor effects in food webs in a southwest
Atlantic - Antarctic
gradient}\label{main-stressor-effects-in-food-webs-in-a-southwest-atlantic---antarctic-gradient}

The most common stressor reported along the southwest Atlantic -
Antarctic gradient is global warming, except for Beagle Channel and
Burdwood Bank, which are more influenced by the introduction of an alien
species and fisheries, respectively (section 3, Table 3). The main
characteristics of global warming in the region, and the most plausible
drivers of change, are: sea warming, glacial retreat, elevated sediment
input in the water column, and reduction of the sea ice extent. These
drivers act in different ways and magnitudes in the studied locations
along the latitudinal gradient. Despite emphasizing global warming in
this section, this does not mean that no other stressors act or interact
with global warming in the study systems, potentially buffering the
overall effect on the food web (e.g.~sea warming and fishery in San
Jorge Gulf). Climate change has led to several well-documented impacts
on marine species regarding distributional shifts induced by warming of
marine currents (Wu et al. 2012; Poloczanska et al. 2013; Vergés et al.
2019). Furthermore, warmer temperatures increase species metabolic rates
(Brown et al. 2004). Changes in metabolic rates can subsequently
translate into shifts in species traits (body size, Vucic-Pestic et al.
(2011); Klein et al. (2018){]}, biomass (Perry et al. 2020) and
distribution (Kortsch et al. 2015). Alterations in the species body size
and distributions have ripple effects on feeding interactions, for
example, it can introduce new feeding interactions (Vergés et al. 2014;
Pecuchet et al. 2020), modify existing ones, and shorten energy pathways
(Bartley et al. 2019; O'Gorman et al. 2019), and reduce trophic
efficiencies (Vucic-Pestic et al. 2011).

In recent years, several new fish (Galván et al. 2022) and
macroinvertebrates species (Vinuesa 2005; López-Gappa 2022) were
registered in Patagonia, mostly in San Jorge Gulf in relation to the
southward range shift of warm-temperate species. This distributional
change is driven by the tropicalization of temperate waters caused by
sea warming (Vergés et al. 2014; Vergés et al. 2019). Because of its
location in the ecotone between two biogeographic provinces, the
Argentine (30°S - 44°S) and the Magellanic (43°S - 55°S) (Balech and
Ehrlich 2008), the San Jorge Gulf is prone to changes in species
composition. We hypothesize that sea warming will alter the food web
structure topologically, by increasing the number of species and
interactions. Newcomers are, in general, mid-trophic level species with
generalist diets, hence an increase in food web connectance may be
expected (Bartley et al. 2019). In another temperate ecosystem, an
increase in the number of fish species, led to an increase in functional
diversity and predation rate (Sgarlatta 2023); consequences that may
also be expected in San Jorge Gulf. Given the short path length of the
San Jorge Gulf food web, the disturbances from the listed stressors are
expected to spread to many species of the food web (Table 3). However,
it has to be acknowledged that the increase in functional diversity
driven by the range expansion of warm-temperate species is contrary to
the process of homogenization and loss of functional diversity in the
area driven by trawl fisheries (Rincón-Díaz et al. 2021).

In the middle of the latitudinal gradient (considered in this study),
the Scotia Sea has experienced one of the largest levels of sea warming
of any polar region (Whitehouse et al. 2008; Atkinson et al. 2019).
López-López et al. (2022) suggested that the southward distributional
shift of generalist predators from the northern towards southern Scotia
Sea increases network connectance of the latter, while decreasing its
modularity. The lower modularity may increase the probability of
perturbations spreading through the network (Stouffer and Bascompte
2011), which may be offset by increased connectance enhancing robustness
to species loss (Dunne et al. 2002b). In the northern Scotia Sea around
South Georgia Island, we suggest that the declining krill biomass driven
by sea warming (Atkinson et al. 2019), ocean acidification and pollution
synergy (Rowlands et al. 2021), will reduce the energy transfer to top
predators like seabirds and marine mammals. However, this may be
buffered since the dominant copepod species have maintained their
distribution (Tarling et al. 2018), but most importantly, showed an
abundance increase in recent decades likely due to reduced predation and
competition for food (Ward et al. 2018). All this is significant for the
structure of the food web given the central role of krill and copepods,
characterized by high degree and mid-trophic position (Table S1).
Overall, the food web's inherent resilience, marked by high connectance
and omnivory, added to the potential compensation for the krill decrease
due to a copepod increase, may buffer against structural changes (Table
3).

In Potter Cove, a fjord-like Antarctic ecosystem, the impacts of climate
change affect many species within the food web. Potter Cove has recently
experienced frequent events of marine heatwaves, i.e.~prolonged periods
of anomalously high sea surface temperatures (Oliver et al. 2018;
Latorre et al. 2023). This has led to decreases in biomasses of
different planktonic functional groups (Garcia et al. 2019; Latorre et
al. 2023). Given the relatively low abundance of phytoplankton,
zooplankton's low degree, and the modular configuration of the food web,
we hypothesize that changes in these nodes, due to increased sea
temperatures, will be retained in the basal pelagic compartment of the
food web and will not expand to higher trophic levels. Benthic primary
producers in Potter Cove are being influence by the decrease in winter
sea ice cover (higher light availability), the increased levels of
sediments in the water column due to glacial melt run-off (lower light
penetration) and the newly free-ice areas available for colonization
associated to glacier retreat. The overall local effect of climate
change on macroalgae is a net increase in their production (Deregibus et
al. 2023). It has been proposed that larger diversity in primary
producers can support a more diverse food web with more specialized
consumers (Iken et al. 2023). We expect to see differential effects of
climate change on hard and soft bottom associated food webs (Cordone et
al. 2020). Given the high relative abundance and the high degree of the
macroalgae functional group (Table S1), we expect a longer hard bottom
food web, wider consumer trophic niches, and that it will become more
stable as sea ice cover decreases and the glacier retreats due to global
warming. In soft bottom areas of the Cove, multiple food web nodes are
being affected by ongoing warming effects: decrease in net primary
production of benthic microalgae (Hoffmann et al. 2019), and changes in
the benthic community biomass, distribution and composition (Sahade et
al. 2015; Pasotti et al. 2015). Given that the Potter Cove food web's
present low connectance and omnivory, we suggest fragility and potential
trophic cascade effects (Marina et al. 2018) with pronounced changes in
energy fluxes (Table 3).

The southernmost location of the latitudinal gradient is the Weddell
Sea, where the main effect of global warming is the decrease in sea ice
extent, with reported anomalies in the past summer seasons (Fretwell et
al. 2023). Declining sea ice extent has reduced the abundance of krill
(Atkinson et al. 2004; Flores et al. 2012), and produced an increase in
phytoplankton productivity (Pinkerton et al. 2021; Isla 2023), altering
the plankton community structure, and benefiting cryptophytes over
diatoms (Lin et al. 2021). Moreover, habitat loss from sea ice decline
will reduce the foraging success and breeding sites of seabirds (e.g.,
snow petrel Pagodroma nivea and emperor penguin Aptenodytes forsteri),
decreasing their biomassess and modifying their distributions. The
projected rise in iceberg scouring is expected to significantly alter
the biomass and community structure of macrobenthos, which in turn will
impact mid-trophic level predators such as demersal fish (Gutt 2001;
Mintenbeck et al. 2012). While we do not anticipate large-scale
topological changes in the food web, local extinctions could lead to
such changes, particularly affecting benthic species (Gutt and
Piepenburg 2003). Given that the impacted species---whether individually
like krill, or collectively like macrobenthos---present a mid-trophic
position, high biomass and high degree (Table S1), we hypothesize that
significant shifts in energy fluxes will occur (Table 3). Additionally,
the food web's low proportion of omnivores suggests reduced system
resilience (Table 1), increasing the likelihood of regime changes (Gutt
et al. 2015).

The two ecosystems in the subantarctic region, the Beagle Channel and
Burdwood Bank, are more affected by other anthropogenic stressors than
warming. Although these areas are being impacted by sea warming,
potentially affecting vertebrate and invertebrate species (Franco et al.
2020), to date no studies exist for Beagle Channel and Burdwood Bank
ecosystems. In the Beagle Channel, the introduction of chinook salmon, a
non-native species, poses a significant risk to the existing food web
(Fernández et al. 2010). We hypothesize that chinook salmon's predation
on Fuegian sprat and black southern cod will disrupt the established
patterns of interaction within the food web. Both of these prey species
are crucial for food web dynamics due to their mid-trophic positions and
relatively high abundance (Table S1). Moreover, we expect that changes
in the black southern cod population will have a more significant impact
on the food web than changes in the Fuegian sprat population, as the
black southern cod has a higher degree (Table S1). Overall, these
disruptions could have far-reaching effects on the ecosystem. This is
particularly concerning given the short path length of the food web,
which means that changes can quickly propagate through the system,
affecting many species and potentially destabilizing the entire network.
This phenomenon is further heightened by the ecosystem's inherent
vulnerability to changes at mid-trophic levels, often referred to as
wasp-waist control (Table 3). In the Burdwood Bank region, fishing
activities may be the main stressor causing shifts in the food web
(Table S2). We hypothesize that a combination of factors will
destabilize the already fragile ecosystem, characterized by low
connectance and low omnivory (Table 1). These factors include a decline
in the biomass of the Patagonian toothfish -a key, highly-connected
species (Table S1)- as well as smaller changes in the biomass of four
mid-level fish species, five top-level seabird species, and over 30
types of benthic macroinvertebrates due to bycatch (Gaitán and Marí
2016; Martínez et al. 2022; Tamini et al. 2023). Additionally,
alterations in the diets of six seabird species, caused by discarded
catch (Tamini et al. 2023), are expected to disrupt the energy flow and
further reduce the stability of the food web (Table 3).

\scriptsize

Table 3. Summary of food web (FW) effects according to the main
stressors reported for each study area. *Industrial trawl fishery ceased
in 2015 remaining artisanal trawling activity.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3194}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Study area}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Stressor}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Food web effects}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{San Jorge Gulf} & & \\
& Fishery* & ↑ FW connectance

↓ FW stability

↓ functional diversity \\
& Sea warming & Shifts in FW topology

↑ FW connectance

↑ functional diversity \\
\textbf{Beagle Channel} & & \\
& Alien species & Shifts in FW topology

↑ spread of perturbations \\
\textbf{Burdwood Bank} & & \\
& Fishery & Shifts in main energy fluxes

↓ FW stability \\
\textbf{Scotia Sea} & & \\
& Sea warming & ↑ FW connectance

↓ FW modularity

↓ energy transfer to high TLs

↑ spread of perturbation \\
\textbf{Potter Cove (Antarctica)} & & \\
& Sea warming & ↓ perturbation spreading \\
& Sea ice decline + glacial retreat & Differential impacts, by
substrate: hard (HS) or soft (SS)

↑ FW chain length \& ↑ trophic niches (HS)

↑ FW fragility \& trophic cascades (SS) \\
\textbf{Weddell Sea (Antarctica)} & & \\
& Sea ice decline + iceberg scouring & Shifts in main energy fluxes

↑ likelihood of regime shifts

↓ resilience \\
\end{longtable}

\normalsize

\subsection{5. Gaps and future
perspectives}\label{gaps-and-future-perspectives}

In the selected study areas along the southwest Atlantic to Antarctic
latitudinal gradient, several stressors may directly affect consumers'
diets triggered by modified environmental conditions (sea warming,
reduced sea ice extent) and new species (due to species' distributional
shifts and introductions). Moreover, the population trends (biomasses
and abundances) of important species are also changing (Funes 2020;
Hindell et al. 2020; Woods et al. 2023) driving shifts in their roles as
either predators or prey (e.g. Belleggia et al. (2017), Pasti et al.
(2021)). These diet and biomass shifts should be investigated in order
to generate reliable predictions of food web responses to multiple
stressors in the southwest Atlantic - Antarctic region. One could argue
that both shifts might increase the complexity of food webs in the short
term by adding generalist predators or new prey (e.g., Cordone et al.
(2023)) or by enabling discard consumption (e.g. Funes et al. (2022)).
However, in the long term, both stressors may lead to the biological
extinction of certain prey and competitors (e.g. Anton et al. (2019)) or
a significant reduction in target and incidental catch species (e.g.
Dulvy et al. (2014)), thereby promoting food web simplification.

Since this review deals with qualitative data of prey-predator
interactions and stressor effects influencing them, adding quantitative
data to the food webs (i.e.~interaction strength) and to the stressors
(i.e.~magnitude) would lead to a better comprehension of how a given
stressor acts on specific species which might translate into food web
effects. In this context, it would be useful to develop quantitative
food web models where the strength of interactions reflect energy fluxes
among species (Nilsson and McCann 2016; Kortsch et al. 2021). Emerging
methods such as bioenergetic food web modeling have been proposed in
this regard and present promising ways to estimate shifts in species
interactions within food webs as a response to stressors (Gellner et al.
2023; Gauzens et al. n.d.). Shifts that can lead to changes in overall
ecosystem functioning and stability.

Regarding knowledge and data gaps on species and their stressors,
especially the Beagle Channel and Burdwood Bank are poorly sampled study
regions. Almost no information exists on the impact of global warming
effects (sea warming, glacial retreat, ocean acidification) on
communities in these ecosystems, though warming of mid-water and bottom
layers have been shown at a regional scale (Franco et al. 2020). Yet in
Beagle Channel recent experimental studies have tested the tolerance of
fishes to scenarios of sea warming and/or acidification suggesting
potential metabolic impacts on species (Lattuca et al. 2018; Lattuca et
al. 2023).

Analyzing the impact of multiple stressors through observational studies
is challenging (Gutt et al. 2021). This complexity arises partly because
of potential for antagonistic effects, where impacts cancel each other
out, or synergistic effects, where the combined impact is greater than
the sum of individual effects (Boyd et al. 2015; Côté et al. 2016).
Moreover, these interactive effects are complicated to handle in the
framework of complex food webs. The number of pathways through which a
species may affect or be affected by other species, and through which
stressors may permeate communities, increases exponentially with the
number of species and interactions in a network (Menge 1995). To tackle
this complexity, Beauchesne et al. (2021) developed a theory-grounded
approach using motifs (i.e.~groups of species that, when put together,
construct whole food webs) to simplify food webs; a methodology that
could be applied to our food web study cases.

\subsection{Conclusions}\label{conclusions}

We have reviewed the main environmental and anthropogenic stressors
acting in six different areas along a large-scale latitudinal gradient,
from temperate Atlantic to cold Antarctic ecosystems. Using a
theoretical framework that combines species and food web level data, we
suggest how warming effects may impact food web structure and
functioning. These qualitative predictions are intended to serve as the
basis for future studies in marine ecosystems of the Southern Hemisphere
that aim at quantifying the magnitude of these stressors and how they
are affecting quantitative food web properties, such as energy fluxes
and stability. There is an urgent need to assess these changes using a
holistic and quantitative framework where the magnitude of stressors and
species interactions are taken into account.

\subsection{Acknowledgements}\label{acknowledgements}

This review is a product of the project ``¿Cuáles son los efectos de los
cambios ambientales antropogénicos en las interacciones tróficas de las
comunidades de los ecosistemas marinos en el gradiente latitudinal
Atlántico Sudoccidental - Antártida?'', funded by Consejo Nacional de
Investigaciones Científicas (CONICET, code PIP 2022-2024 \#0907).

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Acha2004}
Acha, E.M., Mianzan, H.W., Guerrero, R.A., Favero, M., and Bava, J.
2004. Marine fronts at the continental shelves of austral {South
America}: {Physical} and ecological processes. Journal of Marine Systems
\textbf{44}(1): 83--105.
doi:\href{https://doi.org/10.1016/j.jmarsys.2003.09.005}{10.1016/j.jmarsys.2003.09.005}.

\bibitem[\citeproctext]{ref-APN2022}
Administración de Parques Nacionales. 2022. {Plan de gestión AMP
Namuncurá Banco Burdwood}. {Dirección Nacional de Áreas Marinas
Protegidas (DNAMP), Argentina}.

\bibitem[\citeproctext]{ref-Albert2002}
Albert, R., and Barabási, A.-L. 2002. Statistical mechanics of complex
networks. Reviews of Modern Physics \textbf{74}(1): 47--97. {American
Physical Society}.
doi:\href{https://doi.org/10.1103/RevModPhys.74.47}{10.1103/RevModPhys.74.47}.

\bibitem[\citeproctext]{ref-Allega2020}
Allega, L., Braverman, M., Campodónico, S., Carozza, C.R., Cepeda, G.D.,
Colonello, J.H., Derisio, C., Di Mauro, R., Firpo, C.A., Gaitán, E.,
Hozbor, C., Irusta, G., Ivanovic, M., Lagos, N., Lutz, V.A., Marí, N.R.,
Militelli, M.I., Moriondo Danovaro, P., Navarro, G., Orlando, P.,
Pájaro, M., Prandoni, N., Prosdocimi, L., Reta, R., Rico, M.R., Riestra,
C., Ruarte, C., Schejter, L., Schiariti, A., Segura, V., Souto, V.S.,
Temperoni, B., and Verón, E. 2020. {Estado del conocimiento biológico
pesquero de los principales recursos vivos y su ambiente, con relación a
la exploración hidrocarburífera en la Zona Económica Exclusiva argentina
y sus adyacencias}. {Instituto Nacional de Investigación y Desarrollo
Pesquero (INIDEP)}, {Mar del Plata}. {[}accessed 26 September 2023{]}.

\bibitem[\citeproctext]{ref-Allega2019}
Allega, L., Braverman, M.S., Cabreira, A.G., Campodónico, S., Colonello,
J.H., Derisio, C.M., Di Mauro, R.P., Gaitán, E.N., Hozbor, M., Irusta,
G., Lutz, V.A., Marí, N., Militelli, M.I., Moriondo, P., Navarro, G.,
Orlando, P., Pajaro, M., Prandoni García, N.I., Prosdocimi, L., Reta,
R., Rico, R., Riestra, C.M., Segura, V., Schejter, L., Schiariti, A.,
Souto, V., and Verón, E.M. 2019. {Estado del conocimiento biológico
pesquero de los principales recursos vivos y su ambiente, con relación a
la exploración hidrocarburífera en la Zona Económica Exclusiva Argentina
y adyacencias}. {INIDEP}. {[}accessed 10 August 2023{]}.

\bibitem[\citeproctext]{ref-Anton2019}
Anton, A., Geraldi, N.R., Lovelock, C.E., Apostolaki, E.T., Bennett, S.,
Cebrian, J., Krause-Jensen, D., Marbà, N., Martinetto, P., Pandolfi,
J.M., Santana-Garcon, J., and Duarte, C.M. 2019. Global ecological
impacts of marine exotic species. Nature Ecology \& Evolution
\textbf{3}(5): 787--800. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41559-019-0851-0}{10.1038/s41559-019-0851-0}.

\bibitem[\citeproctext]{ref-Antoni2020}
Antoni, J.S., Almandoz, G.O., Ferrario, M.E., Hernando, M.P., Varela,
D.E., Rozema, P.D., Buma, A.G.J., Paparazzo, F.E., and Schloss, I.R.
2020. Response of a natural {Antarctic} phytoplankton assemblage to
changes in temperature and salinity. Journal of Experimental Marine
Biology and Ecology \textbf{532}: 151444.
doi:\href{https://doi.org/10.1016/j.jembe.2020.151444}{10.1016/j.jembe.2020.151444}.

\bibitem[\citeproctext]{ref-Atkinson2019}
Atkinson, A., Hill, S.L., Pakhomov, E.A., Siegel, V., Reiss, C.S., Loeb,
V.J., Steinberg, D.K., Schmidt, K., Tarling, G.A., Gerrish, L., and
Sailley, S.F. 2019. Krill ({Euphausia} superba) distribution contracts
southward during rapid regional warming. Nature Climate Change
\textbf{9}(2): 142--147. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41558-018-0370-z}{10.1038/s41558-018-0370-z}.

\bibitem[\citeproctext]{ref-Atkinson2004}
Atkinson, A., Siegel, V., Pakhomov, E., and Rothery, P. 2004. Long-term
decline in krill stock and increase in salps within the {Southern
Ocean}. Nature \textbf{432}(7013): 100--103. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/nature02996}{10.1038/nature02996}.

\bibitem[\citeproctext]{ref-Balech2008}
Balech, E., and Ehrlich, M.D. 2008. {Esquema biogeográfico del Mar
Argentino}. {Mar del Plata: Instituto Nacional de Investigación y
Desarrollo Pesquero (INIDEP)}. {[}accessed 14 August 2023{]}.

\bibitem[\citeproctext]{ref-Bartley2019}
Bartley, T.J., McCann, K.S., Bieg, C., Cazelles, K., Granados, M.,
Guzzo, M.M., MacDougall, A.S., Tunney, T.D., and McMeans, B.C. 2019.
Food web rewiring in a changing world. Nature Ecology \& Evolution
\textbf{3}(3): 345--354. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41559-018-0772-3}{10.1038/s41559-018-0772-3}.

\bibitem[\citeproctext]{ref-Bauer2022}
Bauer, B., Berti, E., Ryser, R., Gauzens, B., Hirt, M.R., Rosenbaum, B.,
Digel, C., Ott, D., Scheu, S., and Brose, U. 2022. Biotic filtering by
species' interactions constrains food-web variability across spatial and
abiotic gradients. Ecology Letters \textbf{25}(5): 1225--1236.
doi:\href{https://doi.org/10.1111/ele.13995}{10.1111/ele.13995}.

\bibitem[\citeproctext]{ref-Beauchesne2021}
Beauchesne, D., Cazelles, K., Archambault, P., Dee, L.E., and Gravel, D.
2021. On the sensitivity of food webs to multiple stressors. Ecology
Letters \textbf{24}(10): 2219--2237.
doi:\href{https://doi.org/10.1111/ele.13841}{10.1111/ele.13841}.

\bibitem[\citeproctext]{ref-Belgrano2005}
Belgrano, A., Scharler, S.E.R.C.U.M., Scharler, U.M., Dunne, J., and
Ulanowicz, R.E. 2005. Aquatic {Food Webs}: {An Ecosystem Approach}. {OUP
Oxford}.

\bibitem[\citeproctext]{ref-Belleggia2017}
Belleggia, M., Giberto, D., and Bremec, C. 2017. Adaptation of diet in a
changed environment: {Increased} consumption of lobster krill {Munida}
gregaria ({Fabricius}, 1793) by {Argentine} hake. Marine Ecology
\textbf{38}(4): e12445.
doi:\href{https://doi.org/10.1111/maec.12445}{10.1111/maec.12445}.

\bibitem[\citeproctext]{ref-Bovcon2013}
Bovcon, N.D., Góngora, M.E., Marinao, C., and González-Zevallos, D.
2013. Composición de las capturas y descartes generados en la pesca de
merluza común {Merluccius} hubbsi y langostino patagónico {Pleoticus}
muelleri: Un caso de estudio en la flota fresquera de altura del {Golfo
San Jorge}, {Chubut}, {Argentina}. Revista de biología marina y
oceanografía \textbf{48}(2): 303--319. {Universidad de Valparaíso.
Facultad de Ciencias del Mar}.
doi:\href{https://doi.org/10.4067/S0718-19572013000200010}{10.4067/S0718-19572013000200010}.

\bibitem[\citeproctext]{ref-Boyd2015}
Boyd, P.W., Lennartz, S.T., Glover, D.M., and Doney, S.C. 2015.
Biological ramifications of climate-change-mediated oceanic
multi-stressors. Nature Climate Change \textbf{5}(1): 71--79. {Nature
Publishing Group}.
doi:\href{https://doi.org/10.1038/nclimate2441}{10.1038/nclimate2441}.

\bibitem[\citeproctext]{ref-Braeckman2021}
Braeckman, U., Pasotti, F., Hoffmann, R., Vázquez, S., Wulff, A.,
Schloss, I.R., Falk, U., Deregibus, D., Lefaible, N., Torstensson, A.,
Al-Handal, A., Wenzhöfer, F., and Vanreusel, A. 2021. Glacial melt
disturbance shifts community metabolism of an {Antarctic} seafloor
ecosystem from net autotrophy to heterotrophy. Communications Biology
\textbf{4}(1): 148.
doi:\href{https://doi.org/10.1038/s42003-021-01673-6}{10.1038/s42003-021-01673-6}.

\bibitem[\citeproctext]{ref-Braithwaite2015}
Braithwaite, J.E., Meeuwig, J.J., Letessier, T.B., Jenner, K.C.S., and
Brierley, A.S. 2015. From sea ice to blubber: Linking whale condition to
krill abundance using historical whaling records. Polar Biology
\textbf{38}(8): 1195--1202.
doi:\href{https://doi.org/10.1007/s00300-015-1685-0}{10.1007/s00300-015-1685-0}.

\bibitem[\citeproctext]{ref-Brown2004}
Brown, J.H., Gillooly, J.F., Allen, A.P., Savage, V.M., and West, G.B.
2004. Toward a {Metabolic Theory} of {Ecology}. Ecology \textbf{85}(7):
1771--1789. doi:\href{https://doi.org/10.1890/03-9000}{10.1890/03-9000}.

\bibitem[\citeproctext]{ref-Bruder2019}
Bruder, A., Frainer, A., Rota, T., and Primicerio, R. 2019. The
{Importance} of {Ecological Networks} in {Multiple-Stressor Research}
and {Management}. Frontiers in Environmental Science \textbf{7}.
{[}accessed 29 August 2023{]}.

\bibitem[\citeproctext]{ref-Bruno2023}
Bruno, D.O., Riccialdelli, L., Acha, E.M., and Fernández, D.A. 2023.
Seasonal variation of autochthonous and allochthonous carbon sources for
the first levels of the {Beagle Channel} food web. Journal of Marine
Systems \textbf{239}: 103859.
doi:\href{https://doi.org/10.1016/j.jmarsys.2023.103859}{10.1016/j.jmarsys.2023.103859}.

\bibitem[\citeproctext]{ref-Chown2022}
Chown, S.L., Leihy, R.I., Naish, T.R., Brooks, C.M., Convey, P., Henley,
B.J., Mackintosh, A.N., Phillips, L.M., Kennicutt II, M.C., and Grant,
S.M. 2022. Antarctic climate change and the environment: A decadal
synopsis and recommendations for action.

\bibitem[\citeproctext]{ref-Ciancio2008}
Ciancio, J.E., Pascual, M.A., Botto, F., Frere, E., and Iribarne, O.
2008. Trophic relationships of exotic anadromous salmonids in the
southern {Patagonian Shelf} as inferred from stable isotopes. Limnology
and Oceanography \textbf{53}(2): 788--798.
doi:\href{https://doi.org/10.4319/lo.2008.53.2.0788}{10.4319/lo.2008.53.2.0788}.

\bibitem[\citeproctext]{ref-Cirtwill2015}
Cirtwill, A.R., Stouffer, D.B., and Romanuk, T.N. 2015. Latitudinal
gradients in biotic niche breadth vary across ecosystem types.
Proceedings of the Royal Society B: Biological Sciences
\textbf{282}(1819): 20151589. {Royal Society}.
doi:\href{https://doi.org/10.1098/rspb.2015.1589}{10.1098/rspb.2015.1589}.

\bibitem[\citeproctext]{ref-Collins2010}
Collins, M.A., Brickle, P., Brown, J., and Belchier, M. 2010. Chapter
{Four} - {The Patagonian Toothfish}: {Biology}, {Ecology} and {Fishery}.
\emph{In} Advances in {Marine Biology}. \emph{Edited by} M. Lesser.
{Academic Press}. pp. 227--300.
doi:\href{https://doi.org/10.1016/B978-0-12-381015-1.00004-6}{10.1016/B978-0-12-381015-1.00004-6}.

\bibitem[\citeproctext]{ref-Commendatore2012}
Commendatore, M.G., Nievas, M.L., Amin, O., and Esteves, J.L. 2012.
Sources and distribution of aliphatic and polyaromatic hydrocarbons in
coastal sediments from the {Ushuaia Bay} ({Tierra} del {Fuego},
{Patagonia}, {Argentina}). Marine Environmental Research \textbf{74}:
20--31.
doi:\href{https://doi.org/10.1016/j.marenvres.2011.11.010}{10.1016/j.marenvres.2011.11.010}.

\bibitem[\citeproctext]{ref-Constable2014}
Constable, A.J., Melbourne-Thomas, J., Corney, S.P., Arrigo, K.R.,
Barbraud, C., Barnes, D.K.A., Bindoff, N.L., Boyd, P.W., Brandt, A.,
Costa, D.P., Davidson, A.T., Ducklow, H.W., Emmerson, L., Fukuchi, M.,
Gutt, J., Hindell, M.A., Hofmann, E.E., Hosie, G.W., Iida, T., Jacob,
S., Johnston, N.M., Kawaguchi, S., Kokubun, N., Koubbi, P., Lea, M.-A.,
Makhado, A., Massom, R.A., Meiners, K., Meredith, M.P., Murphy, E.J.,
Nicol, S., Reid, K., Richerson, K., Riddle, M.J., Rintoul, S.R., Smith,
W.O., Southwell, C., Stark, J.S., Sumner, M., Swadling, K.M., Takahashi,
K.T., Trathan, P.N., Welsford, D.C., Weimerskirch, H., Westwood, K.J.,
Wienecke, B.C., Wolf-Gladrow, D., Wright, S.W., Xavier, J.C., and
Ziegler, P. 2014. Climate change and {Southern Ocean} ecosystems {I}:
How changes in physical habitats directly affect marine biota. Global
Change Biology \textbf{20}(10): 3004--3025.
doi:\href{https://doi.org/10.1111/gcb.12623}{10.1111/gcb.12623}.

\bibitem[\citeproctext]{ref-Cordone2023}
Cordone, G., Galván, D.E., and Momo, F.R. 2023. Impacts of an invasion
by green crab {Carcinus} maenas on the intertidal food web of a
{Patagonian} rocky shore, {Argentina}. Marine Ecology Progress Series
\textbf{713}: 97--115.
doi:\href{https://doi.org/10.3354/meps14336}{10.3354/meps14336}.

\bibitem[\citeproctext]{ref-Cordone2020}
Cordone, G., Salinas, V., Marina, T.I., Doyle, S.R., Pasotti, F.,
Saravia, L.A., and Momo, F.R. 2020. Green vs brown food web: {Effects}
of habitat type on multidimensional stability proxies for a
highly-resolved {Antarctic} food web. Food Webs \textbf{25}: e00166.
doi:\href{https://doi.org/10.1016/j.fooweb.2020.e00166}{10.1016/j.fooweb.2020.e00166}.

\bibitem[\citeproctext]{ref-Correa2008}
Correa, C., and Gross, M.R. 2008. Chinook salmon invade southern {South
America}. Biological Invasions \textbf{10}(5): 615--639.
doi:\href{https://doi.org/10.1007/s10530-007-9157-2}{10.1007/s10530-007-9157-2}.

\bibitem[\citeproctext]{ref-Cossi2021}
Cossi, P.F., Ojeda, M., Chiesa, I.L., Rimondino, G.N., Fraysse, C.,
Calcagno, J., and Pérez, A.F. 2021. First evidence of microplastics in
the {Marine Protected Area Namuncurá} at {Burdwood Bank}, {Argentina}: A
study on {Henricia} obesa and {Odontaster} penicillatus
({Echinodermata}: {Asteroidea}). Polar Biology \textbf{44}(12):
2277--2287.
doi:\href{https://doi.org/10.1007/s00300-021-02959-5}{10.1007/s00300-021-02959-5}.

\bibitem[\citeproctext]{ref-Cote2016}
Côté, I.M., Darling, E.S., and Brown, C.J. 2016. Interactions among
ecosystem stressors and their importance in conservation. Proceedings of
the Royal Society B: Biological Sciences \textbf{283}(1824): 20152592.
{Royal Society}.
doi:\href{https://doi.org/10.1098/rspb.2015.2592}{10.1098/rspb.2015.2592}.

\bibitem[\citeproctext]{ref-deHaro2022}
de Haro, J.C., Perez Orsi, H., Cané, S., Di Pangracio, A., Falabella,
V., and Sapoznikow, A. 2022. {Informe colaborativo sobre el Estado de
situación. Riesgos e impactos de la prospección sísmica en el Mar
Argentino}. {Foro para la Conservación del Mar Patagónico y Áreas de
Influencia}.

\bibitem[\citeproctext]{ref-Deregibus2023}
Deregibus, D., Campana, G.L., Neder, C., Barnes, D.K.A., Zacher, K.,
Piscicelli, J.M., Jerosch, K., and Quartino, M.L. 2023. Potential
macroalgal expansion and blue carbon gains with northern {Antarctic
Peninsula} glacial retreat. Marine Environmental Research \textbf{189}:
106056.
doi:\href{https://doi.org/10.1016/j.marenvres.2023.106056}{10.1016/j.marenvres.2023.106056}.

\bibitem[\citeproctext]{ref-Deregibus2017}
Deregibus, D., Quartino, M.L., Zacher, K., Campana, G.L., and Barnes,
D.K.A. 2017. Understanding the link between sea ice, ice scour and
{Antarctic} benthic biodiversity\textendash the need for cross-station
and international collaboration. Polar Record \textbf{53}(2): 143--152.
doi:\href{https://doi.org/10.1017/S0032247416000875}{10.1017/S0032247416000875}.

\bibitem[\citeproctext]{ref-Deregibus2020}
Deregibus, D., Zacher, K., Bartsch, I., Campana, G.L., Momo, F.R.,
Wiencke, C., Gómez, I., and Quartino, M.L. 2020. Carbon {Balance Under}
a {Changing Light Environment}. \emph{In} Antarctic {Seaweeds}:
{Diversity}, {Adaptation} and {Ecosystem Services}. \emph{Edited by} I.
Gómez and P. Huovinen. {Springer International Publishing}, {Cham}. pp.
173--191.
doi:\href{https://doi.org/10.1007/978-3-030-39448-6_9}{10.1007/978-3-030-39448-6\_9}.

\bibitem[\citeproctext]{ref-DiMauro2022}
Di Mauro, R., Castillo, S., Pérez, A., Iachetti, C.M., Silva, L., Tomba,
J.P., and Chiesa, I.L. 2022. Anthropogenic microfibers are highly
abundant at the {Burdwood Bank} seamount, a protected sub-{Antarctic}
environment in the {Southwestern Atlantic Ocean}. Environmental
Pollution \textbf{306}: 119364.
doi:\href{https://doi.org/10.1016/j.envpol.2022.119364}{10.1016/j.envpol.2022.119364}.

\bibitem[\citeproctext]{ref-Dodino2022}
Dodino, S., Riccialdelli, L., Polito, M.J., Pütz, K., Brasso, R.L., and
Raya Rey, A. 2022. Mercury exposure driven by geographic and trophic
factors in {Magellanic} penguins from {Tierra} del {Fuego}. Marine
Pollution Bulletin \textbf{174}: 113184.
doi:\href{https://doi.org/10.1016/j.marpolbul.2021.113184}{10.1016/j.marpolbul.2021.113184}.

\bibitem[\citeproctext]{ref-Duarte2011}
Duarte, C.A., Giarratano, E., Amin, O.A., and Comoglio, L.I. 2011. Heavy
metal concentrations and biomarkers of oxidative stress in native
mussels ({Mytilus} edulis chilensis) from {Beagle Channel} coast
({Tierra} del {Fuego}, {Argentina}). Marine Pollution Bulletin
\textbf{62}(8): 1895--1904.
doi:\href{https://doi.org/10.1016/j.marpolbul.2011.05.031}{10.1016/j.marpolbul.2011.05.031}.

\bibitem[\citeproctext]{ref-Duffy2007}
Duffy, J.E., Cardinale, B.J., France, K.E., McIntyre, P.B., Thébault,
E., and Loreau, M. 2007. The functional role of biodiversity in
ecosystems: Incorporating trophic complexity. Ecology Letters
\textbf{10}(6): 522--538.
doi:\href{https://doi.org/10.1111/j.1461-0248.2007.01037.x}{10.1111/j.1461-0248.2007.01037.x}.

\bibitem[\citeproctext]{ref-Dulvy2014}
Dulvy, N.K., Fowler, S.L., Musick, J.A., Cavanagh, R.D., Kyne, P.M.,
Harrison, L.R., Carlson, J.K., Davidson, L.N., Fordham, S.V., Francis,
M.P., Pollock, C.M., Simpfendorfer, C.A., Burgess, G.H., Carpenter,
K.E., Compagno, L.J., Ebert, D.A., Gibson, C., Heupel, M.R.,
Livingstone, S.R., Sanciangco, J.C., Stevens, J.D., Valenti, S., and
White, W.T. 2014. Extinction risk and conservation of the world's sharks
and rays. eLife \textbf{3}: e00590. {eLife Sciences Publications, Ltd}.
doi:\href{https://doi.org/10.7554/eLife.00590}{10.7554/eLife.00590}.

\bibitem[\citeproctext]{ref-Dunne2002}
Dunne, J.A., Williams, R.J., and Martinez, N.D. 2002a. Network structure
and biodiversity loss in food webs: Robustness increases with
connectance. Ecology Letters \textbf{5}(4): 558--567.
doi:\href{https://doi.org/10.1046/j.1461-0248.2002.00354.x}{10.1046/j.1461-0248.2002.00354.x}.

\bibitem[\citeproctext]{ref-Dunne2002a}
Dunne, J.A., Williams, R.J., and Martinez, N.D. 2002b. Food-web
structure and network theory: {The} role of connectance and size.
Proceedings of the National Academy of Sciences \textbf{99}(20):
12917--12922. {Proceedings of the National Academy of Sciences}.
doi:\href{https://doi.org/10.1073/pnas.192407699}{10.1073/pnas.192407699}.

\bibitem[\citeproctext]{ref-Fernandez2010}
Fernández, D.A., Ciancio, J., Ceballos, S.G., Riva-Rossi, C., and
Pascual, M.A. 2010. Chinook salmon ({Oncorhynchus} tshawytscha,
{Walbaum} 1792) in the {Beagle Channel}, {Tierra} del {Fuego}: The onset
of an invasion. Biological Invasions \textbf{12}(9): 2991--2997.
doi:\href{https://doi.org/10.1007/s10530-010-9731-x}{10.1007/s10530-010-9731-x}.

\bibitem[\citeproctext]{ref-Ferreira2021}
Ferreira, M.F., Lo Nostro, F.L., Fernández, D.A., and Genovese, G. 2021.
Endocrine disruption in the sub {Antarctic} fish {Patagonotothen}
tessellata ({Perciformes}, {Notothenidae}) from {Beagle Channel}
associated to anthropogenic impact. Marine Environmental Research
\textbf{171}: 105478.
doi:\href{https://doi.org/10.1016/j.marenvres.2021.105478}{10.1016/j.marenvres.2021.105478}.

\bibitem[\citeproctext]{ref-Fioramonti2022}
Fioramonti, N.E., Ribeiro Guevara, S., Becker, Y.A., and Riccialdelli,
L. 2022. Mercury transfer in coastal and oceanic food webs from the
{Southwest Atlantic Ocean}. Marine Pollution Bulletin \textbf{175}:
113365.
doi:\href{https://doi.org/10.1016/j.marpolbul.2022.113365}{10.1016/j.marpolbul.2022.113365}.

\bibitem[\citeproctext]{ref-Flores2012}
Flores, H., Atkinson, A., Kawaguchi, S., Krafft, B., Milinevsky, G.,
Nicol, S., Reiss, C., Tarling, G., Werner, R., Bravo Rebolledo, E.,
Cirelli, V., Cuzin-Roudy, J., Fielding, S., Van Franeker, J.,
Groeneveld, J., Haraldsson, M., Lombana, A., Marschoff, E., Meyer, B.,
Pakhomov, E., Van De Putte, A., Rombolá, E., Schmidt, K., Siegel, V.,
Teschke, M., Tonkes, H., Toullec, J., Trathan, P., Tremblay, N., and
Werner, T. 2012. Impact of climate change on {Antarctic} krill. Marine
Ecology Progress Series \textbf{458}: 1--19.
doi:\href{https://doi.org/10.3354/meps09831}{10.3354/meps09831}.

\bibitem[\citeproctext]{ref-Franco2020}
Franco, B.C., Combes, V., and González Carman, V. 2020. Subsurface
{Ocean Warming Hotspots} and {Potential Impacts} on {Marine Species}:
{The Southwest South Atlantic Ocean Case Study}. Frontiers in Marine
Science \textbf{7}. {[}accessed 10 August 2023{]}.

\bibitem[\citeproctext]{ref-Fretwell2023}
Fretwell, P.T., Boutet, A., and Ratcliffe, N. 2023. Record low 2022
{Antarctic} sea ice led to catastrophic breeding failure of emperor
penguins. Communications Earth \& Environment \textbf{4}(1): 1--6.
{Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s43247-023-00927-x}{10.1038/s43247-023-00927-x}.

\bibitem[\citeproctext]{ref-Fuentes2016}
Fuentes, V., Alurralde, G., Meyer, B., Aguirre, G.E., Canepa, A., Wölfl,
A.-C., Hass, H.C., Williams, G.N., and Schloss, I.R. 2016. Glacial
melting: An overlooked threat to {Antarctic} krill. Scientific Reports
\textbf{6}(1): 27234. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/srep27234}{10.1038/srep27234}.

\bibitem[\citeproctext]{ref-Funes2020}
Funes, M. 2020, March. Efectos de la pesca de arrastre sobre la
estructura trófica del norte del {Golfo San Jorge}. PhD thesis.

\bibitem[\citeproctext]{ref-Funes2019}
Funes, M., Marinao, C., and Galván, D.E. 2019. Does trawl fisheries
affect the diet of fishes? {A} stable isotope analysis approach*.
Isotopes in Environmental and Health Studies \textbf{55}(4): 327--343.
{Taylor \& Francis}.
doi:\href{https://doi.org/10.1080/10256016.2019.1626381}{10.1080/10256016.2019.1626381}.

\bibitem[\citeproctext]{ref-Funes2022}
Funes, M., Saravia, L.A., Cordone, G., Iribarne, O.O., and Galván, D.E.
2022. Network analysis suggests changes in food web stability produced
by bottom trawl fishery in {Patagonia}. Scientific Reports
\textbf{12}(1): 10876. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41598-022-14363-y}{10.1038/s41598-022-14363-y}.

\bibitem[\citeproctext]{ref-Gaitan2016}
Gaitán, E., and Marí, N. 2016. {Análisis de las comunidades bentónicas
asociadas a capturas de la flota comercial dirigida a Macruronus
magellanicus}. {Instituto Nacional de Investigación y Desarrollo
Pesquero (INIDEP)}.

\bibitem[\citeproctext]{ref-Galvan2022}
Galván, D.E., Bovcon, N.D., Cochia, P.D., González, R.A., Lattuca, M.E.,
Reinaldo, M.O., Rincón-Díaz, M.P., Romero, M.A., Vanella, F.A., Venerus,
L.A., and Svendsen, G.M. 2022. Changes in the {Specific} and
{Biogeographic Composition} of {Coastal Fish Assemblages} in
{Patagonia}, {Driven} by {Climate Change}, {Fishing}, and {Invasion} by
{Alien Species}. \emph{In} Global {Change} in {Atlantic Coastal
Patagonian Ecosystems}: {A Journey Through Time}. \emph{Edited by} E.W.
Helbling, M.A. Narvarte, R.A. González, and V.E. Villafañe. {Springer
International Publishing}, {Cham}. pp. 205--231.
doi:\href{https://doi.org/10.1007/978-3-030-86676-1_9}{10.1007/978-3-030-86676-1\_9}.

\bibitem[\citeproctext]{ref-Garcia2019}
Garcia, M.D., Fernández Severini, M.D., Spetter, C., López Abbate, M.C.,
Tartara, M.N., Nahuelhual, E.G., Marcovecchio, J.E., Schloss, I.R., and
Hoffmeyer, M.S. 2019. Effects of glacier melting on the planktonic
communities of two {Antarctic} coastal areas ({Potter Cove} and {Hope
Bay}) in summer. Regional Studies in Marine Science \textbf{30}: 100731.
doi:\href{https://doi.org/10.1016/j.rsma.2019.100731}{10.1016/j.rsma.2019.100731}.

\bibitem[\citeproctext]{ref-Garcia2016}
Garcia, M.D., Hoffmeyer, M.S., Abbate, M.C.L., Barría De Cao, M.S.,
Pettigrosso, R.E., Almandoz, G.O., Hernando, M.P., and Schloss, I.R.
2016. Micro- and mesozooplankton responses during two contrasting
summers in a coastal {Antarctic} environment. Polar Biology
\textbf{39}(1): 123--137.
doi:\href{https://doi.org/10.1007/s00300-015-1678-z}{10.1007/s00300-015-1678-z}.

\bibitem[\citeproctext]{ref-Garcia-Borboroglu2008}
García-Borboroglu, P., Boersma, P.D., Reyes, L., and Skewgar, E. 2008.
Petroleum {Pollution} and {Penguins}: {Marine Conservation Tools} to
{Reduce} the {Problem}. \emph{In} Marine {Pollution}: {New} research,
Hofer, T.N. {Nova Science Publishers Inc.}, {New York, USA}. pp.
339--356.

\bibitem[\citeproctext]{ref-Gauzens2023}
Gauzens, B., Brose, U., Delmas, E., and Berti, E. (n.d.). {ATNr}:
{Allometric Trophic Network} models in {R}. Methods in Ecology and
Evolution \textbf{n/a}(n/a).
doi:\href{https://doi.org/10.1111/2041-210X.14212}{10.1111/2041-210X.14212}.

\bibitem[\citeproctext]{ref-Gellner2023}
Gellner, G., McCann, K., and Hastings, A. 2023. Stable diverse food webs
become more common when interactions are more biologically constrained.
Proceedings of the National Academy of Sciences \textbf{120}(31):
e2212061120. {Proceedings of the National Academy of Sciences}.
doi:\href{https://doi.org/10.1073/pnas.2212061120}{10.1073/pnas.2212061120}.

\bibitem[\citeproctext]{ref-Gil2011}
Gil, M.N., Torres, A.I., Amin, O., and Esteves, J.L. 2011. Assessment of
recent sediment influence in an urban polluted subantarctic coastal
ecosystem. {Beagle Channel} ({Southern Argentina}). Marine Pollution
Bulletin \textbf{62}(1): 201--207.
doi:\href{https://doi.org/10.1016/j.marpolbul.2010.10.004}{10.1016/j.marpolbul.2010.10.004}.

\bibitem[\citeproctext]{ref-Glembocki2015}
Glembocki, N.G., Williams, G.N., Góngora, M.E., Gagliardini, D.A., and
Orensanz, J.M.(Lobo). 2015. Synoptic oceanography of {San Jorge Gulf}
({Argentina}): {A} template for {Patagonian} red shrimp ({Pleoticus}
muelleri) spatial dynamics. Journal of Sea Research \textbf{95}: 22--35.
doi:\href{https://doi.org/10.1016/j.seares.2014.10.011}{10.1016/j.seares.2014.10.011}.

\bibitem[\citeproctext]{ref-Gongora2012}
Góngora, M.E., González-Zevallos, D., Pettovello, A., and Mendía, L.
2012. Caracterización de las principales pesquerías del golfo {San Jorge
Patagonia}, {Argentina}. Latin american journal of aquatic research
\textbf{40}(1): 1--11. {Pontificia Universidad Católica de Valparaíso.
Facultad de Recursos Naturales. Escuela de Ciencias del Mar}.
{[}accessed 14 August 2023{]}.

\bibitem[\citeproctext]{ref-Gonzalez-Zevallos2006}
González-Zevallos, D., and Yorio, P. 2006. Seabird use of discards and
incidental captures at the {Argentine} hake trawl fishery in the {Golfo
San Jorge}, {Argentina}. Marine Ecology Progress Series \textbf{316}:
175--183.
doi:\href{https://doi.org/10.3354/meps316175}{10.3354/meps316175}.

\bibitem[\citeproctext]{ref-Gorini2021}
Gorini, F.L., Lukaszewicz, G., and Giussi, A.R. 2021. {Actualización de
la estadística pesquera de peces demersales australes en el Atlántico
Sudoccidental (período 2008-2020)}. \{Informe
T\textbackslash\textquotesingle ecnico Oficial\}, {Instituto Nacional de
Investigación y Desarrollo Pesquero (INIDEP)}, {Mar del Plata}.

\bibitem[\citeproctext]{ref-Guihou2020}
Guihou, K., Piola, A.R., Palma, E.D., and Chidichimo, M.P. 2020.
Dynamical connections between large marine ecosystems of austral {South
America} based on numerical simulations. Ocean Science \textbf{16}(2):
271--290. {Copernicus GmbH}.
doi:\href{https://doi.org/10.5194/os-16-271-2020}{10.5194/os-16-271-2020}.

\bibitem[\citeproctext]{ref-Gutt2001a}
Gutt, J. 2001. On the direct impact of ice on marine benthic
communities, a review. Polar Biology \textbf{24}(8): 553--564.
doi:\href{https://doi.org/10.1007/s003000100262}{10.1007/s003000100262}.

\bibitem[\citeproctext]{ref-Gutt2015}
Gutt, J., Bertler, N., Bracegirdle, T.J., Buschmann, A., Comiso, J.,
Hosie, G., Isla, E., Schloss, I.R., Smith, C.R., Tournadre, J., and
Xavier, J.C. 2015. The {Southern Ocean} ecosystem under multiple climate
change stresses - an integrated circumpolar assessment. Global Change
Biology \textbf{21}(4): 1434--1453.
doi:\href{https://doi.org/10.1111/gcb.12794}{10.1111/gcb.12794}.

\bibitem[\citeproctext]{ref-Gutt2021}
Gutt, J., Isla, E., Xavier, J.C., Adams, B.J., Ahn, I.-Y., Cheng,
C.-H.C., Colesie, C., Cummings, V.J., di Prisco, G., Griffiths, H.,
Hawes, I., Hogg, I., McIntyre, T., Meiners, K.M., Pearce, D.A., Peck,
L., Piepenburg, D., Reisinger, R.R., Saba, G.K., Schloss, I.R., Signori,
C.N., Smith, C.R., Vacchi, M., Verde, C., and Wall, D.H. 2021. Antarctic
ecosystems in transition \textendash{} life between stresses and
opportunities. Biological Reviews \textbf{96}(3): 798--821.
doi:\href{https://doi.org/10.1111/brv.12679}{10.1111/brv.12679}.

\bibitem[\citeproctext]{ref-Gutt2003}
Gutt, J., and Piepenburg, D. 2003. Scale-dependent impact on diversity
of {Antarctic} benthos caused by grounding of icebergs. Marine Ecology
Progress Series \textbf{253}: 77--83.
doi:\href{https://doi.org/10.3354/meps253077}{10.3354/meps253077}.

\bibitem[\citeproctext]{ref-Gutt2001}
Gutt, J., and Starmans, A. 2001. Quantification of iceberg impact and
benthic recolonisation patterns in the {Weddell Sea} ({Antarctica}).
Polar Biology \textbf{24}(8): 615--619.
doi:\href{https://doi.org/10.1007/s003000100263}{10.1007/s003000100263}.

\bibitem[\citeproctext]{ref-Hilborn2017}
Hilborn, R., Amoroso, R.O., Bogazzi, E., Jensen, O.P., Parma, A.M.,
Szuwalski, C., and Walters, C.J. 2017. When does fishing forage species
affect their predators? Fisheries Research \textbf{191}: 211--221.
doi:\href{https://doi.org/10.1016/j.fishres.2017.01.008}{10.1016/j.fishres.2017.01.008}.

\bibitem[\citeproctext]{ref-Hindell2020}
Hindell, M.A., Reisinger, R.R., Ropert-Coudert, Y., Hückstädt, L.A.,
Trathan, P.N., Bornemann, H., Charrassin, J.-B., Chown, S.L., Costa,
D.P., Danis, B., Lea, M.-A., Thompson, D., Torres, L.G., Van de Putte,
A.P., Alderman, R., Andrews-Goff, V., Arthur, B., Ballard, G., Bengtson,
J., Bester, M.N., Blix, A.S., Boehme, L., Bost, C.-A., Boveng, P.,
Cleeland, J., Constantine, R., Corney, S., Crawford, R.J.M., Dalla Rosa,
L., de Bruyn, P.J.N., Delord, K., Descamps, S., Double, M., Emmerson,
L., Fedak, M., Friedlaender, A., Gales, N., Goebel, M.E., Goetz, K.T.,
Guinet, C., Goldsworthy, S.D., Harcourt, R., Hinke, J.T., Jerosch, K.,
Kato, A., Kerry, K.R., Kirkwood, R., Kooyman, G.L., Kovacs, K.M.,
Lawton, K., Lowther, A.D., Lydersen, C., Lyver, P.O., Makhado, A.B.,
Márquez, M.E.I., McDonald, B.I., McMahon, C.R., Muelbert, M.,
Nachtsheim, D., Nicholls, K.W., Nordøy, E.S., Olmastroni, S., Phillips,
R.A., Pistorius, P., Plötz, J., Pütz, K., Ratcliffe, N., Ryan, P.G.,
Santos, M., Southwell, C., Staniland, I., Takahashi, A., Tarroux, A.,
Trivelpiece, W., Wakefield, E., Weimerskirch, H., Wienecke, B., Xavier,
J.C., Wotherspoon, S., Jonsen, I.D., and Raymond, B. 2020. Tracking of
marine predators to protect {Southern Ocean} ecosystems. Nature
\textbf{580}(7801): 87--92. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41586-020-2126-y}{10.1038/s41586-020-2126-y}.

\bibitem[\citeproctext]{ref-Hoffmann2019}
Hoffmann, R., Al-Handal, A.Y., Wulff, A., Deregibus, D., Zacher, K.,
Quartino, M.L., Wenzhöfer, F., and Braeckman, U. 2019. Implications of
{Glacial Melt-Related Processes} on the {Potential Primary Production}
of a {Microphytobenthic Community} in {Potter Cove} ({Antarctica}).
Frontiers in Marine Science \textbf{6}: 655.
doi:\href{https://doi.org/10.3389/fmars.2019.00655}{10.3389/fmars.2019.00655}.

\bibitem[\citeproctext]{ref-Iken2023}
Iken, K., Amsler, C.D., Gorman, K.B., Klein, A.G., Galloway, A.W.E.,
Amsler, M.O., Heiser, S., Whippo, R., Lowe, A.T., Schram, J.B.,
Schneider, Z.X., and McClintock, J.B. 2023. Macroalgal input into the
coastal food web along a gradient of seasonal sea ice cover along the
{Western Antarctic Peninsula}. Marine Ecology Progress Series
\textbf{718}: 1--22.
doi:\href{https://doi.org/10.3354/meps14388}{10.3354/meps14388}.

\bibitem[\citeproctext]{ref-Irons2000}
Irons, D.B., Kendall, S.J., Erickson, W.P., McDonald, L.L., and Lance,
B.K. 2000. Nine {Years After} the {Exxon Valdez Oil Spill}: {Effects} on
{Marine Bird Populations} in {Prince William Sound}, {Alaska}. The
Condor \textbf{102}(4): 723--737.
doi:\href{https://doi.org/10.1093/condor/102.4.723}{10.1093/condor/102.4.723}.

\bibitem[\citeproctext]{ref-Isla2023}
Isla, E. 2023. Animal\textendash{{Energy Relationships}} in a {Changing
Ocean}: {The Case} of {Continental Shelf Macrobenthic Communities} on
the {Weddell Sea} and the {Vicinity} of the {Antarctic Peninsula}.
Biology \textbf{12}(5): 659. {Multidisciplinary Digital Publishing
Institute}.
doi:\href{https://doi.org/10.3390/biology12050659}{10.3390/biology12050659}.

\bibitem[\citeproctext]{ref-Jacob2011}
Jacob, U., Thierry, A., Brose, U., Arntz, W.E., Berg, S., Brey, T.,
Fetzer, I., Jonsson, T., Mintenbeck, K., Möllmann, C., Petchey, O.L.,
Riede, J.O., and Dunne, J.A. 2011. The {Role} of {Body Size} in {Complex
Food Webs}: {A Cold Case}. \emph{In} Advances in {Ecological Research}.
\emph{Edited by} A. Belgrano. {Academic Press}. pp. 181--223.
doi:\href{https://doi.org/10.1016/B978-0-12-386475-8.00005-8}{10.1016/B978-0-12-386475-8.00005-8}.

\bibitem[\citeproctext]{ref-Jerosch2018}
Jerosch, K., Pehlke, H., Monien, P., Scharf, F., Weber, L., Kuhn, G.,
Braun, M.H., and Abele, D. 2018. Benthic meltwater fjord habitats formed
by rapid glacier recession on {King George Island}, {Antarctica}.
Philosophical Transactions of the Royal Society A: Mathematical,
Physical and Engineering Sciences \textbf{376}(2122): 20170178.
doi:\href{https://doi.org/10.1098/rsta.2017.0178}{10.1098/rsta.2017.0178}.

\bibitem[\citeproctext]{ref-Jordan2007}
Jordán, F., Benedek, Z., and Podani, J. 2007. Quantifying positional
importance in food webs: {A} comparison of centrality indices.
Ecological Modelling \textbf{205}(1): 270--275.
doi:\href{https://doi.org/10.1016/j.ecolmodel.2007.02.032}{10.1016/j.ecolmodel.2007.02.032}.

\bibitem[\citeproctext]{ref-Kaminskyinprep}
Kaminsky, J., Bagur, M., Boraso, A., Schloss, I.R., Buschmann, A.H., and
Quartino, M.L. in prep. Kelp recruitment declines in a sub-{Antarctic}
ecosystem with high sediment inputs and changes in the understory algae.

\bibitem[\citeproctext]{ref-Kawaguchi2009}
Kawaguchi, S., Nicol, S., and Press, A. 2009. Direct effects of climate
change on the {Antarctic} krill fishery. {University Of Tasmania}.
{[}accessed 13 September 2023{]}.

\bibitem[\citeproctext]{ref-Klein2018}
Klein, E.S., Hill, S.L., Hinke, J.T., Phillips, T., and Watters, G.M.
2018. Impacts of rising sea temperature on krill increase risks for
predators in the {Scotia Sea}. PLOS ONE \textbf{13}(1): e0191011.
{Public Library of Science}.
doi:\href{https://doi.org/10.1371/journal.pone.0191011}{10.1371/journal.pone.0191011}.

\bibitem[\citeproctext]{ref-Kortsch2021}
Kortsch, S., Frelat, R., Pecuchet, L., Olivier, P., Putnis, I.,
Bonsdorff, E., Ojaveer, H., Jurgensone, I., Strāķe, S., Rubene, G.,
Krūze, Ē., and Nordström, M.C. 2021. Disentangling temporal food web
dynamics facilitates understanding of ecosystem functioning. Journal of
Animal Ecology \textbf{90}(5): 1205--1216.
doi:\href{https://doi.org/10.1111/1365-2656.13447}{10.1111/1365-2656.13447}.

\bibitem[\citeproctext]{ref-Kortsch2019}
Kortsch, S., Primicerio, R., Aschan, M., Lind, S., Dolgov, A.V., and
Planque, B. 2019. Food-web structure varies along environmental
gradients in a high-latitude marine ecosystem. Ecography \textbf{42}(2):
295--308.
doi:\href{https://doi.org/10.1111/ecog.03443}{10.1111/ecog.03443}.

\bibitem[\citeproctext]{ref-Kortsch2015}
Kortsch, S., Primicerio, R., Fossheim, M., Dolgov, A.V., and Aschan, M.
2015. Climate change alters the structure of arctic marine food webs due
to poleward shifts of boreal generalists. Proceedings of the Royal
Society B: Biological Sciences \textbf{282}(1814): 20151546.
doi:\href{https://doi.org/10.1098/rspb.2015.1546}{10.1098/rspb.2015.1546}.

\bibitem[\citeproctext]{ref-Kratina2012}
Kratina, P., LeCraw, R.M., Ingram, T., and Anholt, B.R. 2012. Stability
and persistence of food webs with omnivory: {Is} there a general
pattern? Ecosphere \textbf{3}(6): art50.
doi:\href{https://doi.org/10.1890/ES12-00121.1}{10.1890/ES12-00121.1}.

\bibitem[\citeproctext]{ref-Latorre2023}
Latorre, M.P., Iachetti, C.M., Schloss, I.R., Antoni, J., Malits, A., De
La Rosa, F., De Troch, M., Garcia, M.D., Flores-Melo, X., Romero, S.I.,
Gil, M.N., and Hernando, M. 2023. Summer heatwaves affect coastal
{Antarctic} plankton metabolism and community structure. Journal of
Experimental Marine Biology and Ecology \textbf{567}: 151926.
doi:\href{https://doi.org/10.1016/j.jembe.2023.151926}{10.1016/j.jembe.2023.151926}.

\bibitem[\citeproctext]{ref-Lattuca2018}
Lattuca, M.E., Boy, C.C., Vanella, F.A., Barrantes, M.E., and Fernández,
D.A. 2018. Thermal responses of three native fishes from estuarine areas
of the {Beagle Channel}, and their implications for climate change.
Hydrobiologia \textbf{808}(1): 235--249.
doi:\href{https://doi.org/10.1007/s10750-017-3424-8}{10.1007/s10750-017-3424-8}.

\bibitem[\citeproctext]{ref-Lattuca2023}
Lattuca, M.E., Vanella, F.A., Malanga, G., Rubel, M.D., Manríquez, P.H.,
Torres, R., Alter, K., Marras, S., Peck, M.A., Domenici, P., and
Fernández, D.A. 2023. Ocean acidification and seasonal temperature
extremes combine to impair the thermal physiology of a sub-{Antarctic}
fish. Science of The Total Environment \textbf{856}: 159284.
doi:\href{https://doi.org/10.1016/j.scitotenv.2022.159284}{10.1016/j.scitotenv.2022.159284}.

\bibitem[\citeproctext]{ref-Lin2021}
Lin, Y., Moreno, C., Marchetti, A., Ducklow, H., Schofield, O., Delage,
E., Meredith, M., Li, Z., Eveillard, D., Chaffron, S., and Cassar, N.
2021. Decline in plankton diversity and carbon flux with reduced sea ice
extent along the {Western Antarctic Peninsula}. Nature Communications
\textbf{12}(1): 4948. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41467-021-25235-w}{10.1038/s41467-021-25235-w}.

\bibitem[\citeproctext]{ref-Llorca2012}
Llorca, M., Farré, M., Tavano, M.S., Alonso, B., Koremblit, G., and
Barceló, D. 2012. Fate of a broad spectrum of perfluorinated compounds
in soils and biota from {Tierra} del {Fuego} and {Antarctica}.
Environmental Pollution \textbf{163}: 158--166.
doi:\href{https://doi.org/10.1016/j.envpol.2011.10.027}{10.1016/j.envpol.2011.10.027}.

\bibitem[\citeproctext]{ref-Lopez-Gappa2022}
López-Gappa, J. 2022. The {Impact} of {Global Change} on {Marine Benthic
Invertebrates}. \emph{In} Global {Change} in {Atlantic Coastal
Patagonian Ecosystems}: {A Journey Through Time}. \emph{Edited by} E.W.
Helbling, M.A. Narvarte, R.A. González, and V.E. Villafañe. {Springer
International Publishing}, {Cham}. pp. 177--204.
doi:\href{https://doi.org/10.1007/978-3-030-86676-1_8}{10.1007/978-3-030-86676-1\_8}.

\bibitem[\citeproctext]{ref-Lopez-Lopez2022}
López-López, L., Genner, M.J., Tarling, G.A., Saunders, R.A., and
O'Gorman, E.J. 2022. Ecological {Networks} in the {Scotia Sea}:
{Structural Changes Across Latitude} and {Depth}. Ecosystems
\textbf{25}(2): 457--470.
doi:\href{https://doi.org/10.1007/s10021-021-00665-1}{10.1007/s10021-021-00665-1}.

\bibitem[\citeproctext]{ref-Marina2018}
Marina, T.I., Salinas, V., Cordone, G., Campana, G., Moreira, E.,
Deregibus, D., Torre, L., Sahade, R., Tatián, M., Barrera Oro, E., De
Troch, M., Doyle, S., Quartino, M.L., Saravia, L.A., and Momo, F.R.
2018. The {Food Web} of {Potter Cove} ({Antarctica}): Complexity,
structure and function. Estuarine, Coastal and Shelf Science
\textbf{200}: 141--151.
doi:\href{https://doi.org/10.1016/j.ecss.2017.10.015}{10.1016/j.ecss.2017.10.015}.

\bibitem[\citeproctext]{ref-Marinainrev}
Marina, T.I., Schloss, I.R., Lovrich, G.A., Boy, C.C., Bruno, D.O.,
Capitanio, F.L., Delpiani, S.M., Díaz de Astarloa, J.M., Fraysse, C.,
García Alonso, V.A., Raya Rey, A., Schejter, L., Spinelli, M., Tatián,
M., Urteaga, D., and Riccialdelli, L. in rev. The complex network of
trophic interactions in a {Subantarctic} oceanic {Marine Protected
Area}. Marine Ecology Progress Series.

\bibitem[\citeproctext]{ref-Martinez2022}
Martínez, P.A., Wöhler, O.C., Troccoli, G.H., Di Marco, E., and Maydana,
L. 2022. {Descripción de las capturas incidentales de granaderos
presentes en los lances dirigidos a merluza negra en la pesquería
argentina de arrastre: período 2010-2019.} {Instituto Nacional de
Investigación y Desarrollo Pesquero (INIDEP)}. {[}accessed 10 August
2023{]}.

\bibitem[\citeproctext]{ref-Matano2019}
Matano, R.P., Palma, E.D., and Combes, V. 2019. The {Burdwood Bank
Circulation}. Journal of Geophysical Research: Oceans \textbf{124}(10):
6904--6926.
doi:\href{https://doi.org/10.1029/2019JC015001}{10.1029/2019JC015001}.

\bibitem[\citeproctext]{ref-Matano2010}
Matano, R.P., Palma, E.D., and Piola, A.R. 2010. The influence of the
{Brazil} and {Malvinas Currents} on the {Southwestern Atlantic Shelf}
circulation. Ocean Science \textbf{6}(4): 983--995. {Copernicus GmbH}.
doi:\href{https://doi.org/10.5194/os-6-983-2010}{10.5194/os-6-983-2010}.

\bibitem[\citeproctext]{ref-Menge1995}
Menge, B.A. 1995. Indirect {Effects} in {Marine Rocky Intertidal
Interaction Webs}: {Patterns} and {Importance}. Ecological Monographs
\textbf{65}(1): 21--74.
doi:\href{https://doi.org/10.2307/2937158}{10.2307/2937158}.

\bibitem[\citeproctext]{ref-Meredith2018}
Meredith, M.P., Falk, U., Bers, A.V., Mackensen, A., Schloss, I.R., Ruiz
Barlett, E., Jerosch, K., Silva Busso, A., and Abele, D. 2018. Anatomy
of a glacial meltwater discharge event in an {Antarctic} cove.
Philosophical Transactions of the Royal Society A: Mathematical,
Physical and Engineering Sciences \textbf{376}(2122): 20170163.
doi:\href{https://doi.org/10.1098/rsta.2017.0163}{10.1098/rsta.2017.0163}.

\bibitem[\citeproctext]{ref-Michel2019}
Michel, L.N., Danis, B., Dubois, P., Eleaume, M., Fournier, J., Gallut,
C., Jane, P., and Lepoint, G. 2019. Increased sea ice cover alters food
web structure in {East Antarctica}. Scientific Reports \textbf{9}(1):
8062. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41598-019-44605-5}{10.1038/s41598-019-44605-5}.

\bibitem[\citeproctext]{ref-Mintenbeck2012}
Mintenbeck, K., Barrera-Oro, E.R., Brey, T., Jacob, U., Knust, R., Mark,
F.C., Moreira, E., Strobel, A., and Arntz, W.E. 2012. Impact of {Climate
Change} on {Fishes} in {Complex Antarctic Ecosystems}. \emph{In}
Advances in {Ecological Research}. \emph{Edited by} U. Jacob and G.
Woodward. {Academic Press}, {Burlington}. pp. 361--426.
doi:\href{https://doi.org/10.1016/B978-0-12-396992-7.00006-X}{10.1016/B978-0-12-396992-7.00006-X}.

\bibitem[\citeproctext]{ref-Montoya2009}
Montoya, JoséM., Woodward, G., Emmerson, M.C., and Solé, R.V. 2009.
Press perturbations and indirect effects in real food webs. Ecology
\textbf{90}(9): 2426--2433.
doi:\href{https://doi.org/10.1890/08-0657.1}{10.1890/08-0657.1}.

\bibitem[\citeproctext]{ref-Murphy2007}
Murphy, E.J., Trathan, P.N., Watkins, J.L., Reid, K., Meredith, M.P.,
Forcada, J., Thorpe, S.E., Johnston, N.M., and Rothery, P. 2007.
Climatically driven fluctuations in {Southern Ocean} ecosystems.
Proceedings of the Royal Society B: Biological Sciences
\textbf{274}(1629): 3057--3067. {Royal Society}.
doi:\href{https://doi.org/10.1098/rspb.2007.1180}{10.1098/rspb.2007.1180}.

\bibitem[\citeproctext]{ref-Murphy2006}
Murphy, E.j., Watkins, J.l., Trathan, P.n., Reid, K., Meredith, M.p.,
Thorpe, S.e., Johnston, N.m., Clarke, A., Tarling, G.a., Collins, M.a.,
Forcada, J., Shreeve, R.s., Atkinson, A., Korb, R., Whitehouse, M.j.,
Ward, P., Rodhouse, P.g., Enderlein, P., Hirst, A.g., Martin, A.r.,
Hill, S.l., Staniland, I.j., Pond, D.w., Briggs, D.r., Cunningham, N.j.,
and Fleming, A.h. 2006. Spatial and temporal operation of the {Scotia
Sea} ecosystem: A review of large-scale links in a krill centred food
web. Philosophical Transactions of the Royal Society B: Biological
Sciences \textbf{362}(1477): 113--148. {Royal Society}.
doi:\href{https://doi.org/10.1098/rstb.2006.1957}{10.1098/rstb.2006.1957}.

\bibitem[\citeproctext]{ref-Nardi2019}
Nardi, C.F., Fernández, D.A., Vanella, F.A., and Chalde, T. 2019. The
expansion of exotic {Chinook} salmon ({Oncorhynchus} tshawytscha) in the
extreme south of {Patagonia}: An environmental {DNA} approach.
Biological Invasions \textbf{21}(4): 1415--1425.
doi:\href{https://doi.org/10.1007/s10530-018-1908-8}{10.1007/s10530-018-1908-8}.

\bibitem[\citeproctext]{ref-Nilsson2016}
Nilsson, K.A., and McCann, K.S. 2016. Interaction strength
revisited\textemdash clarifying the role of energy flux for food web
stability. Theoretical Ecology \textbf{9}(1): 59--71.
doi:\href{https://doi.org/10.1007/s12080-015-0282-8}{10.1007/s12080-015-0282-8}.

\bibitem[\citeproctext]{ref-Nowacek2015}
Nowacek, D.P., Clark, C.W., Mann, D., Miller, P.J., Rosenbaum, H.C.,
Golden, J.S., Jasny, M., Kraska, J., and Southall, B.L. 2015. Marine
seismic surveys and ocean noise: Time for coordinated and prudent
planning. Frontiers in Ecology and the Environment \textbf{13}(7):
378--386. doi:\href{https://doi.org/10.1890/130286}{10.1890/130286}.

\bibitem[\citeproctext]{ref-OGorman2012}
O'Gorman, E.J., Fitch, J.E., and Crowe, T.P. 2012. Multiple
anthropogenic stressors and the structural properties of food webs.
Ecology \textbf{93}(3): 441--448.
doi:\href{https://doi.org/10.1890/11-0982.1}{10.1890/11-0982.1}.

\bibitem[\citeproctext]{ref-OGorman2019}
O'Gorman, E.J., Petchey, O.L., Faulkner, K.J., Gallo, B., Gordon,
T.A.C., Neto-Cerejeira, J., Ólafsson, J.S., Pichler, D.E., Thompson,
M.S.A., and Woodward, G. 2019. A simple model predicts how warming
simplifies wild food webs. Nature Climate Change \textbf{9}(8):
611--616. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41558-019-0513-x}{10.1038/s41558-019-0513-x}.

\bibitem[\citeproctext]{ref-Ojeda2021}
Ojeda, M., Cossi, P.F., Rimondino, G.N., Chiesa, I.L., Boy, C.C., and
Pérez, A.F. 2021. Microplastics pollution in the intertidal limpet,
{Nacella} magellanica, from {Beagle Channel} ({Argentina}). Science of
The Total Environment \textbf{795}: 148866.
doi:\href{https://doi.org/10.1016/j.scitotenv.2021.148866}{10.1016/j.scitotenv.2021.148866}.

\bibitem[\citeproctext]{ref-Oliver2018}
Oliver, E.C.J., Donat, M.G., Burrows, M.T., Moore, P.J., Smale, D.A.,
Alexander, L.V., Benthuysen, J.A., Feng, M., Sen Gupta, A., Hobday,
A.J., Holbrook, N.J., Perkins-Kirkpatrick, S.E., Scannell, H.A., Straub,
S.C., and Wernberg, T. 2018. Longer and more frequent marine heatwaves
over the past century. Nature Communications \textbf{9}(1): 1324.
{Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41467-018-03732-9}{10.1038/s41467-018-03732-9}.

\bibitem[\citeproctext]{ref-Olivier2019}
Olivier, P., Frelat, R., Bonsdorff, E., Kortsch, S., Kröncke, I.,
Möllmann, C., Neumann, H., Sell, A.F., and Nordström, M.C. 2019.
Exploring the temporal variability of a food web using long-term
biomonitoring data. Ecography \textbf{42}(12): 2107--2121.
doi:\href{https://doi.org/10.1111/ecog.04461}{10.1111/ecog.04461}.

\bibitem[\citeproctext]{ref-Orgeira2021}
Orgeira, J.L., Alvarez, F., and Salvó, C.S. 2021. The same pathway to
the {Weddell Sea} birdlife, after 65 years: Similarities in the species
composition, richness and abundances. Czech Polar Reports
\textbf{11}(2): 291--304.
doi:\href{https://doi.org/10.5817/CPR2021-2-20}{10.5817/CPR2021-2-20}.

\bibitem[\citeproctext]{ref-Orr2020}
Orr, J.A., Vinebrooke, R.D., Jackson, M.C., Kroeker, K.J., Kordas, R.L.,
Mantyka-Pringle, C., Van den Brink, P.J., De Laender, F., Stoks, R.,
Holmstrup, M., Matthaei, C.D., Monk, W.A., Penk, M.R., Leuzinger, S.,
Schäfer, R.B., and Piggott, J.J. 2020. Towards a unified study of
multiple stressors: Divisions and common goals across research
disciplines. Proceedings of the Royal Society B: Biological Sciences
\textbf{287}(1926): 20200421. {Royal Society}.
doi:\href{https://doi.org/10.1098/rspb.2020.0421}{10.1098/rspb.2020.0421}.

\bibitem[\citeproctext]{ref-Paine1996}
Paine, R.T., Ruesink, J.L., Sun, A., Soulanille, E.L., Wonham, M.J.,
Harley, C.D.G., Brumbaugh, D.R., and Secord, D.L. 1996. {TROUBLE ON
OILED WATERS}: {Lessons} from the {\emph{Exxon Valdez}} {Oil Spill}.
Annual Review of Ecology and Systematics \textbf{27}(1): 197--235.
doi:\href{https://doi.org/10.1146/annurev.ecolsys.27.1.197}{10.1146/annurev.ecolsys.27.1.197}.

\bibitem[\citeproctext]{ref-Pallin2023}
Pallin, L.J., Kellar, N.M., Steel, D., Botero-Acosta, N., Baker, C.S.,
Conroy, J.A., Costa, D.P., Johnson, C.M., Johnston, D.W., Nichols, R.C.,
Nowacek, D.P., Read, A.J., Savenko, O., Schofield, O.M., Stammerjohn,
S.E., Steinberg, D.K., and Friedlaender, A.S. 2023. A surplus no more?
{Variation} in krill availability impacts reproductive rates of
{Antarctic} baleen whales. Global Change Biology \textbf{29}(8):
2108--2121.
doi:\href{https://doi.org/10.1111/gcb.16559}{10.1111/gcb.16559}.

\bibitem[\citeproctext]{ref-Pante2023}
Pante, E., Simon-Bouhet, B., and Irisson, J. 2023. Marmap: {Import},
{Plot} and {Analyze Bathymetric} and {Topographic Data}.

\bibitem[\citeproctext]{ref-Pasotti2015}
Pasotti, F., Saravia, L.A., Troch, M.D., Tarantelli, M.S., Sahade, R.,
and Vanreusel, A. 2015. Benthic {Trophic Interactions} in an {Antarctic
Shallow Water Ecosystem Affected} by {Recent Glacier Retreat}. PLOS ONE
\textbf{10}(11): e0141742.
doi:\href{https://doi.org/10.1371/journal.pone.0141742}{10.1371/journal.pone.0141742}.

\bibitem[\citeproctext]{ref-Pasti2021}
Pasti, A.T., Bovcon, N.D., Ruibal-Núñez, J., Navoa, X., Jacobi, K.J.,
and Galván, D.E. 2021. The diet of {Mustelus} schmitti in areas with and
without commercial bottom trawling ({Central Patagonia}, {Southwestern
Atlantic}): {Is} it evidence of trophic interaction with the
{Patagonian} shrimp fishery? Food Webs \textbf{29}: e00214.
doi:\href{https://doi.org/10.1016/j.fooweb.2021.e00214}{10.1016/j.fooweb.2021.e00214}.

\bibitem[\citeproctext]{ref-Pauli2021}
Pauli, N.-C., Flintrop, C.M., Konrad, C., Pakhomov, E.A., Swoboda, S.,
Koch, F., Wang, X.-L., Zhang, J.-C., Brierley, A.S., Bernasconi, M.,
Meyer, B., and Iversen, M.H. 2021. Krill and salp faecal pellets
contribute equally to the carbon flux at the {Antarctic Peninsula}.
Nature Communications \textbf{12}(1): 7168. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41467-021-27436-9}{10.1038/s41467-021-27436-9}.

\bibitem[\citeproctext]{ref-Pecuchet2022}
Pecuchet, L., Jørgensen, L.L., Dolgov, A.V., Eriksen, E., Husson, B.,
Skern-Mauritzen, M., and Primicerio, R. 2022. Spatio-temporal turnover
and drivers of bentho-demersal community and food web structure in a
high-latitude marine ecosystem. Diversity and Distributions
\textbf{28}(12): 2503--2520.
doi:\href{https://doi.org/10.1111/ddi.13580}{10.1111/ddi.13580}.

\bibitem[\citeproctext]{ref-Pecuchet2020}
Pecuchet, L., Lindegren, M., Kortsch, S., Całkiewicz, J., Jurgensone,
I., Margonski, P., Otto, S.A., Putnis, I., Strāķe, S., and Nordström,
M.C. 2020. Spatio-temporal dynamics of multi-trophic communities reveal
ecosystem-wide functional reorganization. Ecography \textbf{43}(2):
197--208.
doi:\href{https://doi.org/10.1111/ecog.04643}{10.1111/ecog.04643}.

\bibitem[\citeproctext]{ref-Perez2020}
Pérez, A.F., Ojeda, M., Rimondino, G.N., Chiesa, I.L., Di Mauro, R.,
Boy, C.C., and Calcagno, J.A. 2020. First report of microplastics
presence in the mussel {Mytilus} chilensis from {Ushuaia Bay} ({Beagle
Channel}, {Tierra} del {Fuego}, {Argentina}). Marine Pollution Bulletin
\textbf{161}: 111753.
doi:\href{https://doi.org/10.1016/j.marpolbul.2020.111753}{10.1016/j.marpolbul.2020.111753}.

\bibitem[\citeproctext]{ref-Perry2020}
Perry, F.A., Kawaguchi, S., Atkinson, A., Sailley, S.F., Tarling, G.A.,
Mayor, D.J., Lucas, C.H., King, R., and Cooper, A. 2020.
Temperature\textendash{{Induced Hatch Failure}} and {Nauplii
Malformation} in {Antarctic Krill}. Frontiers in Marine Science
\textbf{7}. {[}accessed 11 August 2023{]}.

\bibitem[\citeproctext]{ref-Peterson2001}
Peterson, C.H. 2001. The {``{Exxon Valdez}''} oil spill in {Alaska}:
{Acute}, indirect and chronic effects on the ecosystem. \emph{In}
Advances in {Marine Biology}. {Academic Press}. pp. 1--103.
doi:\href{https://doi.org/10.1016/S0065-2881(01)39008-9}{10.1016/S0065-2881(01)39008-9}.

\bibitem[\citeproctext]{ref-Pineda-Metz2020}
Pineda-Metz, S.E.A., Gerdes, D., and Richter, C. 2020. Benthic fauna
declined on a whitening {Antarctic} continental shelf. Nature
Communications \textbf{11}(1): 2226. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/s41467-020-16093-z}{10.1038/s41467-020-16093-z}.

\bibitem[\citeproctext]{ref-Pinkerton2021}
Pinkerton, M.H., Boyd, P.W., Deppeler, S., Hayward, A., Höfer, J., and
Moreau, S. 2021. Evidence for the {Impact} of {Climate Change} on
{Primary Producers} in the {Southern Ocean}. Frontiers in Ecology and
Evolution \textbf{9}. {[}accessed 11 September 2023{]}.

\bibitem[\citeproctext]{ref-Poloczanska2013}
Poloczanska, E.S., Brown, C.J., Sydeman, W.J., Kiessling, W., Schoeman,
D.S., Moore, P.J., Brander, K., Bruno, J.F., Buckley, L.B., Burrows,
M.T., Duarte, C.M., Halpern, B.S., Holding, J., Kappel, C.V., O'Connor,
M.I., Pandolfi, J.M., Parmesan, C., Schwing, F., Thompson, S.A., and
Richardson, A.J. 2013. Global imprint of climate change on marine life.
Nature Climate Change \textbf{3}(10): 919--925. {Nature Publishing
Group}.
doi:\href{https://doi.org/10.1038/nclimate1958}{10.1038/nclimate1958}.

\bibitem[\citeproctext]{ref-Quartino2008}
Quartino, M.L., and Boraso de Zaixso, A.L. 2008. Summer macroalgal
biomass in {Potter Cove}, {South Shetland Islands}, {Antarctica}: Its
production and flux to the ecosystem. Polar Biology \textbf{31}(3):
281--294.
doi:\href{https://doi.org/10.1007/s00300-007-0356-1}{10.1007/s00300-007-0356-1}.

\bibitem[\citeproctext]{ref-Quartino2013}
Quartino, M.L., Deregibus, D., Campana, G.L., Latorre, G.E.J., and Momo,
F.R. 2013. Evidence of {Macroalgal Colonization} on {Newly Ice-Free
Areas} following {Glacial Retreat} in {Potter Cove} ({South Shetland
Islands}), {Antarctica}. PLoS ONE \textbf{8}(3): e58223.
doi:\href{https://doi.org/10.1371/journal.pone.0058223}{10.1371/journal.pone.0058223}.

\bibitem[\citeproctext]{ref-Raymond2011}
Raymond, B. 2011. A circumpolar pelagic regionalisation of the {Southern
Ocean}. {CCAMLR WS-MPA}, {Brest}, {France}, 29 {Aug}\textendash 2 {Sep}
2011. {Document WS-MPA-11}/6. {http://data.aad.gov.au/regionalisation}.

\bibitem[\citeproctext]{ref-Riccialdelli2020}
Riccialdelli, L., Becker, Y.A., Fioramonti, N.E., Torres, M., Bruno,
D.O., Rey, A.R., and Fernández, D.A. 2020. Trophic structure of southern
marine ecosystems: A comparative isotopic analysis from the {Beagle
Channel} to the oceanic {Burdwood Bank} area under a wasp-waist
assumption. Marine Ecology Progress Series \textbf{655}: 1--27.
doi:\href{https://doi.org/10.3354/meps13524}{10.3354/meps13524}.

\bibitem[\citeproctext]{ref-Riccialdelli2013}
Riccialdelli, L., Newsome, S.D., Dellabianca, N.A., Bastida, R., Fogel,
M.L., and Goodall, R.N.P. 2013. Ontogenetic diet shift in {Commerson}'s
dolphin ({Cephalorhynchus} commersonii commersonii) off {Tierra} del
{Fuego}. Polar Biology \textbf{36}(5): 617--627.
doi:\href{https://doi.org/10.1007/s00300-013-1289-5}{10.1007/s00300-013-1289-5}.

\bibitem[\citeproctext]{ref-Rincon-Diaz2021}
Rincón-Díaz, M.P., Bovcon, N.D., Cochia, P.D., Góngora, M.E., and
Galván, D.E. 2021. Fish functional diversity as an indicator of
resilience to industrial fishing in {Patagonia Argentina}. Journal of
Fish Biology \textbf{99}(5): 1650--1667.
doi:\href{https://doi.org/10.1111/jfb.14873}{10.1111/jfb.14873}.

\bibitem[\citeproctext]{ref-Rodriguez2022}
Rodriguez, I.D., Marina, T.I., Schloss, I.R., and Saravia, L.A. 2022.
Marine food webs are more complex but less stable in sub-{Antarctic}
({Beagle Channel}, {Argentina}) than in {Antarctic} ({Potter Cove},
{Antarctic Peninsula}) regions. Marine Environmental Research
\textbf{174}: 105561.
doi:\href{https://doi.org/10.1016/j.marenvres.2022.105561}{10.1016/j.marenvres.2022.105561}.

\bibitem[\citeproctext]{ref-Rowlands2021}
Rowlands, E., Galloway, T., Cole, M., Lewis, C., Peck, V., Thorpe, S.,
and Manno, C. 2021. The {Effects} of {Combined Ocean Acidification} and
{Nanoplastic Exposures} on the {Embryonic Development} of {Antarctic
Krill}. Frontiers in Marine Science \textbf{8}. {[}accessed 11 August
2023{]}.

\bibitem[\citeproctext]{ref-Ruckamp2011}
Rückamp, M., Braun, M., Suckro, S., and Blindow, N. 2011. Observed
glacial changes on the {King George Island} ice cap, {Antarctica}, in
the last decade. Global and Planetary Change \textbf{79}(1-2): 99--109.
doi:\href{https://doi.org/10.1016/j.gloplacha.2011.06.009}{10.1016/j.gloplacha.2011.06.009}.

\bibitem[\citeproctext]{ref-RuibalNunez2020}
Ruibal Nuñez, J. 2020. {Impacto ecológico de la actividad pesquera en
las poblaciones de condrictios en el litoral de la Provincia de Chubut y
Golfo San Jorge}. PhD thesis, Universidad Nacional de Mar del Plata
(UNMdP), {Argentina}.

\bibitem[\citeproctext]{ref-Sahade2015}
Sahade, R., Lagger, C., Torre, L., Momo, F., Monien, P., Schloss, I.,
Barnes, D.K.A., Servetto, N., Tarantelli, S., Tatián, M., Zamboni, N.,
and Abele, D. 2015. Climate change and glacier retreat drive shifts in
an {Antarctic} benthic ecosystem. Science Advances \textbf{1}(10):
e1500050. {American Association for the Advancement of Science}.
doi:\href{https://doi.org/10.1126/sciadv.1500050}{10.1126/sciadv.1500050}.

\bibitem[\citeproctext]{ref-Saunders2019}
Saunders, R.A., Hill, S.L., Tarling, G.A., and Murphy, E.J. 2019.
Myctophid {Fish} ({Family Myctophidae}) {Are Central Consumers} in the
{Food Web} of the {Scotia Sea} ({Southern Ocean}). Frontiers in Marine
Science \textbf{6}. {[}accessed 11 August 2023{]}.

\bibitem[\citeproctext]{ref-Schejter2021}
Schejter, L., and Albano, M. 2021. Benthic communities at the marine
protected area {Namuncurá}/{Burdwood} bank, {SW Atlantic Ocean}:
Detection of vulnerable marine ecosystems and contributions to the
assessment of the rezoning process. Polar Biology \textbf{44}(10):
2023--2037.
doi:\href{https://doi.org/10.1007/s00300-021-02936-y}{10.1007/s00300-021-02936-y}.

\bibitem[\citeproctext]{ref-Schloss2012}
Schloss, I.R., Abele, D., Moreau, S., Demers, S., Bers, A.V., González,
O., and Ferreyra, G.A. 2012. Response of phytoplankton dynamics to
19-year (1991\textendash 2009) climate trends in {Potter Cove}
({Antarctica}). Journal of Marine Systems \textbf{92}(1): 53--66.
doi:\href{https://doi.org/10.1016/j.jmarsys.2011.10.006}{10.1016/j.jmarsys.2011.10.006}.

\bibitem[\citeproctext]{ref-Schloss2023}
Schloss, I.R., Pizarro, G., Cadaillon, A.M., Giesecke, R., Hernando,
M.P., Almandoz, G.O., Latorre, M.P., Malits, A., Flores-Melo, X.,
Saravia, L.A., Martín, J., Guzmán, L., Iachetti, C.M., and Ruiz, C.
2023. Alexandrium catenella dynamics and paralytic shellfish toxins
distribution along the {Beagle Channel} (southern {Patagonia}). Journal
of Marine Systems \textbf{239}: 103856.
doi:\href{https://doi.org/10.1016/j.jmarsys.2022.103856}{10.1016/j.jmarsys.2022.103856}.

\bibitem[\citeproctext]{ref-Schmidt2011}
Schmidt, K., Atkinson, A., Steigenberger, S., Fielding, S., Lindsay,
M.C.M., Pond, D.W., Tarling, G.A., Klevjer, T.A., Allen, C.S., Nicol,
S., and Achterberg, E.P. 2011. Seabed foraging by {Antarctic} krill:
{Implications} for stock assessment, bentho-pelagic coupling, and the
vertical transfer of iron. Limnology and Oceanography \textbf{56}(4):
1411--1428.
doi:\href{https://doi.org/10.4319/lo.2011.56.4.1411}{10.4319/lo.2011.56.4.1411}.

\bibitem[\citeproctext]{ref-Seco2021}
Seco, J., Aparício, S., Brierley, A.S., Bustamante, P., Ceia, F.R.,
Coelho, J.P., Philips, R.A., Saunders, R.A., Fielding, S., Gregory, S.,
Matias, R., Pardal, M.A., Pereira, E., Stowasser, G., Tarling, G.A., and
Xavier, J.C. 2021. Mercury biomagnification in a {Southern Ocean} food
web. Environmental Pollution \textbf{275}: 116620.
doi:\href{https://doi.org/10.1016/j.envpol.2021.116620}{10.1016/j.envpol.2021.116620}.

\bibitem[\citeproctext]{ref-Sgarlatta2023}
Sgarlatta, M. 2023. Mechanisms driving tropicalization of temperate reef
fish assemblages. Thesis, UNSW Sydney.
doi:\href{https://doi.org/10.26190/unsworks/25104}{10.26190/unsworks/25104}.

\bibitem[\citeproctext]{ref-Smale2008}
Smale, D.A., and Barnes, D.K.A. 2008. Likely responses of the
{Antarctic} benthos to climate-related changes in physical disturbance
during the 21st century, based primarily on evidence from the {West
Antarctic Peninsula} region. Ecography \textbf{31}(3): 289--305.
{Blackwell Publishing Ltd}.
doi:\href{https://doi.org/10.1111/j.0906-7590.2008.05456.x}{10.1111/j.0906-7590.2008.05456.x}.

\bibitem[\citeproctext]{ref-Smale2008a}
Smale, D.A., Brown, K.M., Barnes, D.K.A., Fraser, K.P.P., and Clarke, A.
2008. Ice {Scour Disturbance} in {Antarctic Waters}. Science
\textbf{321}(5887): 371--371. {American Association for the Advancement
of Science}.
doi:\href{https://doi.org/10.1126/science.1158647}{10.1126/science.1158647}.

\bibitem[\citeproctext]{ref-Soeffker2022}
Soeffker, M., Hollyman, P.R., Collins, M.A., Hogg, O.T., Riley, A.,
Laptikhovsky, V., Earl, T., Roberts, J., MacLeod, E., Belchier, M., and
Darby, C. 2022. Contrasting life-history traits of two toothfish
({Dissostichus} spp.) species at their range edge around the {South
Sandwich Islands}. Deep Sea Research Part II: Topical Studies in
Oceanography \textbf{201}: 105098.
doi:\href{https://doi.org/10.1016/j.dsr2.2022.105098}{10.1016/j.dsr2.2022.105098}.

\bibitem[\citeproctext]{ref-Stouffer2011}
Stouffer, D.B., and Bascompte, J. 2011. Compartmentalization increases
food-web persistence. Proceedings of the National Academy of Sciences
\textbf{108}(9): 3648--3652. {Proceedings of the National Academy of
Sciences}.
doi:\href{https://doi.org/10.1073/pnas.1014353108}{10.1073/pnas.1014353108}.

\bibitem[\citeproctext]{ref-Tamini2023}
Tamini, L.L., Dellacasa, R.F., Chavez, L.N., Marinao, C.J., Góngora,
M.E., Crawford, R., and Frere, E. 2023. Bird scaring lines reduce
seabird mortality in mid-water and bottom trawlers in {Argentina}. ICES
Journal of Marine Science: fsad109.
doi:\href{https://doi.org/10.1093/icesjms/fsad109}{10.1093/icesjms/fsad109}.

\bibitem[\citeproctext]{ref-Tarling2018}
Tarling, G.A., Ward, P., and Thorpe, S.E. 2018. Spatial distributions of
{Southern Ocean} mesozooplankton communities have been resilient to
long-term surface warming. Global Change Biology \textbf{24}(1):
132--142.
doi:\href{https://doi.org/10.1111/gcb.13834}{10.1111/gcb.13834}.

\bibitem[\citeproctext]{ref-Teixido2002}
Teixidó, N., Garrabou, J., and Arntz, W.E. 2002. Spatial pattern
quantification of {Antarctic} benthic communities using landscape
indices. Marine Ecology Progress Series \textbf{242}: 1--14.
doi:\href{https://doi.org/10.3354/meps242001}{10.3354/meps242001}.

\bibitem[\citeproctext]{ref-Thompson2012}
Thompson, R.M., Brose, U., Dunne, J.A., Hall, R.O., Hladyz, S.,
Kitching, R.L., Martinez, N.D., Rantala, H., Romanuk, T.N., Stouffer,
D.B., and Tylianakis, J.M. 2012. Food webs: Reconciling the structure
and function of biodiversity. Trends in Ecology \& Evolution
\textbf{27}(12): 689--697. {Elsevier}.
doi:\href{https://doi.org/10.1016/j.tree.2012.08.005}{10.1016/j.tree.2012.08.005}.

\bibitem[\citeproctext]{ref-Thompson2007}
Thompson, R.M., Hemberg, M., Starzomski, B.M., and Shurin, J.B. 2007.
Trophic {Levels} and {Trophic Tangles}: {The Prevalence} of {Omnivory}
in {Real Food Webs}. Ecology \textbf{88}(3): 612--617.
doi:\href{https://doi.org/10.1890/05-1454}{10.1890/05-1454}.

\bibitem[\citeproctext]{ref-Trathan2023}
Trathan, P.N. 2023. The future of the {South Georgia} and {South
Sandwich Islands} marine protected area in a changing environment: {The}
choice between industrial fisheries, or ecosystem protection. Marine
Policy \textbf{155}: 105773.
doi:\href{https://doi.org/10.1016/j.marpol.2023.105773}{10.1016/j.marpol.2023.105773}.

\bibitem[\citeproctext]{ref-Trathan2021}
Trathan, P.N., Fielding, S., Hollyman, P.R., Murphy, E.J.,
Warwick-Evans, V., and Collins, M.A. 2021. Enhancing the ecosystem
approach for the fishery for {Antarctic} krill within the complex,
variable, and changing ecosystem at {South Georgia}. ICES Journal of
Marine Science \textbf{78}(6): 2065--2081.
doi:\href{https://doi.org/10.1093/icesjms/fsab092}{10.1093/icesjms/fsab092}.

\bibitem[\citeproctext]{ref-Turner2020}
Turner, J., Guarino, M.V., Arnatt, J., Jena, B., Marshall, G.J.,
Phillips, T., Bajish, C.C., Clem, K., Wang, Z., Andersson, T., Murphy,
E.J., and Cavanagh, R. 2020. Recent {Decrease} of {Summer Sea Ice} in
the {Weddell Sea}, {Antarctica}. Geophysical Research Letters
\textbf{47}(11): e2020GL087127.
doi:\href{https://doi.org/10.1029/2020GL087127}{10.1029/2020GL087127}.

\bibitem[\citeproctext]{ref-Verges2019}
Vergés, A., McCosker, E., Mayer-Pinto, M., Coleman, M.A., Wernberg, T.,
Ainsworth, T., and Steinberg, P.D. 2019. Tropicalisation of temperate
reefs: {Implications} for ecosystem functions and management actions.
Functional Ecology \textbf{33}(6): 1000--1013.
doi:\href{https://doi.org/10.1111/1365-2435.13310}{10.1111/1365-2435.13310}.

\bibitem[\citeproctext]{ref-Verges2014}
Vergés, A., Steinberg, P.D., Hay, M.E., Poore, A.G.B., Campbell, A.H.,
Ballesteros, E., Heck, K.L., Booth, D.J., Coleman, M.A., Feary, D.A.,
Figueira, W., Langlois, T., Marzinelli, E.M., Mizerek, T., Mumby, P.J.,
Nakamura, Y., Roughan, M., van Sebille, E., Gupta, A.S., Smale, D.A.,
Tomas, F., Wernberg, T., and Wilson, S.K. 2014. The tropicalization of
temperate marine ecosystems: Climate-mediated changes in herbivory and
community phase shifts. Proceedings of the Royal Society B: Biological
Sciences \textbf{281}(1789): 20140846. {Royal Society}.
doi:\href{https://doi.org/10.1098/rspb.2014.0846}{10.1098/rspb.2014.0846}.

\bibitem[\citeproctext]{ref-Vinuesa2005}
Vinuesa, J.H. 2005. Distribución de crustáceos decápodos y estomatópodos
del golfo {San Jorge}, {Argentina}. Revista de biología marina y
oceanografía \textbf{40}(1): 7--21. {Universidad de Valparaíso. Facultad
de Ciencias del Mar}.
doi:\href{https://doi.org/10.4067/S0718-19572005000100002}{10.4067/S0718-19572005000100002}.

\bibitem[\citeproctext]{ref-Vinuesa2007}
Vinuesa, J.H., and Varisco, M. 2007. Trophic ecology of the lobster
krill {Munida} gregaria in {San Jorge Gulf}, {Argentina}.
Investigaciones marinas \textbf{35}(2): 25--34. {Escuela de Ciencias del
Mar. Pontificia Universidad Católica de Valparaíso}.
doi:\href{https://doi.org/10.4067/S0717-71782007000200003}{10.4067/S0717-71782007000200003}.

\bibitem[\citeproctext]{ref-Vucic-Pestic2011}
Vucic-Pestic, O., Ehnes, R.B., Rall, B.C., and Brose, U. 2011. Warming
up the system: Higher predator feeding rates but lower energetic
efficiencies. Global Change Biology \textbf{17}(3): 1301--1310.
doi:\href{https://doi.org/10.1111/j.1365-2486.2010.02329.x}{10.1111/j.1365-2486.2010.02329.x}.

\bibitem[\citeproctext]{ref-Ward2018}
Ward, P., Tarling, G.A., and Thorpe, S.E. 2018. Temporal changes in
abundances of large calanoid copepods in the {Scotia Sea}: Comparing the
1930s with contemporary times. Polar Biology \textbf{41}(11):
2297--2310.
doi:\href{https://doi.org/10.1007/s00300-018-2369-3}{10.1007/s00300-018-2369-3}.

\bibitem[\citeproctext]{ref-Wege2021}
Wege, M., Salas, L., and LaRue, M. 2021. Ice matters: {Life-history}
strategies of two {Antarctic} seals dictate climate change eventualities
in the {Weddell Sea}. Global Change Biology \textbf{27}(23): 6252--6262.
doi:\href{https://doi.org/10.1111/gcb.15828}{10.1111/gcb.15828}.

\bibitem[\citeproctext]{ref-Whitehouse2008}
Whitehouse, M.J., Meredith, M.P., Rothery, P., Atkinson, A., Ward, P.,
and Korb, R.E. 2008. Rapid warming of the ocean around {South Georgia},
{Southern Ocean}, during the 20th century: {Forcings}, characteristics
and implications for lower trophic levels. Deep Sea Research Part I:
Oceanographic Research Papers \textbf{55}(10): 1218--1228.
doi:\href{https://doi.org/10.1016/j.dsr.2008.06.002}{10.1016/j.dsr.2008.06.002}.

\bibitem[\citeproctext]{ref-Whitworth1980}
Whitworth, T. 1980. Zonation and geostrophic flow of the {Antarctic}
circumpolar current at {Drake Passage}. Deep Sea Research Part A.
Oceanographic Research Papers \textbf{27}(7): 497--507.
doi:\href{https://doi.org/10.1016/0198-0149(80)90036-9}{10.1016/0198-0149(80)90036-9}.

\bibitem[\citeproctext]{ref-Williams2000}
Williams, R.J., and Martinez, N.D. 2000. Simple rules yield complex food
webs. Nature \textbf{404}(6774): 180--183. {Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/35004572}{10.1038/35004572}.

\bibitem[\citeproctext]{ref-Wood2015}
Wood, S.A., Russell, R., Hanson, D., Williams, R.J., and Dunne, J.A.
2015. Effects of spatial scale of sampling on food web structure.
Ecology and Evolution \textbf{5}(17): 3769--3782.
doi:\href{https://doi.org/10.1002/ece3.1640}{10.1002/ece3.1640}.

\bibitem[\citeproctext]{ref-Woods2023}
Woods, B.L., Van de Putte, A.P., Hindell, M.A., Raymond, B., Saunders,
R.A., Walters, A., and Trebilco, R. 2023. Species distribution models
describe spatial variability in mesopelagic fish abundance in the
{Southern Ocean}. Frontiers in Marine Science \textbf{9}. {[}accessed 16
September 2023{]}.

\bibitem[\citeproctext]{ref-Wu2012}
Wu, L., Cai, W., Zhang, L., Nakamura, H., Timmermann, A., Joyce, T.,
McPhaden, M.J., Alexander, M., Qiu, B., Visbeck, M., Chang, P., and
Giese, B. 2012. Enhanced warming over the global subtropical western
boundary currents. Nature Climate Change \textbf{2}(3): 161--166.
{Nature Publishing Group}.
doi:\href{https://doi.org/10.1038/nclimate1353}{10.1038/nclimate1353}.

\bibitem[\citeproctext]{ref-Yorio2009}
Yorio, P. 2009. Marine protected areas, spatial scales, and governance:
Implications for the conservation of breeding seabirds. Conservation
Letters \textbf{2}(4): 171--178.
doi:\href{https://doi.org/10.1111/j.1755-263X.2009.00062.x}{10.1111/j.1755-263X.2009.00062.x}.

\end{CSLReferences}

\end{document}